### 2008-05-24 paper/paper.tex
paper

\documentclass{article}
\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{url}
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\title{A logic-based framework for Digital Signal Processing}

\author{...}

\maketitle

\begin{abstract}
In this paper we describe a logic-based framework for Digital Signal Processing. This framework is implemented on SWI-Prolog and is formed of interfaces with different DSP libraries and APIs extending this logic language ``to speak'' DSP. This approach presents the advantage, in respect of other integrated APIs, of being application implementation independent, so the extended logic language can be loaded an embedded in any application or language interfacing SWI-Prolog. The new DSP capabilities are public as predicates grouped in modules which offer different levels of abstraction. Low-level modules offer predicates to communicate directly with the API. Task-oriented modules define some default routines which specific goals and procedures. Finally we can relate input/output through ``black boxes'' representing a semantic relationship between them. Modules can communicate by using commong semantic data types declared as functors which determine the possible workflows. Such a framework allow us to build complex DSP queries as procedures as well as simplify the extraction of audio processing knowledge with common semantics. Finally, Concurrent Transaction Logic allows us to wrap the procedural meaning of predicate rules into a declarative definition of transactions controlling the execution and updating of the processes.
\emph{abstract} environment.

\end{abstract}

\section{Introduction}

Digital signal processing is a broad paradigm and object of an important research. Within this community there are many particular investigators discovering and developing new algorithms for signal processing and analysis. Many projects have been directed to develop formal APIs providing a framework for algorithms implementation and publication. However these APIs are normally designed as a close environment which requires a specific developer knowledge or as a ``plugin'' library which requires the implementation of a specific host.

This situation leads to many isolated applications, environments and sources of data which work independently from each other. Any effort to combine them to obtain integrated results for analysis, classification and comparison requires tedious studies or ``glue code'' implementations.

In this paper we address this problem by proposing and developing SWI-DSP, a logical framework for \textit{Audio Processing} that extends the SWI-Prolog vocabulary. We rely on a logic programming language with audio analysis capabilities getting rid of the application-dependency paradigm and offering a single logic front-end which is generally more understandable than a closed implementation in any Procedural or Object Oriented language. In addition to this, we are keen on using logic resources from SWI-Prolog to obtain finally a more capable DSP engine.

In \S 2 we summarize and evaluate the alternative APIs for audio analysis. Some of these APIs are hosted in SWI-DSP through the foreign interface providing an open system for developers as well. In \S 3 we start viewing the system from the top and how it defines an interesting workflow which actually solves the most part of the problems of the DSP paradigm. This workflow is determined by semantic types, a binary data standard and procedural declarations of riles. \S4 goes into detail in each task-oriented module describing the communication layers with the APIs through control predicates which require an understanding of the API.

Thus, SWI-DSP extends SWI-Prolog vocabulary to perform audio analysis and processing by defining semantic types and predicates, triggering external computation performed by dedicated APIs, which are ordered in modules with different levels of abstraction. However, this front-end still requires specific SWI-Prolog knowledge to create programs suitable for the concrete needs and it does not offer any efficient tabling mechanism. \S 5 explains how Concurrent Transaction Logic is the solution to provide a better logic front-end combining procedural and declarative meaning of transactions which also comprise data updating. In \S 6 we present related works as alternatives to manage SWI-DSP featuring another user interface than SWI-Prolog. We conclude in \S 7 by assessing the value and scope of this work and introducing future advances.

\section{Audio Analysis tasks and solutions}

Audio analysis  APIs offer formal implementations for development and use of different algorithms and techniques. These algorithms are focused on different tasks that can be classified in these 3 groups:

\begin{itemize}
 \item Processing
 \item Feature Extraction
 \item Classification
\end{itemize}

\subsection{Audio Processing}

This tasks group can be characterized by having audio signals as input and output, so an input signal is processed in such a way that we retrieve an altered, processed version of the original signal. Some examples of audio processing tasks are filtering, effects processing and signal generation.

One famous API is the Linux Audio Developer's Simple Plugin API (LADSPA) which has been succesfully used in many systems and applications and is a widely spread implementation for algorithms development.

\subsection{Feature Extraction}

Feature extraction is an important paradigm in signal research as it focuses on extract qualitative characteristics of the signal in a different domain than the input. Features are classified in different levels of abstraction from frequency characteristics to high-level audio ones.

The Vamp Plugins API offers another standard and utilities set to develop feature extraction plugins that can be hosted in any application able to control the ``plugin lifecycle''.

\subsection{Classification}

From different features we are able to classify different signals.

\subsection{Integrative solutions}

With the same motivation of this work there are some integrative APIs which combines different tasks to perform high-level operations in a closed framework. In the particular domain of Music jMIR and Marsyas present a complex framework for audio analysis and classification.

\section{SWI-DSP workflow}

We find of high interest the workflow architecture for DSP tasks. A complex task is then subdivided into subgoals that are achieved for the specific task-oriented modules. A procedure of this nature can be easily implemented combining high-level predicates with a proper matching of the instantation patterns. These instantation patters must have terms in common which are easily carried forward from one module to another without worring about further conversion. For this purpose, we define semantic datatypes as SWI-Prolog functors. In the same sense, we need to define a standard format for large binary data objects which are highly common in DSP tasks. We define these utils in the \textbf{swilib} module which provides a C++ library and a SWI-Prolog module for their management.

\subsection{Semantic datatypes}

SWI-Prolog functors are defined by their name and their arity. Therefore we need to recognize and set the common concepts and their properties used in audio analysis. In the previous section we have seen how tasks are strongly characterized by the type of input and output. We also want to look at the Music Ontology specification [x] as ontological description of music-related concepts including audio analysis terminology. We distinguish the following semantic datatypes:

\begin{itemize}
 \item 'Signal'(Channels, SampleRate, Samples, [ListOfPCMPerChannel]): any audio data fragment is classified as a signal. Audio data is presented in our system as Pulse Code Modulation (PCM) which is a digital version of an analog signal sampled in equal intervals. Necessary parameters to understand the signal are the number of channels, the sample rate of the sampling and the resulting number of samples.

\item 'Frame'(Channels, SampleRate, StartingSample, [ListOfPCMPerChannel]) as pieces of a signal, frames have an important role in DSP as we need to decompose the signal in smaller frames for an efficient analysis.

\item 'Feature'(Type, Timestamp, Data): extracted characteristics of a signal in some other domain are called features which are the main target of audio analysis at a low level. A feature term is characterized by the feature type, its interval of time and the actual feature data.

\item 'Timestamp'(Start, Duration): audio properties are always necessarily bound to an interval of time defined as a timestamp in respect of the audio signal ``timeline''.

\item 'Parameter'(Name, Value): parametrization of algorithms is one of the most significant issues in DSP as well. Algorithms vary their perfomance depending on the parameters. We want to generalize the parameterization for any task with generic terms.
\end{itemize}

\subsection{Large Binary Objects}

The so called BLOBs are objects wrapping large binary data. The representation of data in SWI-Prolog can be done through lists which can not handle such amount of data (e.g. signals of 1000000 samples) or through BLOB terms which are cumbersome as datatype.

We have implemented a mechanism of \textit{data identifiers} that vary incrementally and are stored in a database updated in each session. These identifiers, with the form \textbf{data\_x}, hide the binary data (which is internally a vector of floats) and allow us to perform different operations over the data easily with prolog predicates: \textit{data\_out(+Id, +File), data\_in(+File, +Id), data\_to\_list(+Id, -List)...} as well as control the binary data database status. Each module is implemented to read/write data in this fashion.

\subsection{Workflow}

These standards allow us to work with different modules without taking care of the specific underlying APIs types or implementation, so that we achieve a logic framework with dedicated semantics to perform audio analysis tasks as showed in \ref{fig:workflow}.

\begin{figure}
\centerline{\framebox{\includegraphics[width=\columnwidth]{workflow.png}}}
\caption{Workflow specification of the modules}
\label{fig:workflow}
\end{figure}

We can then execute the following task as a SWI-Prolog query. \textit{The input audio file is decoded and then we apply a pre-processing in which we amplify the signal. We then apply a key detection algorithm over the signal to retrieve tonal regions}

\begin{verbatim}

my_task(Input, ['Parameter'(gain, 2)], KeyProfile):-
	decode(InputFile, Signal),
	amplify(Signal, 'Parameter'(gain, 2), SignalX2),
	key(SignalX2, KeyProfile).

KeyProfile = ['Feature'(key, 'Timestamp'(0, 23.45), 'Amajor`')|...].

\end{verbatim}

\section{SWI-DSP modules}

Each task module is composed by lower level modules which offer communication with an external API or library in charge of the computations. By hosting APIs we can extend easily SWI-DSP vocabulary as we do not need to assemble each algorithm.

\subsection{swiaudio}

This module is in charge of providing audio data to the system. This module is splitted in two sub-modules: \textbf{swiaudiosource} and \textbf{swiaudiodata}. The former is a powerful module for decoding which interacts with C decoding libraries (mad, soundfile, fishsound, oggz and faad) to extract a \textit{Signal} term given and audio input file supported by the libraries. This process is triggered by a simple call to the predicate \textit{aspl\_decode(+InputFile, -Signal)} which recognizes the file extension and calls a specific library allowing a fast decoding process and even more important an independent term representing the encoded data in a standard way for any format.

The latter is a module which defines utilities typically required in DSP. The most important set of predicates is dedicated to framing. The original signal is decomposed in overlapping blocks of data outputting frame terms through the non-deterministic predicate \textit{framing(+Signal, +StepSize, +BlockSize, -Frame)}. The rest of predicates are mainly oriented to extract information from a Signal or Frame term.

\subsection{feature extraction}

This module defines a standard task-oriented set of predicates for feature extraction. As we have seen in \S 2 the module is characterized for a clear input/output relationship. So we can declare a simple predicate to extract the tempo of an audio signal: tempo(+Input, -Tempo) being Input an instantation of the Signal datatype and Tempo a list of instantations of the Feature datatype. This ``magic'' predicate is triggering a specific computation tool, so we could select the specific tool by extending the predicate: \textit{tempo(+Input, +Tool, -Tempo}.

We rely on the Vamp Plugin API as it offers an efficient plugin and host implementation. Certainly, there is a limitation in the sense that any feature extraction algorithm we may want to use has to be implemented as Vamp Plugin, but as counterpart it is easier to write an algorithm as Vamp Plugin than to interface many APIs or particular algorithm implementations.

The sub-module \textbf{swivamp} is a communication layer based on the SWI-Prolog foreign interface which allow us to query the plugin information, set up the plugin context for the transform, execute the plugin given a Signal term and retrieve the output features for the plugin according to the lifecycle specification. Plugins are wrapped in a similar fashion than the audio data so we can use a plugin instance representation which allow us to have concurrent plugins working at the same time. The previous tempo/3 predicate is expressed as this set of rules:

\begin{verbatim}

tempo(Input, Tool, Tempo):-
        Input = 'Signal'(Ch, Sr, L, PCMs),
        Tool = 'Vamp',
        vamp_feature_of('tempo', Signal, WholeFeature).
vamp_feature_of(Type, Signal, WholeFeature):-
        select_plugin(Type, PluginKey, Output),
        vamp_outputs_for(Signal, PluginKey, Output, Features),
        flatten(Features, WholeFeature).
vamp_outputs_for(Signak, PluginKey, Output, Feature):-
        vmpl_load_plugin_for(PluginKey, Signal, Plugin),
        set_blockSize(Plugin, BlockSize),
        set_stepSize(Plugin, StepSize),
        get_channels(Signal, Channels),
        vmpl_initialize_plugin(Plugin, Channels, StepSize, BlockSize),
        vamp_compute_feature(Signal, StepSize, BlockSize, Outputs, Plugin, WholeFeature).
vamp_compute_feature(Signal, StepSize, BlockSize, Outputs, Plugin, WholeFeature):-
        get_samples_per_channel(Signal, L),
        findall(Features, vamp_process_signal(Signal, L, StepSize, BlockSize, Output, Plugin, Features), FeatureSet),
        get_sample_rate(Signal, SampleRate),
        vmpl_remaining_features(Plugin, L, SampleRate, Output, Remaining),
        append(FeatureSet, Remaining, RawFeatures),
        delete(RawFeatures, [], WholeFeature).
vamp_process_signal(Signal, L, StepSize, BlockSize, Output, Plugin, Features):-
        set_limit_framing(L, StepSize, Limit),
        set_framing(StepSize, L, Limit, Start),
        vmpl_process_block_framing(Plugin, Signal, Start, BlockSize, Output, FeatureSet).

\end{verbatim}

\subsection{audio processing}

This module defines any computation which outputs a Signal term. This

The module swiladspa offers communication with LADSPA which is the most generic and spread plugin API for general audio processing.

\subsection{swiweka}

\section{Towards a computation reasoner}

So far we have presented how SWI-DSP provides a vocabulary of logic predicates to perform audio analysis computations, but this logic front-end is still inappropiate for such tasks. We have organized the modules in communication layers and task-oriented modules, but there is an important trade-off between understandability and specification. The communication layer even offering an application-independent access to the computation libraries still requires a high knowledge of such libraries besides a SWI-Prolog familiarization which is not expected in most part of the potential users. The task-oriented module which is itself very layered provides high-level predicates which are missing some important options at setting-up.

In addition to this, we need require a mechanism to update results and keep track of computation sequences, so that computations are not restricted to one SWI-Prolog session, allowing us to reuse and share them. This section tackles this problem by extending SWI-DSP as an expert system with defined ``DSP scenarios'' and by introducing Concurrent Transaction Logic on top of our extended predicate-logic language.

\subsection{DSP scenarios}

Our DSP workflow is strongly constrained by the predicate instantation patterns which clearly establishes a hierarchy of users depeding on the level of abstraction we are dealing with. Applications provide the user with a nice graphical or command line interface to control the computations. We need therefore to set a ``nice view'' of the system to the potential users which are not (and do not need to be) related to the SWI-DSP or the libraries implementation.

We address this problem by defining ``scenarios'' that control and delimit the user's view of the system without restricting significantly his options. Any computation over an audio signal can be seen as a function input/ouput with a specific setup. The ``scenario'' characterizes such a family of funtions in a generic way so the user just needs to configure the scenario and provide the input to capture the output. Such computations can be wrapped as a signal \textbf{transform} which is declared as a SWI-Prolog functor:

'Transform'(TransformType, 'Engine'(API, PluginName, Version), SampleRate, StepSize, BlockSize, ListOfParameters, Program).

\begin{itemize}

 \item TransformType indicates which sort of task we are carrying out (feature extraction, audio effect...)
 \item 'Engine'(A, Pn, V) identifies the computation engine in charge of the processing of the transform
 \item SampleRate indicates the input sample rate in case it is not constant
 \item StepSize and BlockSize are the framing parameters
 \item ListOfParameters to setup the engine
 \item Program is a predetermined configuration for the engine
\end{itemize}

Then each task module hardcodes the necessary knowledge in form of interpretation rules to read the tranform scenario and set the up the corresponding engine. The predicate \textit{transform(+InputSignal, +TransformContext, -Output)} will provide a generic view of processing to the user with standarized semantics.

Another complex ``scenario'' is classification. We find this approach suitable to control swiweka effectively. Each classification context can be defined by:

'Classification'(Identifier, ListOfCategories, ListOfDataRecords, Classifier, Classification) (have to look at it better)

\subsection{Computation reasoning}

The previous abstraction of computations offers a cleaner view of the system without loosing flexibility but there are more issues to be addressed. The system proposed so far does not provide anything novel in respect of other approaches presented in \S 2 and even could seem even more cumbersome to deal with. The major advantage of using a logic framework as basis of the DSP system is the reasoning capability of such a framework.

This reasoning skills are extended to computation reasoning by using the Concurrent Transaction Logic as superset of the Predicates Logic. Transaction Logic defines the logics that allow us to control and manage the execution of the computation and reason about the effects in these terms:

\begin{itemize}
 \item Integration of updating transactions in queries with formal logics
 \item Integration of declarative and procedural meanings of computational transactions (paths)
\end{itemize}

The extended syntax provides connectives for different program states to create a complex execution path (allowing parallel/serial operations, concurrency control, loops, conditionals, recursion and subroutines). The declarative meaning and the procedural meaning are combined:

computation A executes along path X = A is true on X

Actually we are now in position to define several execution paths for the same computation, control the execution of those paths, reason about the results and the truth of such executions and store the computation sequences and data states.

(need one good example here)

\section{SWI-DSP in real applications}

SWI-DSP defines therefore a logic front-end for computations in DSP ``scenarios''. We have described how Concurrent Transaction Logic converts this framework based on a predicate calculus into a DSP reasoner which control the execution and data of the different scenarios. This system still needs an implementation which allow an external user to control easily the reasoner, access the computations and visualise the results.

\subsection{Languages hosting SWI-Prolog}

.... comment possible languages interfacing prolog...

\subsection{Henry}

Although SWI-Prolog is a well-known and used implementation of Prolog it does not represent the most suitable framework to deal with multimedia data. The multimedia community in general needs a standard machine-readable format with proper semantics to manage data. The Resource Description Framework (RDF) has been proposed as such a format and defines an Standard to represent linked data by means of triples in the form of (Subject, Predicate, Object).

This standard can be implemented with different syntaxes. We prefer to use Notation3 [x] whose syntax allow us to write RDF statements in a human-readable way and also to write ``rules'' as named graphs [g]. This is a suitable framework to represent data that can be shared and published in the so-called ``web of data``.

The Ontology Web Language (OWL) provides the framework to express the knowledge of a particular domain in RDF through concepts and relationships. Thus, descripting semantic models related to audio analysis are not restricted to a local implementation but can be shared and processed by different systems.

The project called Henry implements an N3/Prolog parser and reasoner based on Concurrent Transaction Logic. This agent defines an entailment module capable of reasoning over the SWI-DSP computations, which are identified through Universal Resource Identifiers (URIs), and N3 statements and rules. The semantic web package module for SWI-Prolog interprets SPARQL queries which are passed to Henry which produces RDF descriptions of computations by performing the entailment and triggering SWI-DSP computations.

(any example or figure maybe)

 \section{Conclusion and Future work}

We have presented a powerful reasoner for multimedia analysis implemented as a logic-based framework that aims at converting low-level computations into shareable and distributed knowledge.

\end{document}

### 2008-05-27 trunk/paper/paper.tex

--- 2008-05-24 paper/paper.tex
@@
-\documentclass{article}
+\documentclass[runningheads]{llncs}
+
 \usepackage{amssymb}
@@
 \usepackage{graphicx}
+
 \usepackage{url}
@@
 
+\usepackage{url}
+\urldef{\mails}\path|{david.pastor,yves.raimond, mark.sandler}@elec.qmul.ac.uk|
+%\newcommand{\keywords}[1]{\par\addvspace\baselineskip
+%\noindent\keywordname\enspace\ignorespaces#1}
+
 \begin{document}
 
-\title{A logic-based framework for Digital Signal Processing}
+\mainmatter
 
-\author{...}
+\title{A Logic-Based Framework for Digital Signal Processing}
+
+\titlerunning{A Logic-Based Framework for Digital Signal Processing}
+
+\author{David Pastor\and Yves Raimond\and Mark Sandler}
+
+\institute{Centre for Digital Music, Queen Mary, University of London,\\
+\mails}
 
@@
 Finally, Concurrent Transaction Logic allows us to wrap the procedural meaning of predicate rules into a declarative definition of transactions controlling the execution and updating of the processes.
-\emph{abstract} environment.
-
+\keywords{DSP predicates, computation query, computation reasoner, DSP scenarios, RDF}
 \end{abstract}
 
-\section{Introduction}
+\section{Introduction}\label{sec:intro}
 
-Digital signal processing is a broad paradigm and object of an important research.
-Within this community there are many particular investigators discovering and developing new algorithms for signal processing and analysis.
-Many projects have been directed to develop formal APIs providing a framework for algorithms implementation and publication.
-However these APIs are normally designed as a close environment which requires a specific developer knowledge or as a ``plugin'' library which requires the implementation of a specific host.
+Digital signal processing is a broad field and an object of an important research.
+Withing this research community, particular investigators are discovering and developing new algorithms for signal processing and analysis.
+Many projects aim at developing formal APIs providing a framework for algorithms implementation and publication.
+However, these APIs are normally designed as a closed environment which hardly interacts with other related APIs and applications.
+Although several APIs are wrapped and exported as ``plugin'' libraries which can be embedded in different applications, they still require the implementation of specific and non-interactive hosts.
 
-This situation leads to many isolated applications, environments and sources of data which work independently from each other.
+This situation leads to many isolated applications, environments and sources of data working independently from each other.
 Any effort to combine them to obtain integrated results for analysis, classification and comparison requires tedious studies or ``glue code'' implementations.
 
-In this paper we address this problem by proposing and developing SWI-DSP, a logical framework for \textit{Audio Processing} that extends the SWI-Prolog vocabulary.
-We rely on a logic programming language with audio analysis capabilities getting rid of the application-dependency paradigm and offering a single logic front-end which is generally more understandable than a closed implementation in any Procedural or Object Oriented language.
-In addition to this, we are keen on using logic resources from SWI-Prolog to obtain finally a more capable DSP engine.
+In this paper we address this problematic situation by proposing and developing SWI-DSP, a logical framework for \textit{Audio Processing} (as subset of DSP) that extends the SWI-Prolog vocabulary.
+We rely on a logic programming language with aggregated audio analysis capabilities getting rid of the application-dependency paradigm and offering a single logic front-end which is generally more understandable than a closed implementation in any Procedural or Object Oriented language.
+In addition to this, we are keen on using logic resources to finally achieve a more capable DSP engine.
 
-In \S 2 we summarize and evaluate the alternative APIs for audio analysis.
-Some of these APIs are hosted in SWI-DSP through the foreign interface providing an open system for developers as well.
-In \S 3 we start viewing the system from the top and how it defines an interesting workflow which actually solves the most part of the problems of the DSP paradigm.
-This workflow is determined by semantic types, a binary data standard and procedural declarations of riles.
-\S4 goes into detail in each task-oriented module describing the communication layers with the APIs through control predicates which require an understanding of the API.
+In \S 2 we identify the most common audio analysis tasks and summarize several relevant \textit{open source} APIs aimed at performing such tasks.
+We evaluate the advantages and limitations of these APIs in order to find the requirements for SWI-DSP.
+In \S 3 we start viewing the system from the top and how it defines an interesting workflow that characterizes input/output relationships between task-oriented modules.
+This workflow is determined by semantic types, a binary data standard and procedural declarations of DSP rules.
+\S 4 goes into the implementation of each task-oriented module which are indeed communication layers with dedicated ``plugin'' libraries that are in charge of the actual computation and offer an open development framework.
 
 Thus, SWI-DSP extends SWI-Prolog vocabulary to perform audio analysis and processing by defining semantic types and predicates, triggering external computation performed by dedicated APIs, which are ordered in modules with different levels of abstraction.
-However, this front-end still requires specific SWI-Prolog knowledge to create programs suitable for the concrete needs and it does not offer any efficient tabling mechanism.
-\S 5 explains how Concurrent Transaction Logic is the solution to provide a better logic front-end combining procedural and declarative meaning of transactions which also comprise data updating.
-In \S 6 we present related works as alternatives to manage SWI-DSP featuring another user interface than SWI-Prolog.
+However, this front-end does not present an efficient and manageable framework as it requires specific SWI-Prolog knowledge to create programs suitable for the concrete needs and it does not offer any efficient tabling mechanism.
+We tackle this problem in \S 5 by defining DSP ``scenarios'' which offer a semantic view of the system to the user without losing computational flexibility.
+
+We also describe as a Standard, how Concurrent Transaction Logic is the solution to provide an efficient engine allowing a formal data updating mechanism and computation reasoning capabilities.
+Finally, \S 6 presents related works that use SWI-DSP to create an environment for multimedia data management and visualisation.
 We conclude in \S 7 by assessing the value and scope of this work and introducing future advances.
 
-\section{Audio Analysis tasks and solutions}
+\section{Audio Analysis tasks and solutions}\label{sec:tasks}
 
-Audio analysis  APIs offer formal implementations for development and use of different algorithms and techniques.
+Audio analysis and synthesis APIs offer formal implementations for development and use of different algorithms and techniques.
 These algorithms are focused on different tasks that can be classified in these 3 groups:
@@
 \begin{itemize}
- \item Processing
+ \item Processing and Effects
  \item Feature Extraction
@@
 
-\subsection{Audio Processing}
+\subsection{Audio Processing}\label{subsec:effects}
 
@@
 
-One famous API is the Linux Audio Developer's Simple Plugin API (LADSPA) which has been succesfully used in many systems and applications and is a widely spread implementation for algorithms development.
+One important dedicated API is the Linux Audio Developer's Simple Plugin API (LADSPA) which has been succesfully used in many systems and applications and is a widely spread implementation for algorithms development.
+We host LADSPA in SWI-DSP but there are other interesting libraries like DSSI plugins for effects and LV2 (new generation of LADSPA) which are targets for future extensions.
+(this is too simple i think)
 
-\subsection{Feature Extraction}
+\subsection{Feature Extraction}\label{subsec:feature}
 
 Feature extraction is an important paradigm in signal research as it focuses on extract qualitative characteristics of the signal in a different domain than the input.
-Features are classified in different levels of abstraction from frequency characteristics to high-level audio ones.
+Features are classified in different levels of abstraction from frequency and time domain characteristics to high-level audio ones.
 
-The Vamp Plugins API offers another standard and utilities set to develop feature extraction plugins that can be hosted in any application able to control the ``plugin lifecycle''.
+The Vamp Plugins API offers a standard and utilities to develop feature extraction plugins that can be hosted in any application able to control the ``plugin lifecycle''.
+We host Vamp plugins as basis for feature extraction in SWI-DSP.
+In general plugin libraries are characterized for being easy for development but application-dependent.
+By embedding plugins in SWI-DSP we offer a flexible development framework and make an interactive interface to deal with plugins in several ways.
 
-\subsection{Classification}
+\subsection{Integrative solutions}\label{subsec:inte}
 
-From different features we are able to classify different signals.
+With the same motivation of this work there are some integrative APIs which combines different tasks to perform low high-level operations in a closed implementation.
+CLAM, jMIR and Marsyas present a complex framework collecting tools to perform different DSP tasks.
+The most interesting feature of these frameworks is the possibility to create DSP networks combining specific algorithms and modules thanks to a workflow design.
 
-\subsection{Integrative solutions}
+CLAM also offers a metamodel of DSP processes wich provides semantics to design the networks properly presenting a higher and more manageable level of abstraction.
+However, these APIs still present two important limitations that we pretend to solve:
 
-With the same motivation of this work there are some integrative APIs which combines different tasks to perform high-level operations in a closed framework.
-In the particular domain of Music jMIR and Marsyas present a complex framework for audio analysis and classification.
+\begin{itemize}
+ \item There do not offer a common format to exchange data or an easy way to interact with other applications.
+ \item They offer a procedural system unable to reach higher-level knowledge by itself.
+\end{itemize}
 
-\section{SWI-DSP workflow}
+\subsection{Classification}\label{subsec:classif}
+
+Classification is one the most interesting and complex paradigms in any specific domain exploiting the ``intelligence'' of computational machines.
+The WEKA API collects different machine-learning algorithms and provides a formal interface (in Java) for development and interaction with other systems.
+
+However, this generic API does not provide semantics to create a dataset which is the most intriguing and complex issue for classifying data properly.
+Marsyas and jMÍR integrate WEKA orienting it for music classification purporses by offering a code interface with other modules (Marsyas) or by extending the WEKA file format ARFF (jMIR).
+Both solutions still seem to lack of effective semantics to perform classification of audio data.
+
+\section{SWI-DSP workflow}\label{sec:dspworkflow}
 
@@
 
-\subsection{Semantic datatypes}
+\subsection{Semantic datatypes}\label{subsec:datatypes}
 
@@
 
-\subsection{Large Binary Objects}
+\subsection{Large Binary Objects}\label{subsec:blobs}
 
@@
 
-\subsection{Workflow}
+\subsection{Workflow}\label{subsec:worflow}
 
@@
 
-KeyProfile = ['Feature'(key, 'Timestamp'(0, 23.45), 'Amajor`')|...].
+KeyProfile = ['Feature'(key, 'Timestamp'(0, 23.45), 'Amajor')|...].
 
@@
 
-\section{SWI-DSP modules}
+\section{SWI-DSP modules}\label{sec:modules}
 
 Each task module is composed by lower level modules which offer communication with an external API or library in charge of the computations.
-By hosting APIs we can extend easily SWI-DSP vocabulary as we do not need to assemble each algorithm.
+By hosting ``plugin libraries'' we can extend easily SWI-DSP vocabulary as we do not need to assemble each algorithm.
 
-\subsection{swiaudio}
+\subsection{swiaudio}\label{subsec:swiaudio}
 
@@
 
-\subsection{feature extraction}
+\subsection{feature extraction}\label{subsec:swivamp}
 
@@
 So we can declare a simple predicate to extract the tempo of an audio signal: tempo(+Input, -Tempo) being Input an instantation of the Signal datatype and Tempo a list of instantations of the Feature datatype.
-This ``magic'' predicate is triggering a specific computation tool, so we could select the specific tool by extending the predicate: \textit{tempo(+Input, +Tool, -Tempo}.
+This ``magic'' predicate is triggering a specific computation tool, so we could select the specific tool by extending the predicate: \textit{tempo(+Input, +Tool, -Tempo)}.
 
@@
 Plugins are wrapped in a similar fashion than the audio data so we can use a plugin instance representation which allow us to have concurrent plugins working at the same time.
-The previous tempo/3 predicate is expressed as this set of rules:
+The previous tempo/3 predicate is declared as this set of rules:
 
@@
 
-\subsection{audio processing}
+\subsection{audio processing}\label{subsec:swilasdpa}
 
@@
 
-\subsection{swiweka}
+\subsection{swiweka}\label{subsec:swiweka}
 
-\section{Towards a computation reasoner}
+\section{Towards a computation reasoner}\label{sec:reasoner}
 
@@
 
-\subsection{DSP scenarios}
+\subsection{DSP scenarios}\label{subsec:scenarios}
 
@@
 
-\subsection{Computation reasoning}
+\subsection{Computation reasoning}\label{subsec:ctl}
 
@@
 
-\section{SWI-DSP in real applications}
+\section{SWI-DSP in real applications}\label{sec:realapp}
 
@@
 
-\subsection{Henry}
+\subsection{Henry}\label{subsec:henry}
 
@@
 
- \section{Conclusion and Future work}
+ \section{Conclusion and Future work}\label{sec:conclusion}
 

### 2008-05-27 paper/paper.tex
change to Latex2e

@@
 
-\subsection{audio processing}
+\subsection{audio processing}\label{subsec:swilasdpa}
 
-This module defines any computation which outputs a Signal term. This
+This module defines any computation which outputs a Signal term.
+It comprises any sort of effect or filtering over a signal.
+It reproduces a real time processing by outputting processed frames as they are computed (non-deterministic call to a processing predicate) or stores the processed blocks to return and altered version of the input signal.
 
 The module swiladspa offers communication with LADSPA which is the most generic and spread plugin API for general audio processing.
+The communication layer allows us to query the plugin description, control the set up of the plugin (control and audio ports) and retrieve the processed data.
+Through predefined routines we can offer a more handy access to the plugin or even declared predicates just defined by an input/output relationship.
 
-\subsection{swiweka}
+\subsection{swiweka}\label{subsec:swiweka}
 
-\section{Towards a computation reasoner}
+The WEKA API is interfaced through the package JPL available for SWI-Prolog defining a complex Java-Prolog interface to read/write ARFF files, load ARFF files, instantiate classifiers, execute classification and retrieve the classification scores of the dataset.
+This communication layer offers a really cumbersome and API-dependent module which requires semantics and interpretation rules to become into a useful logic-framework for audio data classification as described in the next section.
+
+\section{Towards a computation reasoner}\label{sec:reasoner}
 
@@
 The communication layer even offering an application-independent access to the computation libraries still requires a high knowledge of such libraries besides a SWI-Prolog familiarization which is not expected in most part of the potential users.
-The task-oriented module which is itself very layered provides high-level predicates which are missing some important options at setting-up.
+The task-oriented module, which is itself very layered, provides high-level predicates which are missing important options at setting-up.
 
-In addition to this, we need require a mechanism to update results and keep track of computation sequences, so that computations are not restricted to one SWI-Prolog session, allowing us to reuse and share them.
-This section tackles this problem by extending SWI-DSP as an expert system with defined ``DSP scenarios'' and by introducing Concurrent Transaction Logic on top of our extended predicate-logic language.
+In addition to this, we require a mechanism to update results and keep track of computation sequences, so that computations are not restricted to one SWI-Prolog session, allowing us to reuse and share them.
+This section tackles this problem by extending SWI-DSP as an expert system able to interpret data within defined ``DSP scenarios'' and by introducing Concurrent Transaction Logic on top of our extended predicate-logic language.
 
-\subsection{DSP scenarios}
+\subsection{DSP scenarios}\label{subsec:scenarios}
 
@@
 
-'Classification'(Identifier, ListOfCategories, ListOfDataRecords, Classifier, Classification) (have to look at it better)
+'Classification'(Identifier, ListOfCategories, ListOfDataRecords, Classifier, ClassificationVector)
 
-\subsection{Computation reasoning}
+\begin{itemize}
+
+ \item Identifier gives a label for the classification set
+ \item 'Category'(Label, Type, ValuesRange) specifies which type of data is going to be classified
+ \item 'DataRecord'(RecordID, ListOfMatchingValues) is a data instance of the dataset 
+ \item 'Classifier'(Name, ...)
+ \item 'Score'(RecordID, ListOfValues) keeps track of each score (vector of scores depeding of the statistical distribution) for every record.
+\end{itemize}
+
+By defining these functors, the engine is able to interpret the data and the user has a clear semantic template to create the classification environment properly.
+We could even define some constraints and predefined functors instanes for a specific domain.
+
+\subsection{Computation reasoning}\label{subsec:ctl}
 

### 2008-05-27 paper/paper.tex

@@
 
+\medskip
+
+\noindent
+
 \begin{verbatim}
@@
 \end{verbatim}
+\noindent
 
@@
 
-.... comment possible languages interfacing prolog...
+Even when there is not current specific work on wrapping SWI-DSP in any other programming language which could make easier to develop a Graphical User Interface (as it probably would restrict the logic capabilities of the framework), we have to mention that SWI-DSP can be wrapped by other languages through the interfaces used for the communication layers (they define both ways of interaction).
 
@@
 
-(any example or figure maybe)
+\subsection{Visualisation}\label{subsec:vis}
+
+Audio data can just be poorly interpreted in a float/binary format.
+Time and frequency domain signals and features in other domains are meaningful for an expert when they are visually represented.
+Although a logic domain does not seem to be the perfect environment to visualise audio data, we can easily save this difficulty.
+
+By controlling SWI-DSP by means of Henry we can easily convert our Prolog types and binary data into RDF (machine-readable) descriptions where each source of data is represented through a link or as a ''literal`` in a basic datatype format.
+RDF descriptions can be easily browsed with the new extensions of classical web browsers for linked data.
+Even more interesting, dedicated agents with an SPARQL point are able to query DSP through the semantic web resources as explained in \ref{subsec:henry}.
+Thus, we are able to visualise our own computations but also a scalable web of data with shared semantics which is enriched with computations from other local machines.
+
+(I dont know if i should mention SV and the new extensions here....)
 
@@
 
-We have presented a powerful reasoner for multimedia analysis implemented as a logic-based framework that aims at converting low-level computations into shareable and distributed knowledge.
+We have presented a powerful reasoner for multimedia analysis implemented as a logic-based framework which allow us to convert low-level computations into shareable and distributed knowledge.
+The system design ''breaks apart`` traditional applications architecture to introduce logic and reasoning machinery on top of the computation engines providing a more intelligent, capable and understandable DSP reasoner.
+This reasoner can be extended and enriched by the open source APIs working on the foundations of the system and controlled by ''Prolog-speaking`` agents.
+SWI-DSP is fairly well positioned for the new web technologies and actually it finds in them the perfect semantic and storing support to overwhelm traditional and closed DSP systems.
 
+Future work and development is oriented to:
+
+\begin{itemize}
+ \item Testing and comparison of performance with other rival systems
+ \item Extending the SWI-DSP vocabulary and capabilities
+\end{itemize}
+
+SWI-DSP vocabulary and cabalities can be extended in several ways.
+First, by interfacing more open source APIs to cover the implementation of any DSP task (not necessarily restricted to audio domain).
+Secondly, by supporting symbolic data (MIDI) input which is a really important source of audio knowledge and providing analysis modules for such an input.
+Finally, other logic paradigms can contribute to enhance SWI-DSP (e.g.
+there are studies that use Inductive Logic Programming systems to extract harmonic rules or rhythm patterns given certain input descriptors [ref]).
+By providing semantics (OWL ontologies cleary point to this goal) to interconnect properly each logical module we can build a hierarchy of modules which help us to extract and share different levels of knowledge from digital signals.
+Such a system could be used either to extract low level cepstral coefficients from an input signal (mathematics level) or to obtain patterns of music (musicology level).
 \end{document}

### 2008-05-30 paper/paper.tex

@@
 \usepackage{url}
-\urldef{\mails}\path|{david.pastor,yves.raimond, mark.sandler}@elec.qmul.ac.uk|
-%\newcommand{\keywords}[1]{\par\addvspace\baselineskip
-%\noindent\keywordname\enspace\ignorespaces#1}
+\urldef{\mails}\path|{david.pastor, yves.raimond, mark.sandler}@elec.qmul.ac.uk|
 
@@
 In this paper we describe a logic-based framework for Digital Signal Processing.
-This framework is implemented on SWI-Prolog and is formed of interfaces with different DSP libraries and APIs extending this logic language ``to speak'' DSP.
-This approach presents the advantage, in respect of other integrated APIs, of being application implementation independent, so the extended logic language can be loaded an embedded in any application or language interfacing SWI-Prolog.
-The new DSP capabilities are public as predicates grouped in modules which offer different levels of abstraction.
-Low-level modules offer predicates to communicate directly with the API.
-Task-oriented modules define some default routines which specific goals and procedures.
-Finally we can relate input/output through ``black boxes'' representing a semantic relationship between them.
-Modules can communicate by using commong semantic data types declared as functors which determine the possible workflows.
-Such a framework allow us to build complex DSP queries as procedures as well as simplify the extraction of audio processing knowledge with common semantics.
-Finally, Concurrent Transaction Logic allows us to wrap the procedural meaning of predicate rules into a declarative definition of transactions controlling the execution and updating of the processes.
-\keywords{DSP predicates, computation query, computation reasoner, DSP scenarios, RDF}
+This framework is implemented in SWI-Prolog and develops interfaces to signal processing libraries providing that with modules ``to speak'' DSP.
+Modules communicate by using common functors, which we use to handle the orchestration of our different analysis components.
+The system provides the user with a unified and flexible front-end to define signal processing workflows.
+These workflows involve a number of heterogeneous software libraries.
+Concurrent Transaction Logic allows us to combine procedural and declarative meaning of queries to execute computations and update data obtaining a powerful query-driven DSP reasoner.
+
+\keywords{DSP predicates, semantic front-end, Concurrent Transaction Logic, computation reasoner, DSP-quering agent}
 \end{abstract}
@@
 
-Digital signal processing is a broad field and an object of an important research.
+Digital signal processing (DSP) is a broad field and an object of an important research.
 Withing this research community, particular investigators are discovering and developing new algorithms for signal processing and analysis.
@@
 In this paper we address this problematic situation by proposing and developing SWI-DSP, a logical framework for \textit{Audio Processing} (as subset of DSP) that extends the SWI-Prolog vocabulary.
-We rely on a logic programming language with aggregated audio analysis capabilities getting rid of the application-dependency paradigm and offering a single logic front-end which is generally more understandable than a closed implementation in any Procedural or Object Oriented language.
-In addition to this, we are keen on using logic resources to finally achieve a more capable DSP engine.
+We rely on a logic programming language with aggregated audio analysis capabilities getting rid of the application-dependency paradigm and offering a single logic front-end.
+Logic resources are also used to achieve a more capable DSP engine.
+The main and primary motivation for a logic-based implementation of such a system is to turn our conception of traditional procedural applications for computation of data towards semantic frameworks to manage knowledge envolving specific computations.
 
@@
 We evaluate the advantages and limitations of these APIs in order to find the requirements for SWI-DSP.
-In \S 3 we start viewing the system from the top and how it defines an interesting workflow that characterizes input/output relationships between task-oriented modules.
-This workflow is determined by semantic types, a binary data standard and procedural declarations of DSP rules.
-\S 4 goes into the implementation of each task-oriented module which are indeed communication layers with dedicated ``plugin'' libraries that are in charge of the actual computation and offer an open development framework.
+In \S 3 we start viewing the system from the top and how it defines an interesting workflows-based architecture that characterizes input/output relationships between task-oriented modules.
+These workflows are possible by means of common functorsa and a binary data Standard.
+\S 4 goes into the implementation of each task-oriented module which are indeed communication layers with dedicated audio processing libraries that are in charge of the actual computation and offer an open development framework.
 
-Thus, SWI-DSP extends SWI-Prolog vocabulary to perform audio analysis and processing by defining semantic types and predicates, triggering external computation performed by dedicated APIs, which are ordered in modules with different levels of abstraction.
+Thus, SWI-DSP extends SWI-Prolog vocabulary to perform audio analysis and processing by defining functors (as common semantic types) and predicates, triggering external computation performed by dedicated APIs, which are ordered in modules with different levels of abstraction.
 However, this front-end does not present an efficient and manageable framework as it requires specific SWI-Prolog knowledge to create programs suitable for the concrete needs and it does not offer any efficient tabling mechanism.
-We tackle this problem in \S 5 by defining DSP ``scenarios'' which offer a semantic view of the system to the user without losing computational flexibility.
+We tackle this problem in \S 5 by defining DSP ``scenarios'' which offer a unified view of the system to the user preserving computational flexibility.
 
-We also describe as a Standard, how Concurrent Transaction Logic is the solution to provide an efficient engine allowing a formal data updating mechanism and computation reasoning capabilities.
+/fixthis
+We also describe, as a Standard, how Concurrent Transaction Logic is the solution to provide an efficient engine allowing a formal data updating mechanism and computation reasoning capabilities.
 Finally, \S 6 presents related works that use SWI-DSP to create an environment for multimedia data management and visualisation.
@@
 
-\subsection{Audio Processing}\label{subsec:effects}
+\subsection{Signal Procesing}\label{subsec:effects}
 
@@
 
-Feature extraction is an important paradigm in signal research as it focuses on extract qualitative characteristics of the signal in a different domain than the input.
-Features are classified in different levels of abstraction from frequency and time domain characteristics to high-level audio ones.
+Feature extraction is an important paradigm in signal research as it focuses on extract qualitative characteristics of the signal.
+Features are classified in different levels of abstraction which may go from frequency and time domain characteristics to high-level audio ones.
 
@@
 In general plugin libraries are characterized for being easy for development but application-dependent.
-By embedding plugins in SWI-DSP we offer a flexible development framework and make an interactive interface to deal with plugins in several ways.
+By embedding plugins in SWI-DSP we offer a flexible development framework and provide an interactive interface to deal with plugins in several ways.
 
@@
      vmpl_load_plugin_for(PluginKey, Signal, Plugin),
-        set_blockSize(Plugin, BlockSize),
-        set_stepSize(Plugin, StepSize),
+     set_blockSize(Plugin, Block),
+     set_stepSize(Plugin, Step),
      get_channels(Signal, Channels),
      vmpl_initialize_plugin(Plugin, Channels, StepSize, BlockSize),
-        vamp_compute_feature(Signal, StepSize, BlockSize, Outputs, Plugin, WholeFeature).
-vamp_compute_feature(Signal, StepSize, BlockSize, Outputs, Plugin, WholeFeature):-
+     vamp_compute_feature(Signal, Step, Block, Outputs, Plugin, WholeFeature).
+vamp_compute_feature(Signal, Step, Block, Outputs, Plugin, Features):-
      get_samples_per_channel(Signal, L),
-        findall(Features, vamp_process_signal(Signal, L, StepSize, BlockSize, Output, Plugin, Features), FeatureSet),
+     findall(F, vamp_process_signal(Signal, L, Step, Block, Output, Plugin, F), FSet),
      get_sample_rate(Signal, SampleRate),
-        vmpl_remaining_features(Plugin, L, SampleRate, Output, Remaining),
-        append(FeatureSet, Remaining, RawFeatures),
-        delete(RawFeatures, [], WholeFeature).
-vamp_process_signal(Signal, L, StepSize, BlockSize, Output, Plugin, Features):-
+     vmpl_remaining_features(Plugin, L, SampleRate, Outputs, Remaining),
+     append(FSet, Remaining, RawFeatures),
+     delete(RawFeatures, [], Features).
+vamp_process_signal(Signal, L, StepSize, Block, Out, Plugin, Features):-
      set_limit_framing(L, StepSize, Limit),
      set_framing(StepSize, L, Limit, Start),
-        vmpl_process_block_framing(Plugin, Signal, Start, BlockSize, Output, FeatureSet).
+     vmpl_process_block_framing(Plugin, Signal, Start, Block, Out, Features).
 
@@
 Such a system could be used either to extract low level cepstral coefficients from an input signal (mathematics level) or to obtain patterns of music (musicology level).
+
+\begin{thebibliography}{4}
+
+\bibitem{distributed} Raimond, Y., Sutton, C., Sandler, M.: A distributed data space for music-related information.
+In: Workshop on the Many Faces of Multimedia Semantic.
+ACM (Multimedia), Augsburg (2007).
+
+\bibitem{ctl} Bonner, A.J., Kifer, M.: Concurrency and Communication in Transaction Logic. In: Joint International Conference and Symposium on Logic Programming, pp. 142--156. Bonn (1996)
+
+\bibitem{tlprog} Bonner, A.J., Kifer, M.: Transaction Logic Programming. In: 10th International Conference on Logic Programming. Budapest (1993)
+
+\bibitem{clam} CLAM, \url{http://www.clam.iua.upf.edu/}
+
+\bibitem{marsyas} Music Analysis, Retrieval and Synthesis for Audio Signals. \url{http://marsyas.sness.net/}
+
+\bibitem{jmir} McKay, C.: jMIR. \url{http://jmir.sourceforge.net/}
+
+\bibitem{mo} Raimond, Y., Abdallah, S., Sandler, M., Frederick, G.: The Music Ontology. In: 8th International Conference of Music Information Retrieval. Vienna (2007)
+
+\bibitem{kmrdf} Source code project of KM-RDF. \url{http://code.google.com/p/km-rdf/w/list}
+
+\bibitem{ladspa} Linux Audio Developer Simple Plugin API. \url{http://www.ladspa.org/}
+
+\bibitem{vamp} Cannam, C.: Vamp Plugins. \url{http://www.vamp-plugins.org/}
+
+\end{thebibliography}
+
 \end{document}

### 2008-06-02 paper/paper.tex
update

@@
 In this paper we describe a logic-based framework for Digital Signal Processing.
-This framework is implemented in SWI-Prolog and develops interfaces to signal processing libraries providing that with modules ``to speak'' DSP.
-Modules communicate by using common functors, which we use to handle the orchestration of our different analysis components.
+This framework is implemented as a set of SWI-Prolog modules developing interfaces to heterogeneous signal processing libraries.
+A set of functors capturing common datatypes and a standard format for binary data are used to mediate our different analysis components.
 The system provides the user with a unified and flexible front-end to define signal processing workflows.
-These workflows involve a number of heterogeneous software libraries.
-Concurrent Transaction Logic allows us to combine procedural and declarative meaning of queries to execute computations and update data obtaining a powerful query-driven DSP reasoner.
+The framework develops a signal processing-enabled reasoner capable of performing the primary tasks through the interfaces: decoding (specific libraries for each format), feature extraction (Vamp plugins), audio processing and effects (LV2/LADSPA plugins) and classification (WEKA machine learning API).
 
-\keywords{DSP predicates, semantic front-end, Concurrent Transaction Logic, computation reasoner, DSP-quering agent}
+\keywords{Digital Signal Processing, SWI-prolog modules, signal processing workflows, processing-enabled reasoner, interfaced open source engines}
 \end{abstract}
@@
 
-Digital signal processing (DSP) is a broad field and an object of an important research.
-Withing this research community, particular investigators are discovering and developing new algorithms for signal processing and analysis.
-Many projects aim at developing formal APIs providing a framework for algorithms implementation and publication.
+Digital signal processing (DSP) is a broad field and an important object of engineering research.
+Within this research community, particular investigators are discovering and developing new algorithms for signal processing and analysis.
+Many projects aim at developing formal APIs providing a framework for algorithms implementation.
 However, these APIs are normally designed as a closed environment which hardly interacts with other related APIs and applications.
@@
 
-In this paper we address this problematic situation by proposing and developing SWI-DSP, a logical framework for \textit{Audio Processing} (as subset of DSP) that extends the SWI-Prolog vocabulary.
+In this paper we address this problematic situation by proposing and developing SWI-DSP, a logic-based framework for \textit{Audio Processing} (as subset of DSP) that extends the SWI-Prolog vocabulary.
 We rely on a logic programming language with aggregated audio analysis capabilities getting rid of the application-dependency paradigm and offering a single logic front-end.
 Logic resources are also used to achieve a more capable DSP engine.
-The main and primary motivation for a logic-based implementation of such a system is to turn our conception of traditional procedural applications for computation of data towards semantic frameworks to manage knowledge envolving specific computations.
+The main and primary motivation for a logic-based implementation of such a system is to turn our conception of traditional procedural applications for computation of data towards semantic frameworks to manage knowledge involving specific computations.
 
@@
 In \S 3 we start viewing the system from the top and how it defines an interesting workflows-based architecture that characterizes input/output relationships between task-oriented modules.
-These workflows are possible by means of common functorsa and a binary data Standard.
-\S 4 goes into the implementation of each task-oriented module which are indeed communication layers with dedicated audio processing libraries that are in charge of the actual computation and offer an open development framework.
+These workflows are possible by means of common functors and a binary data standard format.
+\S 4 goes into the implementation of each task-oriented module that develops communication layers with dedicated audio processing libraries that are in charge of the actual computation and offer an open development framework.
 
-Thus, SWI-DSP extends SWI-Prolog vocabulary to perform audio analysis and processing by defining functors (as common semantic types) and predicates, triggering external computation performed by dedicated APIs, which are ordered in modules with different levels of abstraction.
-However, this front-end does not present an efficient and manageable framework as it requires specific SWI-Prolog knowledge to create programs suitable for the concrete needs and it does not offer any efficient tabling mechanism.
-We tackle this problem in \S 5 by defining DSP ``scenarios'' which offer a unified view of the system to the user preserving computational flexibility.
+Thus, SWI-DSP extends SWI-Prolog vocabulary to perform audio analysis and processing by defining functors (capturing common datatypes) and predicates, triggering external computation performed by dedicated APIs, which are ordered in modules with different levels of abstraction.
+However, this architecture does not present an efficient and manageable framework as it requires specific module knowledge to create programs suitable for the concrete needs.
+We tackle this problem in \S 5 by defining an unified and flexible front-end, composed by DSP ``scenarios'', which allows the user to define workflows with understandable semantics.
 
-/fixthis
-We also describe, as a Standard, how Concurrent Transaction Logic is the solution to provide an efficient engine allowing a formal data updating mechanism and computation reasoning capabilities.
-Finally, \S 6 presents related works that use SWI-DSP to create an environment for multimedia data management and visualisation.
+Data management is an important paradigm involved in many domains and also in the multimedia research domain.
+Computations can not often be retrieved from session to session or are stored in local filesystems unlikely directed to extend any shareable source of knowledge because they are not expressed in a standard format with common semantics.
+This problem also encourages the approach presented in this work, based on a predicate calculus implementation (SWI-Prolog), in two different ways:
+
+\begin{itemize}
+ \item Concurrent Transcation Logic can be used to extend the predicate calculus to provide an unified logic framework for computations and data updating (given a database implementation).
+ \item SWI-Prolog provides an important set of interfaces to other languages (to build user applications embedding the system) and packages that allow us to interact with semantic web technologies which seem to be the best framework to provide a common source of knowledge.
+\end{itemize}
+
+Although SWI-DSP does not develop a data management system itself, we discuss this problem in \S 6 and present related works which contribute to achieve such a multimedia knowledge reasoner/manager.
 We conclude in \S 7 by assessing the value and scope of this work and introducing future advances.

### 2008-06-02 front-end/transform.pl
interpretation rules and a front-end module to deal with the interfaces in an unified way

@@ -1,0 +1,66 @@
+/**
+	General view of signal transforms
+	David Pastor Escuredo, c4dm
+	May 2008
+
+	Contains hardcoded rules for interpretation of a transform predicate
+*/
+:-module(transform, [api/1
+		,	plugin_engine/1
+			]).
+
+
+:-use_module(../feature-extractor/fe).
+:-use_module(../swivamp/vamp).
+:-use_module(../swiladspa/ladspa).
+//load effects
+:-style_check(-discontiguous).
+
+/*'Transform'(TransformType, 'Engine'(API, PluginName, Version), SampleRate, StepSize, BlockSize, ListOfParameters, Program).*/
+
+
+/* Hardcoded information */
+transform_type('feature_extraction').
+transform_type('effect').
+api('vamp').
+api('ladspa').
+
+/**Querying the possible engines**/
+plugin_engine('Engine'(API, PluginName, _Version)):-
+	(vamp_plugin_system(PluginName), API ='vamp', vamp_plugin_version(PluginName, Version));
+	(lpdl_plugin_system(PluginName), API='ladspa').
+
+/**read_parameters**/
+set_parameters(Engine, Plugin, ListOfParameters):-
+	Engine = 'Engine'(A, P, V),
+	(A = 'vamp', vamp_set_parameters(Plugin, ListOfParameters));
+	(A = 'ladspa', ladspa_ser_parameters(Plugin, ListOfParameters)).
+
+/** api-specific routines for parameters setup**/
+vamp_set_parameters(PluginName, Plugin, _).
+vamp_set_parameters(PluginName, Plugin, [H|T]):-
+	H = 'Parameter'(Name, Value),
+	vmpl_pluginParameters(Plugin, List),
+	member(Name, List),
+	vmpl_set_parameter(Plugin, Name, Value),
+	vamp_set_parameters(Plugin, T).
+
+ladspa_set_parameters(PluginName, Plugin, _).
+ladspa_set_parameters(PluginName, Plugin, [H|T]).
+	H = 'Parameter'(Name, Value),
+	ldpl_port_description(PluginName, 'LadspaPort'(T, I, N)),
+	Name = N,
+	((T = 'InputControl', ldpl_set_parameter(Plugin, I, Value));
+	(T = 'OutputControl', ldpl_set_output_control(Plugin, I, Value))).
+	
+/** program selection for vamp */
+set_program(API, Plugin, Program):-
[…]

### 2008-06-04 trunk/vamp-n3/vamp_n3.pl
All fixes and comments

@@ -1,11 +1,0 @@
-:- module(vamp-mapping,[]).
-
-:- consult(library('semweb/rdf_db')).
-:- use_module('../p2r/match').
-
-:- consult('vamp-ns').
-
-
-
-
-

### 2008-06-04 trunk/paper/paper.tex

@@
 In this paper we describe a logic-based framework for Digital Signal Processing.
-This framework is implemented on SWI-Prolog and is formed of interfaces with different DSP libraries and APIs extending this logic language ``to speak'' DSP.
-This approach presents the advantage, in respect of other integrated APIs, of being application implementation independent, so the extended logic language can be loaded an embedded in any application or language interfacing SWI-Prolog.
-The new DSP capabilities are public as predicates grouped in modules which offer different levels of abstraction.
-Low-level modules offer predicates to communicate directly with the API.
-Task-oriented modules define some default routines which specific goals and procedures.
-Finally we can relate input/output through ``black boxes'' representing a semantic relationship between them.
-Modules can communicate by using commong semantic data types declared as functors which determine the possible workflows.
-Such a framework allow us to build complex DSP queries as procedures as well as simplify the extraction of audio processing knowledge with common semantics.
-Finally, Concurrent Transaction Logic allows us to wrap the procedural meaning of predicate rules into a declarative definition of transactions controlling the execution and updating of the processes.
-\keywords{DSP predicates, computation query, computation reasoner, DSP scenarios, RDF}
+This framework is implemented as a set of SWI-Prolog modules developing interfaces to heterogeneous signal processing libraries.
+A set of functors capturing common datatypes and a standard format for binary data are used to mediate our different analysis components.
+The system provides the user with a unified and flexible front-end to define signal processing workflows.
+The framework develops a signal processing-enabled reasoner capable of performing the primary tasks through the interfaces: decoding (specific libraries for each format), feature extraction (Vamp plugins), audio processing and effects (LV2/LADSPA plugins) and classification (WEKA machine learning API).
+
+\keywords{Digital Signal Processing, SWI-prolog modules, signal processing workflows, processing-enabled reasoner, interfaced open source engines}
 \end{abstract}
@@
 
-Digital signal processing is a broad field and an object of an important research.
-Withing this research community, particular investigators are discovering and developing new algorithms for signal processing and analysis.
-Many projects aim at developing formal APIs providing a framework for algorithms implementation and publication.
+Digital signal processing (DSP) is a broad field and an important object of engineering research.
+Within this research community, particular investigators are discovering and developing new algorithms for signal processing and analysis.
+Many projects aim at developing formal APIs providing a framework for algorithms implementation.
 However, these APIs are normally designed as a closed environment which hardly interacts with other related APIs and applications.
@@
 
-In this paper we address this problematic situation by proposing and developing SWI-DSP, a logical framework for \textit{Audio Processing} (as subset of DSP) that extends the SWI-Prolog vocabulary.
-We rely on a logic programming language with aggregated audio analysis capabilities getting rid of the application-dependency paradigm and offering a single logic front-end which is generally more understandable than a closed implementation in any Procedural or Object Oriented language.
-In addition to this, we are keen on using logic resources to finally achieve a more capable DSP engine.
+In this paper we address this problematic situation by proposing and developing SWI-DSP, a logic-based framework for \textit{Audio Processing} (as subset of DSP) that extends the SWI-Prolog vocabulary.
+We rely on a logic programming language with aggregated audio analysis capabilities getting rid of the application-dependency paradigm and offering a single logic front-end.
+Logic resources are also used to achieve a more capable DSP engine.
+The main and primary motivation for a logic-based implementation of such a system is to turn our conception of traditional procedural applications for computation of data towards semantic frameworks to manage knowledge involving specific computations.
 
@@
 We evaluate the advantages and limitations of these APIs in order to find the requirements for SWI-DSP.
-In \S 3 we start viewing the system from the top and how it defines an interesting workflow that characterizes input/output relationships between task-oriented modules.
-This workflow is determined by semantic types, a binary data standard and procedural declarations of DSP rules.
-\S 4 goes into the implementation of each task-oriented module which are indeed communication layers with dedicated ``plugin'' libraries that are in charge of the actual computation and offer an open development framework.
+In \S 3 we start viewing the system from the top and how it defines an interesting workflows-based architecture that characterizes input/output relationships between task-oriented modules.
+These workflows are possible by means of common functors and a binary data standard format.
+\S 4 goes into the implementation of each task-oriented module that develops communication layers with dedicated audio processing libraries that are in charge of the actual computation and offer an open development framework.
 
-Thus, SWI-DSP extends SWI-Prolog vocabulary to perform audio analysis and processing by defining semantic types and predicates, triggering external computation performed by dedicated APIs, which are ordered in modules with different levels of abstraction.
-However, this front-end does not present an efficient and manageable framework as it requires specific SWI-Prolog knowledge to create programs suitable for the concrete needs and it does not offer any efficient tabling mechanism.
-We tackle this problem in \S 5 by defining DSP ``scenarios'' which offer a semantic view of the system to the user without losing computational flexibility.
+Thus, SWI-DSP extends SWI-Prolog vocabulary to perform audio analysis and processing by defining functors (capturing common datatypes) and predicates, triggering external computation performed by dedicated APIs, which are ordered in modules with different levels of abstraction.
+However, this architecture does not present an efficient and manageable framework as it requires specific module knowledge to create programs suitable for the concrete needs.
+We tackle this problem in \S 5 by defining an unified and flexible front-end, composed by DSP ``scenarios'', which allows the user to define workflows with understandable semantics.
 
-We also describe as a Standard, how Concurrent Transaction Logic is the solution to provide an efficient engine allowing a formal data updating mechanism and computation reasoning capabilities.
-Finally, \S 6 presents related works that use SWI-DSP to create an environment for multimedia data management and visualisation.
+Data management is an important paradigm involved in many domains and also in the multimedia research domain.
+Computations can not often be retrieved from session to session or are stored in local filesystems unlikely directed to extend any shareable source of knowledge because they are not expressed in a standard format with common semantics.
+This problem also encourages the approach presented in this work, based on a predicate calculus implementation (SWI-Prolog), in two different ways:
+
+\begin{itemize}
+ \item Concurrent Transcation Logic can be used to extend the predicate calculus to provide an unified logic framework for computations and data updating (given a database implementation).
+ \item SWI-Prolog provides an important set of interfaces to other languages (to build user applications embedding the system) and packages that allow us to interact with semantic web technologies which seem to be the best framework to provide a common source of knowledge.
+\end{itemize}
+
+Although SWI-DSP does not develop a data management system itself, we discuss this problem in \S 6 and present related works which contribute to achieve such a multimedia knowledge reasoner/manager.
 We conclude in \S 7 by assessing the value and scope of this work and introducing future advances.
 
-\section{Audio Analysis tasks and solutions}\label{sec:tasks}
+\section{Audio processing tasks and solutions}\label{sec:tasks}
 
-Audio analysis and synthesis APIs offer formal implementations for development and use of different algorithms and techniques.
-These algorithms are focused on different tasks that can be classified in these 3 groups:
+A particular application of DSP is the study of one-dimensional, time domain signals that in real world often exist as audio signals.
+Audio signals evolve in time and have acoustic properties produced by the signal structure and shape.
+Music and speech are the main domains involving audio signals.
+
+The properties of an analog and continuous signal (acoustic signals) can be studied by analysing their digital (sampled and quantified) versions.
+A one-dimension digital signal is therefore a processable sequence of values in a discretized time domain that represents the audio signal.
+Digital signals allow machine processing useful to modify the orignal signal or extract characterisitics from it.
+We can classify the digital signal processing paradigm in these main tasks:
 
 \begin{itemize}
- \item Processing and Effects
+ \item Effects
  \item Feature Extraction
@@
 
-\subsection{Audio Processing}\label{subsec:effects}
+\subsection{Effects}\label{subsec:effects}
 
-This tasks group can be characterized by having audio signals as input and output, so an input signal is processed in such a way that we retrieve an altered, processed version of the original signal.
-Some examples of audio processing tasks are filtering, effects processing and signal generation.
+A signal may be presented in a raw or non desirable way for its analysis so a basic stage of many systems is the pre-processing.
+The signal is processed to obtain a satisfying version of it by different techniques (filtering, normalization, amplification...).
 
-One important dedicated API is the Linux Audio Developer's Simple Plugin API (LADSPA) which has been succesfully used in many systems and applications and is a widely spread implementation for algorithms development.
+It is very common in applied systems to provide tools that modify acoustic properties of the signal by introducing specific effects.
+Reverberation, delay or distortion are examples of effects that can be digitally obtained reproducing natural or instrumental alterations of the signal.
+
+In practice, we can characterize this group of tasks for its input/output relationships.
+The operation outputs a signal (which may be time, frequency or multidimensional domain) when is fed with an input signal (accepting several domains as well).
+Oscillators and synthesis systems outputs signals without input.
+
+One important dedicated API is the Linux Audio Developer's Simple Plugin API (LADSPA) \cite{ladspa}  which has been succesfully used in many systems and applications and is a widely spread implementation for algorithms development.
 We host LADSPA in SWI-DSP but there are other interesting libraries like DSSI plugins for effects and LV2 (new generation of LADSPA) which are targets for future extensions.
-(this is too simple i think)
 
@@
 
-Feature extraction is an important paradigm in signal research as it focuses on extract qualitative characteristics of the signal in a different domain than the input.
-Features are classified in different levels of abstraction from frequency and time domain characteristics to high-level audio ones.
+Acoustic properties which are the substract of the psychological perspecitve of audio signals are conveyed in the signal and therefore in their digital representations.
+The so-called feature extraction paradigm aims at extracting different levels of audio features to characterize, classify and compare the audio signals (and any signal in general).
+The domain of the feature represents a qualitative side of the signal (tempo, chromatism, timbre...), so features do not convey all the signal information.
+This is interesting as distinct signals may have some very close features and differ much in others.
 
-The Vamp Plugins API offers a standard and utilities to develop feature extraction plugins that can be hosted in any application able to control the ``plugin lifecycle''.
+The Vamp Plugins API \cite{vamp} offers a framework to develop feature extraction plugins that can be hosted in any application able to control the ``plugin lifecycle''.
 We host Vamp plugins as basis for feature extraction in SWI-DSP.
-In general plugin libraries are characterized for being easy for development but application-dependent.
-By embedding plugins in SWI-DSP we offer a flexible development framework and make an interactive interface to deal with plugins in several ways.
+In general plugin libraries are characterized for being suitable for development but application-dependent.
 
@@
 
-With the same motivation of this work there are some integrative APIs which combines different tasks to perform low high-level operations in a closed implementation.
-CLAM, jMIR and Marsyas present a complex framework collecting tools to perform different DSP tasks.
-The most interesting feature of these frameworks is the possibility to create DSP networks combining specific algorithms and modules thanks to a workflow design.
+With the same motivation of this work there are some integrative APIs which combines different tasks to perform low and hig-level operations in a closed implementation.
+CLAM \cite{clam}, jMIR \cite{jmir} and Marsyas \cite{marsyas} present a complex framework collecting tools to perform different DSP tasks.
+The most interesting feature of these frameworks is the possibility to create DSP networks combining specific algorithms and modules thanks to a workflow-based architecture (Marsyas and CLAM) or with a descriptive task format (jMIR).
 
 CLAM also offers a metamodel of DSP processes wich provides semantics to design the networks properly presenting a higher and more manageable level of abstraction.
-However, these APIs still present two important limitations that we pretend to solve:
+However, these APIs still present two important limitations that our work is aimed at solving:
 
 \begin{itemize}
- \item There do not offer a common format to exchange data or an easy way to interact with other applications.
- \item They offer a procedural system unable to reach higher-level knowledge by itself.
+ \item They do not offer a common format to exchange data or an easy way to interact with other applications.
+ \item They are based on a procedural system unable to reach higher-level knowledge by itself.
 \end{itemize}
@@
 Classification is one the most interesting and complex paradigms in any specific domain exploiting the ``intelligence'' of computational machines.
-The WEKA API collects different machine-learning algorithms and provides a formal interface (in Java) for development and interaction with other systems.
+WEKA (Waikato Environment for Knowledge Analysis) \cite{weka} collects different machine-learning algorithms and provides several user interfaces to manage the analysis.
+It aslo provdes a formal interface (in Java) for development and interaction with other systems.
 
@@
 Marsyas and jMÍR integrate WEKA orienting it for music classification purporses by offering a code interface with other modules (Marsyas) or by extending the WEKA file format ARFF (jMIR).
-Both solutions still seem to lack of effective semantics to perform classification of audio data.
+Both solutions still seem to lack of effective semantics and a interactive front-end to perform classification of audio data.
 
-\section{SWI-DSP workflow}\label{sec:dspworkflow}
+\section{SWI-DSP architecture}\label{sec:architecture}
 
@@
      vmpl_load_plugin_for(PluginKey, Signal, Plugin),
-        set_blockSize(Plugin, BlockSize),
-        set_stepSize(Plugin, StepSize),
+     set_blockSize(Plugin, Block),
+     set_stepSize(Plugin, Step),
      get_channels(Signal, Channels),
      vmpl_initialize_plugin(Plugin, Channels, StepSize, BlockSize),
-        vamp_compute_feature(Signal, StepSize, BlockSize, Outputs, Plugin, WholeFeature).
-vamp_compute_feature(Signal, StepSize, BlockSize, Outputs, Plugin, WholeFeature):-
+     vamp_compute_feature(Signal, Step, Block, Outputs, Plugin, WholeFeature).
+vamp_compute_feature(Signal, Step, Block, Outputs, Plugin, Features):-
      get_samples_per_channel(Signal, L),
-        findall(Features, vamp_process_signal(Signal, L, StepSize, BlockSize, Output, Plugin, Features), FeatureSet),
+     findall(F, vamp_process_signal(Signal, L, Step, Block, Output, Plugin, F), FSet),
      get_sample_rate(Signal, SampleRate),
-        vmpl_remaining_features(Plugin, L, SampleRate, Output, Remaining),
-        append(FeatureSet, Remaining, RawFeatures),
-        delete(RawFeatures, [], WholeFeature).
-vamp_process_signal(Signal, L, StepSize, BlockSize, Output, Plugin, Features):-
+     vmpl_remaining_features(Plugin, L, SampleRate, Outputs, Remaining),
+     append(FSet, Remaining, RawFeatures),
+     delete(RawFeatures, [], Features).
+vamp_process_signal(Signal, L, StepSize, Block, Out, Plugin, Features):-
      set_limit_framing(L, StepSize, Limit),
      set_framing(StepSize, L, Limit, Start),
-        vmpl_process_block_framing(Plugin, Signal, Start, BlockSize, Output, FeatureSet).
+     vmpl_process_block_framing(Plugin, Signal, Start, Block, Out, Features).
 
 \end{verbatim}
+\noindent
 
@@
 
-This module defines any computation which outputs a Signal term. This
+This module defines any computation which outputs a Signal term.
+It comprises any sort of effect or filtering over a signal.
+It reproduces a real time processing by outputting processed frames as they are computed (non-deterministic call to a processing predicate) or stores the processed blocks to return and altered version of the input signal.
 
 The module swiladspa offers communication with LADSPA which is the most generic and spread plugin API for general audio processing.
+The communication layer allows us to query the plugin description, control the set up of the plugin (control and audio ports) and retrieve the processed data.
+Through predefined routines we can offer a more handy access to the plugin or even declared predicates just defined by an input/output relationship.
 
 \subsection{swiweka}\label{subsec:swiweka}
+
+The WEKA API is interfaced through the package JPL available for SWI-Prolog defining a complex Java-Prolog interface to read/write ARFF files, load ARFF files, instantiate classifiers, execute classification and retrieve the classification scores of the dataset.
+This communication layer offers a really cumbersome and API-dependent module which requires semantics and interpretation rules to become into a useful logic-framework for audio data classification as described in the next section.
 
@@
 The communication layer even offering an application-independent access to the computation libraries still requires a high knowledge of such libraries besides a SWI-Prolog familiarization which is not expected in most part of the potential users.
-The task-oriented module which is itself very layered provides high-level predicates which are missing some important options at setting-up.
+The task-oriented module, which is itself very layered, provides high-level predicates which are missing important options at setting-up.
 
-In addition to this, we need require a mechanism to update results and keep track of computation sequences, so that computations are not restricted to one SWI-Prolog session, allowing us to reuse and share them.
-This section tackles this problem by extending SWI-DSP as an expert system with defined ``DSP scenarios'' and by introducing Concurrent Transaction Logic on top of our extended predicate-logic language.
+In addition to this, we require a mechanism to update results and keep track of computation sequences, so that computations are not restricted to one SWI-Prolog session, allowing us to reuse and share them.
+This section tackles this problem by extending SWI-DSP as an expert system able to interpret data within defined ``DSP scenarios'' and by introducing Concurrent Transaction Logic on top of our extended predicate-logic language.
 
@@
 
-'Classification'(Identifier, ListOfCategories, ListOfDataRecords, Classifier, Classification) (have to look at it better)
+'Classification'(Identifier, ListOfCategories, ListOfDataRecords, Classifier, ClassificationVector)
+
+\begin{itemize}
+
+ \item Identifier gives a label for the classification set
+ \item 'Category'(Label, Type, ValuesRange) specifies which type of data is going to be classified
+ \item 'DataRecord'(RecordID, ListOfMatchingValues) is a data instance of the dataset 
+ \item 'Classifier'(Name, ...)
+ \item 'Score'(RecordID, ListOfValues) keeps track of each score (vector of scores depeding of the statistical distribution) for every record.
+\end{itemize}
+
+By defining these functors, the engine is able to interpret the data and the user has a clear semantic template to create the classification environment properly.
+We could even define some constraints and predefined functors instanes for a specific domain.
 
@@
 
-We have presented a powerful reasoner for multimedia analysis implemented as a logic-based framework that aims at converting low-level computations into shareable and distributed knowledge.
+We have presented a powerful reasoner for multimedia analysis implemented as a logic-based framework which allow us to convert low-level computations into shareable and distributed knowledge.
+The system design ''breaks apart`` traditional applications architecture to introduce logic and reasoning machinery on top of the computation engines providing a more intelligent, capable and understandable DSP reasoner.
+This reasoner can be extended and enriched by the open source APIs working on the foundations of the system and controlled by ''Prolog-speaking`` agents.
+SWI-DSP is fairly well positioned for the new web technologies and actually it finds in them the perfect semantic and storing support to overwhelm traditional and closed DSP systems.
+
+Future work and development is oriented to:
+
+\begin{itemize}
+ \item Testing and comparison of performance with other rival systems
+ \item Extending the SWI-DSP vocabulary and capabilities
+\end{itemize}
+
+SWI-DSP vocabulary and cabalities can be extended in several ways.
+First, by interfacing more open source APIs to cover the implementation of any DSP task (not necessarily restricted to audio domain).
+Secondly, by supporting symbolic data (MIDI) input which is a really important source of audio knowledge and providing analysis modules for such an input.
+Finally, other logic paradigms can contribute to enhance SWI-DSP (e.g.
+there are studies that use Inductive Logic Programming systems to extract harmonic rules or rhythm patterns given certain input descriptors [ref]).
+By providing semantics (OWL ontologies cleary point to this goal) to interconnect properly each logical module we can build a hierarchy of modules which help us to extract and share different levels of knowledge from digital signals.
+Such a system could be used either to extract low level cepstral coefficients from an input signal (mathematics level) or to obtain patterns of music (musicology level).
+
+\begin{thebibliography}{4}
+
+\bibitem{distributed} Raimond, Y., Sutton, C., Sandler, M.: A distributed data space for music-related information.
+In: Workshop on the Many Faces of Multimedia Semantic.
+ACM (Multimedia), Augsburg (2007).
+
+\bibitem{ctl} Bonner, A.J., Kifer, M.: Concurrency and Communication in Transaction Logic. In: Joint International Conference and Symposium on Logic Programming, pp. 142--156. Bonn (1996)
+
+\bibitem{tlprog} Bonner, A.J., Kifer, M.: Transaction Logic Programming. In: 10th International Conference on Logic Programming. Budapest (1993)
+
+\bibitem{clam} CLAM, \url{http://www.clam.iua.upf.edu/}
+
+\bibitem{marsyas} Music Analysis, Retrieval and Synthesis for Audio Signals. \url{http://marsyas.sness.net/}
+
+\bibitem{jmir} McKay, C.: jMIR. \url{http://jmir.sourceforge.net/}
+
+\bibitem{mo} Raimond, Y., Abdallah, S., Sandler, M., Frederick, G.: The Music Ontology. In: 8th International Conference of Music Information Retrieval. Vienna (2007)
+
+\bibitem{kmrdf} Source code project of KM-RDF. \url{http://code.google.com/p/km-rdf/w/list}
+
+\bibitem{ladspa} Linux Audio Developer Simple Plugin API. \url{http://www.ladspa.org/}
+
+\bibitem{vamp} Cannam, C.: Vamp Plugins. \url{http://www.vamp-plugins.org/}
+
+\bibitem{weka} WEKA home site \url{http://www.cs.waikato.ac.nz/ml/weka/}
+
+\end{thebibliography}
 

### 2008-06-04 paper/paper.tex

@@
 
-\section{Audio Analysis tasks and solutions}\label{sec:tasks}
+\section{Audio processing tasks and solutions}\label{sec:tasks}
 
-Audio analysis and synthesis APIs offer formal implementations for development and use of different algorithms and techniques.
-These algorithms are focused on different tasks that can be classified in these 3 groups:
+A particular application of DSP is the study of one-dimensional, time domain signals that in real world often exist as audio signals.
+Audio signals evolve in time and have acoustic properties produced by the signal structure and shape.
+Music and speech are the main domains involving audio signals.
+
+The properties of an analog and continuous signal (acoustic signals) can be studied by analysing their digital (sampled and quantified) versions.
+A one-dimension digital signal is therefore a processable sequence of values in a discretized time domain that represents the audio signal.
+Digital signals allow machine processing useful to modify the orignal signal or extract characterisitics from it.
+We can classify the digital signal processing paradigm in these main tasks:
 
 \begin{itemize}
- \item Processing and Effects
+ \item Effects
  \item Feature Extraction
@@
 
-\subsection{Signal Procesing}\label{subsec:effects}
+\subsection{Effects}\label{subsec:effects}
 
-This tasks group can be characterized by having audio signals as input and output, so an input signal is processed in such a way that we retrieve an altered, processed version of the original signal.
-Some examples of audio processing tasks are filtering, effects processing and signal generation.
+A signal may be presented in a raw or non desirable way for its analysis so a basic stage of many systems is the pre-processing.
+The signal is processed to obtain a satisfying version of it by different techniques (filtering, normalization, amplification...).
 
-One important dedicated API is the Linux Audio Developer's Simple Plugin API (LADSPA) which has been succesfully used in many systems and applications and is a widely spread implementation for algorithms development.
+It is very common in applied systems to provide tools that modify acoustic properties of the signal by introducing specific effects.
+Reverberation, delay or distortion are examples of effects that can be digitally obtained reproducing natural or instrumental alterations of the signal.
+
+In practice, we can characterize this group of tasks for its input/output relationships.
+The operation outputs a signal (which may be time, frequency or multidimensional domain) when is fed with an input signal (accepting several domains as well).
+Oscillators and synthesis systems outputs signals without input.
+
+One important dedicated API is the Linux Audio Developer's Simple Plugin API (LADSPA) \cite{ladspa}  which has been succesfully used in many systems and applications and is a widely spread implementation for algorithms development.
 We host LADSPA in SWI-DSP but there are other interesting libraries like DSSI plugins for effects and LV2 (new generation of LADSPA) which are targets for future extensions.
-(this is too simple i think)
 
@@
 
-Feature extraction is an important paradigm in signal research as it focuses on extract qualitative characteristics of the signal.
-Features are classified in different levels of abstraction which may go from frequency and time domain characteristics to high-level audio ones.
+Acoustic properties which are the substract of the psychological perspecitve of audio signals are conveyed in the signal and therefore in their digital representations.
+The so-called feature extraction paradigm aims at extracting different levels of audio features to characterize, classify and compare the audio signals (and any signal in general).
+The domain of the feature represents a qualitative side of the signal (tempo, chromatism, timbre...), so features do not convey all the signal information.
+This is interesting as distinct signals may have some very close features and differ much in others.
 
-The Vamp Plugins API offers a standard and utilities to develop feature extraction plugins that can be hosted in any application able to control the ``plugin lifecycle''.
+The Vamp Plugins API \cite{vamp} offers a framework to develop feature extraction plugins that can be hosted in any application able to control the ``plugin lifecycle''.
 We host Vamp plugins as basis for feature extraction in SWI-DSP.
-In general plugin libraries are characterized for being easy for development but application-dependent.
-By embedding plugins in SWI-DSP we offer a flexible development framework and provide an interactive interface to deal with plugins in several ways.
+In general plugin libraries are characterized for being suitable for development but application-dependent.
 
@@
 Marsyas and jMÍR integrate WEKA orienting it for music classification purporses by offering a code interface with other modules (Marsyas) or by extending the WEKA file format ARFF (jMIR).
-Both solutions still seem to lack of effective semantics to perform classification of audio data.
+Both solutions still seem to lack of effective semantics and a interactive front-end to perform classification of audio data.
 
-\section{SWI-DSP workflow}\label{sec:dspworkflow}
+\section{SWI-DSP architecture}\label{sec:architecture}
 
-We find of high interest the workflow architecture for DSP tasks.
-A complex task is then subdivided into subgoals that are achieved for the specific task-oriented modules.
-A procedure of this nature can be easily implemented combining high-level predicates with a proper matching of the instantation patterns.
-These instantation patters must have terms in common which are easily carried forward from one module to another without worring about further conversion.
-For this purpose, we define semantic datatypes as SWI-Prolog functors.
-In the same sense, we need to define a standard format for large binary data objects which are highly common in DSP tasks.
+From the previous comparison and review we consider that a innovating system should satisfy the following requirements:
+
+\begin{itemize}
+ \item Allowing workflows design
+ \item Providing an interface to interact with other related systems and applications
+ \item Featuring an unified, flexible and understandable front-end
+ \item Providing a open-development platform
+ \item Management of shareable data semantically labelled
+ \item Relying on a logic framework which allow us to use and provide new capabilities than traditional procedural implementations: declarative meaning of , reasoning, inferencing, computations as queries,
+\end{itemize}
+
+As initial step, we build a framework able to develop digital signal processing workflows based on a predicate logic implementation (SWI-Prolog) extending its vocabulary by wrapping computations under logic predicates.
+We develop task-oriented modules and define functors capturing the common datatypes allowing the communication between modules and hiding the internal implementation.
+In addition to this, we define a standard format for binary data ``spoken by every SWI-DSP module''.
 We define these utils in the \textbf{swilib} module which provides a C++ library and a SWI-Prolog module for their management.
 
-\subsection{Semantic datatypes}\label{subsec:datatypes}
+\subsection{Datatypes controlling workflows}\label{subsec:datatypes}
 
-SWI-Prolog functors are defined by their name and their arity.
-Therefore we need to recognize and set the common concepts and their properties used in audio analysis.
-In the previous section we have seen how tasks are strongly characterized by the type of input and output.
-We also want to look at the Music Ontology specification [x] as ontological description of music-related concepts including audio analysis terminology.
-We distinguish the following semantic datatypes:
+In section \ref{sec:tasks} we have seen how tasks are strongly characterized by their input/output relationship.
+The system require to define datatypes that guarantee the correct input and output matching.
+We identify these datatypes by analysing the differnt tasks explained.
+We also want to look at the Music Ontology specification \cite{mo} as a ``web ontology'' description of music-related concepts including audio analysis terminology.
+We distinguish the following datatypes:
 
 \begin{itemize}
- \item 'Signal'(Channels, SampleRate, Samples, [ListOfPCMPerChannel]): any audio data fragment is classified as a signal.
-Audio data is presented in our system as Pulse Code Modulation (PCM) which is a digital version of an analog signal sampled in equal intervals.
-Necessary parameters to understand the signal are the number of channels, the sample rate of the sampling and the resulting number of samples.
+Mercury
+Define datatypes
+\end{itemize}
 
-\item 'Frame'(Channels, SampleRate, StartingSample, [ListOfPCMPerChannel]) as pieces of a signal, frames have an important role in DSP as we need to decompose the signal in smaller frames for an efficient analysis.
-
-\item 'Feature'(Type, Timestamp, Data): extracted characteristics of a signal in some other domain are called features which are the main target of audio analysis at a low level.
-A feature term is characterized by the feature type, its interval of time and the actual feature data.
-
-\item 'Timestamp'(Start, Duration): audio properties are always necessarily bound to an interval of time defined as a timestamp in respect of the audio signal ``timeline''.
-
-\item 'Parameter'(Name, Value): parametrization of algorithms is one of the most significant issues in DSP as well.
-Algorithms vary their perfomance depending on the parameters.
-We want to generalize the parameterization for any task with generic terms.
-\end{itemize}
+We implement these datatypes in SWI-Prolog as functors (name/arity) that bind together the attributes of a concept (e.g.
+signal(Channels, SampleRate, Length, [DataIdPerChannel])).
+These datatypes (with the disadvantage of being a non-flexible model) can be interpreted to any module guaranteeing the correct instantation pattern of the predicates and ensuring a valid execution.
 
@@
 
-The so called BLOBs are objects wrapping large binary data.
+Large binary data is present in any multimedia system and needs to be treated in the most efficient and manageable way.
+The so-called BLOBs are objects wrapping large binary data.
 The representation of data in SWI-Prolog can be done through lists which can not handle such amount of data (e.g.
-signals of 1000000 samples) or through BLOB terms which are cumbersome as datatype.
+signals of 1000000 samples) or through SWI-Prolog BLOB terms which are cumbersome as datatype.
 
 We have implemented a mechanism of \textit{data identifiers} that vary incrementally and are stored in a database updated in each session.
-These identifiers, with the form \textbf{data\_x}, hide the binary data (which is internally a vector of floats) and allow us to perform different operations over the data easily with prolog predicates: \textit{data\_out(+Id, +File), data\_in(+File, +Id), data\_to\_list(+Id, -List)...} as well as control the binary data database status.
-Each module is implemented to read/write data in this fashion.
+Each identifier holds the binary representation of an arbitrary large sequence of floating point format data.
+This identifiers-based system is useful in the following ways:
 
-\subsection{Workflow}\label{subsec:worflow}
+\begin{itemize}
+ \item It hides awkward binary representations or inefficient SWI-Prolog lists by handy representations of data as atoms: e.g. \textbf{\_\_data\_435}
+ \item It implements a look-up database whose status can be easily updated and checked
+ \item It provides multiple ways to export/import data to the database through the identifiers.
+Data can be wrapped into or extracted from a blob term, displayed/read as prolog list (length restrictions) and dump into or load from a binary file
+ \item It unifies the ``raw'' data representation allowing the communication at the lowest level between modules 
+\end{itemize}
+
+The following example shows how a decoded audio file can be easily represented by the data identifiers that make easy to work with such a big sequence of data.
+\begin{verbatim}
+:-decode('../myfile.*', Signal).
+Signal = signal(2, 44100, 1200000, ['__data_0', '__data_1']).
+
+:-data_out('__data_0', '../myfilech1').
+:-data_out('__data_1', '../mufilech2').
+
+:-frame(signal(2, 44100, 1200000, ['__data_0', '__data_1']), 200, 5, frame(Ch, Sr, L, [Ch1D, Ch2D])),
+data_to_list(Ch1D, List), length(List, Llen).
+Ch1D = '__data_2'
+List = [6.45, 4.56, 3.23, 9.78, 3.26]
+Llen = 5
+\end{verbatim}
+
+\subsection{Processing workflows}\label{subsec:worflow}
 
@@
 
+\bibitem{weka} WEKA home site \url{http://www.cs.waikato.ac.nz/ml/weka/}
+
 \end{thebibliography}

### 2008-06-05 paper/paper.tex

@@
 \usepackage{url}
-\urldef{\mails}\path|{david.pastor, yves.raimond, mark.sandler}@elec.qmul.ac.uk|
+\urldef{\mails}\path|{david.pastor, yves.raimond, samer.abdallah, mark.sandler}@elec.qmul.ac.uk|
 
@@
 
-\author{David Pastor\and Yves Raimond\and Mark Sandler}
+\author{David Pastor\and Yves Raimond\and Samer Abdallah\and Mark Sandler}
 
@@
  \item Providing an interface to interact with other related systems and applications
- \item Featuring an unified, flexible and understandable front-end
- \item Providing a open-development platform
+ \item Featuring an unified, flexible, interactive and understandable front-end
+ \item Providing an open-development platform
  \item Management of shareable data semantically labelled
- \item Relying on a logic framework which allow us to use and provide new capabilities than traditional procedural implementations: declarative meaning of , reasoning, inferencing, computations as queries,
+ \item Enhancing traditional procedural implementations by building a logic-based framework of declarative nature allowing reasoning and inferencing.
 \end{itemize}
@@
 
+This layout allow us to wrap fast and efficient implementations of algorithms (in some imperative language) to provide predicates that determine the relationship between defined types (as black box computation) giving a ``functional programming'' view of the system preserving computational efficiency.
+This functional programming approach (which allow us to create workflows as rules with a procedural meaning) is motivated by the advantages of functional languages like Haskell \cite{haskell} that present a semantically oriented view of computations instead of a machine-oriented one.
+There are also several works that develop such a typed-functional language on a logic framework like lambda-Prolog \cite{lambda} or Mercury \cite{mercury}.
+However, our approach results to be more suitable to develop an open platform for research and development.
+External engines based on open source APIs provide faster computations than any functional or logic/strong-typed language and a suitable developers interface.
+Furthermore, SWI-Prolog features a very powerful, interactive and flexible front-end which is also a requirement for the proposed system.
+
 \subsection{Datatypes controlling workflows}\label{subsec:datatypes}
@@
 In section \ref{sec:tasks} we have seen how tasks are strongly characterized by their input/output relationship.
-The system require to define datatypes that guarantee the correct input and output matching.
-We identify these datatypes by analysing the differnt tasks explained.
+The system requires datatypes that guarantee the correct input and output matching.
+We identify these datatypes by analysing the tasks classification.
 We also want to look at the Music Ontology specification \cite{mo} as a ``web ontology'' description of music-related concepts including audio analysis terminology.
@@
 \begin{itemize}
-Mercury
-Define datatypes
+ \item signal: this type represents an audio signal in a physical timeline. T
 \end{itemize}
@@
 signal(Channels, SampleRate, Length, [DataIdPerChannel])).
-These datatypes (with the disadvantage of being a non-flexible model) can be interpreted to any module guaranteeing the correct instantation pattern of the predicates and ensuring a valid execution.
+These datatypes (with the disadvantage of being a non-flexible model) can be interpreted by any module guaranteeing the correct instantation pattern of the predicates and ensuring a valid execution.
 
@@
 \begin{verbatim}
-:-decode('../myfile.*', Signal).
+?-decode('../myfile.*', Signal).
 Signal = signal(2, 44100, 1200000, ['__data_0', '__data_1']).
 
-:-data_out('__data_0', '../myfilech1').
-:-data_out('__data_1', '../mufilech2').
+?-data_out('__data_0', '../myfilech1').
+?-data_out('__data_1', '../mufilech2').
 
-:-frame(signal(2, 44100, 1200000, ['__data_0', '__data_1']), 200, 5, frame(Ch, Sr, L, [Ch1D, Ch2D])),
+?-get_frame(signal(2, 44100, 1200000, ['__data_0', '__data_1']), 200, 5,
+frame(Ch, Sr, L, [Ch1D, Ch2D])),
 data_to_list(Ch1D, List), length(List, Llen).

### 2008-06-05 trunk/paper/paper.tex

@@
 
-In this paper we address this problematic situation by proposing and developing SWI-DSP, a logic-based framework for \textit{Audio Processing} (as subset of DSP) that extends the SWI-Prolog vocabulary.
+In this paper we address this problematic situation by proposing and developing SWI-DSP, a logic-based framework for \textit{Audio Processing} (as subset of DSP) that extends the SWI-Prolog \cite{swi} vocabulary.
 We rely on a logic programming language with aggregated audio analysis capabilities getting rid of the application-dependency paradigm and offering a single logic front-end.
@@
 
-We find of high interest the workflow architecture for DSP tasks.
-A complex task is then subdivided into subgoals that are achieved for the specific task-oriented modules.
-A procedure of this nature can be easily implemented combining high-level predicates with a proper matching of the instantation patterns.
-These instantation patters must have terms in common which are easily carried forward from one module to another without worring about further conversion.
-For this purpose, we define semantic datatypes as SWI-Prolog functors.
-In the same sense, we need to define a standard format for large binary data objects which are highly common in DSP tasks.
+From the previous comparison and review we consider that a innovating system should satisfy the following requirements:
+
+\begin{itemize}
+ \item Allowing workflows design
+ \item Providing an interface to interact with other related systems and applications
+ \item Featuring an unified, flexible, interactive and understandable front-end
+ \item Providing an open-development platform
+ \item Management of shareable data semantically labelled
+ \item Enhancing traditional procedural implementations by building a logic-based framework of declarative nature allowing reasoning and inferencing.
+\end{itemize}
+
+As initial step, we build a framework able to develop digital signal processing workflows based on a predicate logic implementation (SWI-Prolog) extending its vocabulary by wrapping computations under logic predicates.
+We develop task-oriented modules and define functors capturing the common datatypes allowing the communication between modules and hiding the internal implementation.
+In addition to this, we define a standard format for binary data ``spoken by every SWI-DSP module''.
 We define these utils in the \textbf{swilib} module which provides a C++ library and a SWI-Prolog module for their management.
 
-\subsection{Semantic datatypes}\label{subsec:datatypes}
+This layout allow us to wrap fast and efficient implementations of algorithms (in some imperative language) to provide predicates that determine the relationship between defined types (as black box computation) giving a ``functional programming'' view of the system preserving computational efficiency.
+This functional programming approach (which allow us to create workflows as rules with a procedural meaning) is motivated by the advantages of functional languages like Haskell \cite{haskell} that present a semantically oriented view of computations instead of a machine-oriented one.
+There are also several works that develop such a typed-functional language on a logic framework like lambda-Prolog \cite{lambda} or Mercury \cite{mercury}.
+However, our approach results to be more suitable to develop an open platform for research and development.
+External engines based on open source APIs provide faster computations than any functional or logic/strong-typed language and a suitable developers interface.
+Furthermore, SWI-Prolog features a very powerful, interactive and flexible front-end which is also a requirement for the proposed system.
 
-SWI-Prolog functors are defined by their name and their arity.
-Therefore we need to recognize and set the common concepts and their properties used in audio analysis.
-In the previous section we have seen how tasks are strongly characterized by the type of input and output.
-We also want to look at the Music Ontology specification [x] as ontological description of music-related concepts including audio analysis terminology.
-We distinguish the following semantic datatypes:
+\subsection{Datatypes controlling workflows}\label{subsec:datatypes}
+
+In section \ref{sec:tasks} we have seen how tasks are strongly characterized by their input/output relationship.
+The system requires datatypes that guarantee the correct input and output matching.
+We identify these datatypes by analysing the tasks classification.
+We also want to look at the Music Ontology specification \cite{mo} as a ``web ontology'' description of music-related concepts including audio analysis terminology.
+We distinguish the following datatypes:
 
 \begin{itemize}
- \item 'Signal'(Channels, SampleRate, Samples, [ListOfPCMPerChannel]): any audio data fragment is classified as a signal.
-Audio data is presented in our system as Pulse Code Modulation (PCM) which is a digital version of an analog signal sampled in equal intervals.
-Necessary parameters to understand the signal are the number of channels, the sample rate of the sampling and the resulting number of samples.
+ \item signal: this type represents an audio signal in a physical timeline. 
+ \item feature: this type represents some side of the audio signal properties
+ \item parameter: this type represents the different variables of the computation context (computations are not only a one-to-one function)
+ \item timestamp: this type represent the reference of a data object to a timeline
+\end{itemize}
 
-\item 'Frame'(Channels, SampleRate, StartingSample, [ListOfPCMPerChannel]) as pieces of a signal, frames have an important role in DSP as we need to decompose the signal in smaller frames for an efficient analysis.
-
-\item 'Feature'(Type, Timestamp, Data): extracted characteristics of a signal in some other domain are called features which are the main target of audio analysis at a low level.
-A feature term is characterized by the feature type, its interval of time and the actual feature data.
-
-\item 'Timestamp'(Start, Duration): audio properties are always necessarily bound to an interval of time defined as a timestamp in respect of the audio signal ``timeline''.
-
-\item 'Parameter'(Name, Value): parametrization of algorithms is one of the most significant issues in DSP as well.
-Algorithms vary their perfomance depending on the parameters.
-We want to generalize the parameterization for any task with generic terms.
-\end{itemize}
+We implement these datatypes in SWI-Prolog as functors (name/arity) that bind together the attributes of a concept (e.g.
+signal(Channels, SampleRate, Length, [DataIdPerChannel])).
+These datatypes (with the disadvantage of being a non-flexible model) can be interpreted by any module guaranteeing the correct instantation pattern of the predicates and ensuring a valid execution.
 
@@
 
-The so called BLOBs are objects wrapping large binary data.
+Large binary data is present in any multimedia system and needs to be treated in the most efficient and manageable way.
+The so-called BLOBs are objects wrapping large binary data.
 The representation of data in SWI-Prolog can be done through lists which can not handle such amount of data (e.g.
-signals of 1000000 samples) or through BLOB terms which are cumbersome as datatype.
+signals of 1000000 samples) or through SWI-Prolog BLOB terms which are cumbersome as datatype.
 
 We have implemented a mechanism of \textit{data identifiers} that vary incrementally and are stored in a database updated in each session.
-These identifiers, with the form \textbf{data\_x}, hide the binary data (which is internally a vector of floats) and allow us to perform different operations over the data easily with prolog predicates: \textit{data\_out(+Id, +File), data\_in(+File, +Id), data\_to\_list(+Id, -List)...} as well as control the binary data database status.
-Each module is implemented to read/write data in this fashion.
+Each identifier holds the binary representation of an arbitrary large sequence of floating point format data.
+This identifiers-based system is useful in the following ways:
 
-\subsection{Workflow}\label{subsec:worflow}
+\begin{itemize}
+ \item It hides awkward binary representations or inefficient SWI-Prolog lists by handy representations of data as atoms: e.g. \textbf{\_\_data\_435}
+ \item It implements a look-up database whose status can be easily updated and checked
+ \item It provides multiple ways to export/import data to the database through the identifiers.
+Data can be wrapped into or extracted from a blob term, displayed/read as prolog list (length restrictions) and dump into or load from a binary file
+ \item It unifies the ``raw'' data representation allowing the communication at the lowest level between modules 
+\end{itemize}
 
-These standards allow us to work with different modules without taking care of the specific underlying APIs types or implementation, so that we achieve a logic framework with dedicated semantics to perform audio analysis tasks as showed in \ref{fig:workflow}.
+The following example shows how a decoded audio file can be easily represented by the data identifiers that make easy to work with such a big sequence of data.
+\begin{verbatim}
+?-decode('../myfile.*', Signal).
+Signal = signal(2, 44100, 1200000, ['__data_0', '__data_1']).
+
+?-data_out('__data_0', '../myfilech1').
+?-data_out('__data_1', '../mufilech2').
+
+?-get_frame(signal(2, 44100, 1200000, ['__data_0', '__data_1']), 200, 5,
+frame(Ch, Sr, L, [Ch1D, Ch2D])),
+data_to_list(Ch1D, List), length(List, Llen).
+Ch1D = '__data_2'
+List = [6.45, 4.56, 3.23, 9.78, 3.26]
+Llen = 5
+\end{verbatim}
+
+\subsection{Processing workflows}\label{subsec:worflow}
 
@@
 \centerline{\framebox{\includegraphics[width=\columnwidth]{workflow.png}}}
-\caption{Workflow specification of the modules}
+\caption{Example of how to create a workflow: compare the tempo of a signal and a reverbered version of it}
 \label{fig:workflow}
@@
 
-We can then execute the following task as a SWI-Prolog query.
-\textit{The input audio file is decoded and then we apply a pre-processing in which we amplify the signal.
-We then apply a key detection algorithm over the signal to retrieve tonal regions}
-
-\begin{verbatim}
-
-my_task(Input, ['Parameter'(gain, 2)], KeyProfile):-
-	decode(InputFile, Signal),
-	amplify(Signal, 'Parameter'(gain, 2), SignalX2),
-	key(SignalX2, KeyProfile).
-
-KeyProfile = ['Feature'(key, 'Timestamp'(0, 23.45), 'Amajor')|...].
-
-\end{verbatim}
+Figure \ref{fig:workflow} shows a possible workflow that is designed by using the functional framework formed of datatypes, a standard format for data and the modules of SWI-Prolog predicates which are described in \ref{sec:modules}.
 
@@
 
-Each task module is composed by lower level modules which offer communication with an external API or library in charge of the computations.
-By hosting ``plugin libraries'' we can extend easily SWI-DSP vocabulary as we do not need to assemble each algorithm.
+Each task (functional) module is composed by lower level modules which offer communication with an external API or library in charge of the computations.
+By hosting ``plugin libraries'' we can extend easily the SWI-DSP vocabulary (new plugins can be added without code modifications).
 
@@
 The original signal is decomposed in overlapping blocks of data outputting frame terms through the non-deterministic predicate \textit{framing(+Signal, +StepSize, +BlockSize, -Frame)}.
+This predicate calculates the number of possible frames using the length of the Signal and the StepSize and creates a list of natural indexes which are passed to the low-level framing predicate in a non-deterministic way using member/2.
 The rest of predicates are mainly oriented to extract information from a Signal or Frame term.
@@
 So we can declare a simple predicate to extract the tempo of an audio signal: tempo(+Input, -Tempo) being Input an instantation of the Signal datatype and Tempo a list of instantations of the Feature datatype.
-This ``magic'' predicate is triggering a specific computation tool, so we could select the specific tool by extending the predicate: \textit{tempo(+Input, +Tool, -Tempo)}.
+This predicate that returns a tempo-related list of features given an input signal is indeed triggering a specific computation tool, so we could select the specific tool by extending the predicate: \textit{tempo(+Input, +Tool, -Tempo)}.
 
@@
      flatten(Features, WholeFeature).
-vamp_outputs_for(Signak, PluginKey, Output, Feature):-
+vamp_outputs_for(Signal, PluginKey, Output, Feature):-
      vmpl_load_plugin_for(PluginKey, Signal, Plugin),
@@
 
-\section{Towards a computation reasoner}\label{sec:reasoner}
+\section{DSP front-end}\label{sec:frontend}
 
-So far we have presented how SWI-DSP provides a vocabulary of logic predicates to perform audio analysis computations, but this logic front-end is still inappropiate for such tasks.
+So far we have presented how SWI-DSP provides a vocabulary of logic predicates and datatypes to perform audio analysis imitating a functional language, but this aproach is still inappropiate for such tasks.
 We have organized the modules in communication layers and task-oriented modules, but there is an important trade-off between understandability and specification.
 The communication layer even offering an application-independent access to the computation libraries still requires a high knowledge of such libraries besides a SWI-Prolog familiarization which is not expected in most part of the potential users.
-The task-oriented module, which is itself very layered, provides high-level predicates which are missing important options at setting-up.
+The datatypes are not the only necessary semantic resource to provide an actual workbench for DSP.
+We require a front-end that reflects the experimental domain providing an interactive ``laboratory''.
 
-In addition to this, we require a mechanism to update results and keep track of computation sequences, so that computations are not restricted to one SWI-Prolog session, allowing us to reuse and share them.
-This section tackles this problem by extending SWI-DSP as an expert system able to interpret data within defined ``DSP scenarios'' and by introducing Concurrent Transaction Logic on top of our extended predicate-logic language.
-
-\subsection{DSP scenarios}\label{subsec:scenarios}
-
-Our DSP workflow is strongly constrained by the predicate instantation patterns which clearly establishes a hierarchy of users depeding on the level of abstraction we are dealing with.
-Applications provide the user with a nice graphical or command line interface to control the computations.
-We need therefore to set a ``nice view'' of the system to the potential users which are not (and do not need to be) related to the SWI-DSP or the libraries implementation.
-
-We address this problem by defining ``scenarios'' that control and delimit the user's view of the system without restricting significantly his options.
-Any computation over an audio signal can be seen as a function input/ouput with a specific setup.
-The ``scenario'' characterizes such a family of funtions in a generic way so the user just needs to configure the scenario and provide the input to capture the output.
-Such computations can be wrapped as a signal \textbf{transform} which is declared as a SWI-Prolog functor:
-
-'Transform'(TransformType, 'Engine'(API, PluginName, Version), SampleRate, StepSize, BlockSize, ListOfParameters, Program).
+We define some ``DSP scenarios'' that reflect the researching activity.
+This scenarios define the necessary patterns and datatypes (implemented as predicates and functors) to allow the user to actually perform high-level research preserving the necessary flexibility.
+The elemental scenarios are:
 
 \begin{itemize}
-
- \item TransformType indicates which sort of task we are carrying out (feature extraction, audio effect...)
- \item 'Engine'(A, Pn, V) identifies the computation engine in charge of the processing of the transform
- \item SampleRate indicates the input sample rate in case it is not constant
- \item StepSize and BlockSize are the framing parameters
- \item ListOfParameters to setup the engine
- \item Program is a predetermined configuration for the engine
+ \item Transform: This scenario provides a pattern to apply any computation over an audio signal.
+It specifices by a datatype the framing parameters, the running engine, the engine configuration and parameters, the sample rate (if variable), the type of transform and the domain specifications.
+These factors are captured with a functor used to link input/output (with full flexibility) in transform(+Input, +Transform, -Output)
+ \item Classification: This scenario is provided for a proper setup of classification.
+This scenario specifies the classification domain, the categories (names, types and ranges), the classification schema and the training set.
+It also provides a pattern for the input data set which is interpreted along with the classification environment to return the resulting
 \end{itemize}
 
-Then each task module hardcodes the necessary knowledge in form of interpretation rules to read the tranform scenario and set the up the corresponding engine.
-The predicate \textit{transform(+InputSignal, +TransformContext, -Output)} will provide a generic view of processing to the user with standarized semantics.
+The patterns and functors are interpreted by hardcoded prior knowledge which starts up the subsequent steps to provide the queried output.
+More complex scenarios can be created in order to achieve higher or broader semantics for research activity.
+This interactive logical framework can be interfaced, tranlasted or embedded into other environments by means of the wide range of SWI-Prolog tools (see \ref{subsec:henry}).
 
-Another complex ``scenario'' is classification. We find this approach suitable to control swiweka effectively. Each classification context can be defined by:
+\section{Data management and computation reasoning}\label{subsec:datamanage}
 
-'Classification'(Identifier, ListOfCategories, ListOfDataRecords, Classifier, ClassificationVector)
+We still need to address, probably, the most important issue to accomplish the proposed architecture.
+We could say that such a system is indeed aimed at managing complex and large data, which is related through computational ``blocks'', and allowing reasoning about it.
+This goal requires the following achievements: an efficient storage of data objects and the relationships (predicates) linking them, mechanisms to control the concurrency, consistency and side-effects of the computations and a high-level interface to interact with them.
 
-\begin{itemize}
+We do not develop such manager within SWI-DSP as the database implementation actually depends on the needs of the host agent interacting with SWI-DSP.
+However, supposed a database we can describe the architecture in detail.
+Data is stored in the database by ``tabling'' predicates (which link the input, the context and the output).
+Tabled predicates (we table predicates that comprise a big load of computations) are stored in the database the first time they are called so data can be retrieved automatically without running the external engines simulating not functional but semantic relationships of complex (audio) data.
+For this purpose, the data identifiers mechanism seems to be ideal as data ids are linked (and binary data is dumped and loaded from external files relieving memory) and stored wraping complex data as a declarative-oriented units of knowledge.
+An example of a tabled predicate is:
 
- \item Identifier gives a label for the classification set
- \item 'Category'(Label, Type, ValuesRange) specifies which type of data is going to be classified
- \item 'DataRecord'(RecordID, ListOfMatchingValues) is a data instance of the dataset 
- \item 'Classifier'(Name, ...)
- \item 'Score'(RecordID, ListOfValues) keeps track of each score (vector of scores depeding of the statistical distribution) for every record.
-\end{itemize}
+\begin{verbatim}
+?-table(decode('../myfile', signal(2, 44100, 1200000, ['__data_345', '__data_346']))).
+\end{verbatim}
 
-By defining these functors, the engine is able to interpret the data and the user has a clear semantic template to create the classification environment properly.
-We could even define some constraints and predefined functors instanes for a specific domain.
+Our functional-view system can be seen now as a knowledge source that can be managed by combining procedural declarations of execution paths and declarative transactions and updates of data by means of Concurrent Transaction Logic.
 
-\subsection{Computation reasoning}\label{subsec:ctl}
-
-The previous abstraction of computations offers a cleaner view of the system without loosing flexibility but there are more issues to be addressed.
-The system proposed so far does not provide anything novel in respect of other approaches presented in \S 2 and even could seem even more cumbersome to deal with.
-The major advantage of using a logic framework as basis of the DSP system is the reasoning capability of such a framework.
-
-This reasoning skills are extended to computation reasoning by using the Concurrent Transaction Logic as superset of the Predicates Logic.
-Transaction Logic defines the logics that allow us to control and manage the execution of the computation and reason about the effects in these terms:
-
-\begin{itemize}
- \item Integration of updating transactions in queries with formal logics
- \item Integration of declarative and procedural meanings of computational transactions (paths)
-\end{itemize}
-
-The extended syntax provides connectives for different program states to create a complex execution path (allowing parallel/serial operations, concurrency control, loops, conditionals, recursion and subroutines).
-The declarative meaning and the procedural meaning are combined:
-
-computation A executes along path X = A is true on X
-
-Actually we are now in position to define several execution paths for the same computation, control the execution of those paths, reason about the results and the truth of such executions and store the computation sequences and data states.
-
-(need one good example here)
-
-\section{SWI-DSP in real applications}\label{sec:realapp}
-
-SWI-DSP defines therefore a logic front-end for computations in DSP ``scenarios''.
-We have described how Concurrent Transaction Logic converts this framework based on a predicate calculus into a DSP reasoner which control the execution and data of the different scenarios.
-This system still needs an implementation which allow an external user to control easily the reasoner, access the computations and visualise the results.
-
-\subsection{Languages hosting SWI-Prolog}
-
-Even when there is not current specific work on wrapping SWI-DSP in any other programming language which could make easier to develop a Graphical User Interface (as it probably would restrict the logic capabilities of the framework), we have to mention that SWI-DSP can be wrapped by other languages through the interfaces used for the communication layers (they define both ways of interaction).
+\subsection{Concurrent Transaction Logic}\label{sec:ctl}
 

### 2008-06-06 paper/paper.tex
draft version of the paper

@@
 This functional programming approach (which allow us to create workflows as rules with a procedural meaning) is motivated by the advantages of functional languages like Haskell \cite{haskell} that present a semantically oriented view of computations instead of a machine-oriented one.
-There are also several works that develop such a typed-functional language on a logic framework like lambda-Prolog \cite{lambda} or Mercury \cite{mercury}.
+There are also several works that develop such a typed-functional language on a logic framework like lambda-Prolog or Mercury \cite{mercury}.
 However, our approach results to be more suitable to develop an open platform for research and development.
@@
 \begin{itemize}
- \item signal: this type represents an audio signal in a physical timeline. T
+ \item signal: this type represents an audio signal in a physical timeline. 
+ \item feature: this type represents some side of the audio signal properties
+ \item parameter: this type represents the different variables of the computation context (computations are not only a one-to-one function)
+ \item timestamp: this type represent the reference of a data object to a timeline
 \end{itemize}
@@
 
-These standards allow us to work with different modules without taking care of the specific underlying APIs types or implementation, so that we achieve a logic framework with dedicated semantics to perform audio analysis tasks as showed in \ref{fig:workflow}.
-
 \begin{figure}
 \centerline{\framebox{\includegraphics[width=\columnwidth]{workflow.png}}}
-\caption{Workflow specification of the modules}
+\caption{Example of how to create a workflow: compare the tempo of a signal and a reverbered version of it}
 \label{fig:workflow}
@@
 
-\section{Towards a computation reasoner}\label{sec:reasoner}
+\section{DSP front-end}\label{sec:frontend}
 
-So far we have presented how SWI-DSP provides a vocabulary of logic predicates to perform audio analysis computations, but this logic front-end is still inappropiate for such tasks.
+So far we have presented how SWI-DSP provides a vocabulary of logic predicates and datatypes to perform audio analysis imitating a functional language, but this aproach is still inappropiate for such tasks.
 We have organized the modules in communication layers and task-oriented modules, but there is an important trade-off between understandability and specification.
 The communication layer even offering an application-independent access to the computation libraries still requires a high knowledge of such libraries besides a SWI-Prolog familiarization which is not expected in most part of the potential users.
-The task-oriented module, which is itself very layered, provides high-level predicates which are missing important options at setting-up.
+The datatypes are not the only necessary semantic resource to provide an actual workbench for DSP.
+We require a front-end that reflects the experimental domain providing an interactive ``laboratory''.
 
-In addition to this, we require a mechanism to update results and keep track of computation sequences, so that computations are not restricted to one SWI-Prolog session, allowing us to reuse and share them.
-This section tackles this problem by extending SWI-DSP as an expert system able to interpret data within defined ``DSP scenarios'' and by introducing Concurrent Transaction Logic on top of our extended predicate-logic language.
-
-\subsection{DSP scenarios}\label{subsec:scenarios}
-
-Our DSP workflow is strongly constrained by the predicate instantation patterns which clearly establishes a hierarchy of users depeding on the level of abstraction we are dealing with.
-Applications provide the user with a nice graphical or command line interface to control the computations.
-We need therefore to set a ``nice view'' of the system to the potential users which are not (and do not need to be) related to the SWI-DSP or the libraries implementation.
-
-We address this problem by defining ``scenarios'' that control and delimit the user's view of the system without restricting significantly his options.
-Any computation over an audio signal can be seen as a function input/ouput with a specific setup.
-The ``scenario'' characterizes such a family of funtions in a generic way so the user just needs to configure the scenario and provide the input to capture the output.
-Such computations can be wrapped as a signal \textbf{transform} which is declared as a SWI-Prolog functor:
-
-'Transform'(TransformType, 'Engine'(API, PluginName, Version), SampleRate, StepSize, BlockSize, ListOfParameters, Program).
+We define some ``DSP scenarios'' that reflect the researching activity.
+This scenarios define the necessary patterns and datatypes (implemented as predicates and functors) to allow the user to actually perform high-level research preserving the necessary flexibility.
+The elemental scenarios are:
 
 \begin{itemize}
-
- \item TransformType indicates which sort of task we are carrying out (feature extraction, audio effect...)
- \item 'Engine'(A, Pn, V) identifies the computation engine in charge of the processing of the transform
- \item SampleRate indicates the input sample rate in case it is not constant
- \item StepSize and BlockSize are the framing parameters
- \item ListOfParameters to setup the engine
- \item Program is a predetermined configuration for the engine
+ \item Transform: This scenario provides a pattern to apply any computation over an audio signal.
+It specifices by a datatype the framing parameters, the running engine, the engine configuration and parameters, the sample rate (if variable), the type of transform and the domain specifications.
+These factors are captured with a functor used to link input/output (with full flexibility) in transform(+Input, +Transform, -Output)
+ \item Classification: This scenario is provided for a proper setup of classification.
+This scenario specifies the classification domain, the categories (names, types and ranges), the classification schema and the training set.
+It also provides a pattern for the input data set which is interpreted along with the classification environment to return the resulting
 \end{itemize}
 
-Then each task module hardcodes the necessary knowledge in form of interpretation rules to read the tranform scenario and set the up the corresponding engine.
-The predicate \textit{transform(+InputSignal, +TransformContext, -Output)} will provide a generic view of processing to the user with standarized semantics.
+The patterns and functors are interpreted by hardcoded prior knowledge which starts up the subsequent steps to provide the queried output.
+More complex scenarios can be created in order to achieve higher or broader semantics for research activity.
+This interactive logical framework can be interfaced, tranlasted or embedded into other environments by means of the wide range of SWI-Prolog tools (see \ref{subsec:henry}).
 
-Another complex ``scenario'' is classification. We find this approach suitable to control swiweka effectively. Each classification context can be defined by:
+\section{Data management and computation reasoning}\label{subsec:datamanage}
 
-'Classification'(Identifier, ListOfCategories, ListOfDataRecords, Classifier, ClassificationVector)
+We still need to address, probably, the most important issue to accomplish the proposed architecture.
+We could say that such a system is indeed aimed at managing complex and large data, which is related through computational ``blocks'', and allowing reasoning about it.
+This goal requires the following achievements: an efficient storage of data objects and the relationships (predicates) linking them, mechanisms to control the concurrency, consistency and side-effects of the computations and a high-level interface to interact with them.
 
-\begin{itemize}
+We do not develop such manager within SWI-DSP as the database implementation actually depends on the needs of the host agent interacting with SWI-DSP.
+However, supposed a database we can describe the architecture in detail.
+Data is stored in the database by ``tabling'' predicates (which link the input, the context and the output).
+Tabled predicates (we table predicates that comprise a big load of computations) are stored in the database the first time they are called so data can be retrieved automatically without running the external engines simulating not functional but semantic relationships of complex (audio) data.
+For this purpose, the data identifiers mechanism seems to be ideal as data ids are linked (and binary data is dumped and loaded from external files relieving memory) and stored wraping complex data as a declarative-oriented units of knowledge.
+An example of a tabled predicate is:
 
- \item Identifier gives a label for the classification set
- \item 'Category'(Label, Type, ValuesRange) specifies which type of data is going to be classified
- \item 'DataRecord'(RecordID, ListOfMatchingValues) is a data instance of the dataset 
- \item 'Classifier'(Name, ...)
- \item 'Score'(RecordID, ListOfValues) keeps track of each score (vector of scores depeding of the statistical distribution) for every record.
-\end{itemize}
+\begin{verbatim}
+?-table(decode('../myfile', signal(2, 44100, 1200000, ['__data_345', '__data_346']))).
+\end{verbatim}
 
-By defining these functors, the engine is able to interpret the data and the user has a clear semantic template to create the classification environment properly.
-We could even define some constraints and predefined functors instanes for a specific domain.
+The system that was relying in a functional view of computations can be seen now as a knowledge source that can be managed by combining procedural declarations of execution paths and declarative transactions and updates of data by means of Concurrent Transaction Logic.
+By adopting a formalized mechanism to manage data we are not only achieving a higher level of abstraction but we provide the system with tools to store/export/import data.
+The system is then open to interact with other environments and applications (saving data format/type conversion issues) allowing a researcher or user to extend his workspace and data resources.
 
-\subsection{Computation reasoning}\label{subsec:ctl}
+\subsection{Concurrent Transaction Logic}\label{sec:ctl}
 
-The previous abstraction of computations offers a cleaner view of the system without loosing flexibility but there are more issues to be addressed.
-The system proposed so far does not provide anything novel in respect of other approaches presented in \S 2 and even could seem even more cumbersome to deal with.
-The major advantage of using a logic framework as basis of the DSP system is the reasoning capability of such a framework.
+Concurrent Transaction Logic \cite{ctl} extends the predicate logic to unify in a declarative formalism the control of an underlying database (SWI-Prolog and other prolog implementations just rely on assert and retract predicates for data updating, being very limited).
+Concurrent Transaction Logic allows a declarative definition of execution paths and data updates by providing new connectives to express complex data transactions: parallel and serial conjunctions, execution in isolation and executability checking.
+Rules define possible execution paths for a goal with a higher-level meaning than just declared procedures as we are certainly dealing with actions (and their effects) instead of sequences of predicates.
 
-This reasoning skills are extended to computation reasoning by using the Concurrent Transaction Logic as superset of the Predicates Logic.
-Transaction Logic defines the logics that allow us to control and manage the execution of the computation and reason about the effects in these terms:
+Therefore, we can prevent side-effects before modifying the data state, reason about the actions and their effects, control the execution of different processes and compare different paths for the same goal.
+Concurrent transaction logic applied to complex data/computational environments allows and helps to develop ``virtual laboratory workbenches'' \cite{virtual}.
 
-\begin{itemize}
- \item Integration of updating transactions in queries with formal logics
- \item Integration of declarative and procedural meanings of computational transactions (paths)
-\end{itemize}
-
-The extended syntax provides connectives for different program states to create a complex execution path (allowing parallel/serial operations, concurrency control, loops, conditionals, recursion and subroutines).
-The declarative meaning and the procedural meaning are combined:
-
-computation A executes along path X = A is true on X
-
-Actually we are now in position to define several execution paths for the same computation, control the execution of those paths, reason about the results and the truth of such executions and store the computation sequences and data states.
-
-(need one good example here)
-
-\section{SWI-DSP in real applications}\label{sec:realapp}
-
-SWI-DSP defines therefore a logic front-end for computations in DSP ``scenarios''.
-We have described how Concurrent Transaction Logic converts this framework based on a predicate calculus into a DSP reasoner which control the execution and data of the different scenarios.
-This system still needs an implementation which allow an external user to control easily the reasoner, access the computations and visualise the results.
-
-\subsection{Languages hosting SWI-Prolog}
-
-Even when there is not current specific work on wrapping SWI-DSP in any other programming language which could make easier to develop a Graphical User Interface (as it probably would restrict the logic capabilities of the framework), we have to mention that SWI-DSP can be wrapped by other languages through the interfaces used for the communication layers (they define both ways of interaction).
+SWI-DSP combined with a data manager based on Concurrent Transaction Logic (CTL) develops a multimedia knowledge system to drive research with semantics and logic formalism exploiting reasoning and declarative definition of computations and providing an interactive and shareable source of knowledge.
 
@@
 
-Although SWI-Prolog is a well-known and used implementation of Prolog it does not represent the most suitable framework to deal with multimedia data.
-The multimedia community in general needs a standard machine-readable format with proper semantics to manage data.
-The Resource Description Framework (RDF) has been proposed as such a format and defines an Standard to represent linked data by means of triples in the form of (Subject, Predicate, Object).
+We present a related project so-called Henry which develops a real CTL-based agent aimed at creating a distributed multimedia knowledge workspace built in semantic web technologies.
+The multimedia community needs a standard machine-readable format with proper semantics to manage complex data involved in any sub-domain, so data can be exchanged and shared by different systems enriching the common knowledge and resources.
 
+The Resource Description Framework (RDF) developed as the standard format for new (data-oriented) web technologies expresses linked data by means of triples in the form of (Subject, Predicate, Object).
 This standard can be implemented with different syntaxes.
-We prefer to use Notation3 [x] whose syntax allow us to write RDF statements in a human-readable way and also to write ``rules'' as named graphs [g].
-This is a suitable framework to represent data that can be shared and published in the so-called ``web of data``.
-
+We prefer Notation3 \cite{n3} whose syntax allow us to write RDF statements in a human-readable way and also to write ``rules'' as named graphs.
 The Ontology Web Language (OWL) provides the framework to express the knowledge of a particular domain in RDF through concepts and relationships.
 Thus, descripting semantic models related to audio analysis are not restricted to a local implementation but can be shared and processed by different systems.
+These elements form a suitable framework to represent and manage data and algorithms in a ``web of (complex) data''.
 
-The project called Henry implements an N3/Prolog parser and reasoner based on Concurrent Transaction Logic.
-This agent defines an entailment module capable of reasoning over the SWI-DSP computations, which are identified through Universal Resource Identifiers (URIs), and N3 statements and rules.
-The semantic web package module for SWI-Prolog interprets SPARQL queries which are passed to Henry which produces RDF descriptions of computations by performing the entailment and triggering SWI-DSP computations.
-
-\subsection{Visualisation}\label{subsec:vis}
-
-Audio data can just be poorly interpreted in a float/binary format.
-Time and frequency domain signals and features in other domains are meaningful for an expert when they are visually represented.
-Although a logic domain does not seem to be the perfect environment to visualise audio data, we can easily save this difficulty.
-
-By controlling SWI-DSP by means of Henry we can easily convert our Prolog types and binary data into RDF (machine-readable) descriptions where each source of data is represented through a link or as a ''literal`` in a basic datatype format.
-RDF descriptions can be easily browsed with the new extensions of classical web browsers for linked data.
-Even more interesting, dedicated agents with an SPARQL point are able to query DSP through the semantic web resources as explained in \ref{subsec:henry}.
-Thus, we are able to visualise our own computations but also a scalable web of data with shared semantics which is enriched with computations from other local machines.
-
-(I dont know if i should mention SV and the new extensions here....)
+Henry develops an N3/Prolog parser and reasoner based on CTL. (Text for Yves). Talk about results and performance...
+(talk about data visualisation?)
 
@@
 
+(may have to reword this after all the changes)
 We have presented a powerful reasoner for multimedia analysis implemented as a logic-based framework which allow us to convert low-level computations into shareable and distributed knowledge.
@@
 
+\bibitem{virtual} Bonner, A.J., Shrufi, A., Rozen, S.: LabFlow-1: a Database Benchmark for High-Throughput Workflow Management.
+In: 5th International Conference on Extending Database Technology, pp.
+463-478.
+Springer-Verlag, Lecture Notes in Computer Science, volume 1057.
+Avignon (1996).
+
 \bibitem{clam} CLAM, \url{http://www.clam.iua.upf.edu/}
@@
 
+\bibitem{mercury} Mercury Home Project \url{http://www.cs.mu.oz.au/research/mercury/index.html}
+
+\bibitem{haskell} Haskell \url{http://www.haskell.org/}
+
+\bibitem{swi} SWI-Prolog Home \url{http://www.swi-prolog.org/}
+
+\bibitem{n3} Notation3 \url{http://www.w3.org/DesignIssues/Notation3}
+
 \end{thebibliography}

### 2008-06-06 trunk/paper/paper.tex
Just a few rewording Your sentences are really too long :-)

@@
 The system provides the user with a unified and flexible front-end to define signal processing workflows.
-The framework develops a signal processing-enabled reasoner capable of performing the primary tasks through the interfaces: decoding (specific libraries for each format), feature extraction (Vamp plugins), audio processing and effects (LV2/LADSPA plugins) and classification (WEKA machine learning API).
+The framework develops a signal processing-enabled reasoner capable of performing the primary tasks through the following interfaces: decoding (specific libraries for each format), feature extraction (Vamp plugins), audio processing and effects (LV2/LADSPA plugins) and classification (WEKA machine learning API).
 
@@
 
-In this paper we address this problematic situation by proposing and developing SWI-DSP, a logic-based framework for \textit{Audio Processing} (as subset of DSP) that extends the SWI-Prolog \cite{swi} vocabulary.
+In this paper we address this problematic situation by proposing and developing SWI-DSP, a logic-based framework for \textit{Audio Processing} (as subset of DSP) that extends SWI-Prolog \cite{swi}.
 We rely on a logic programming language with aggregated audio analysis capabilities getting rid of the application-dependency paradigm and offering a single logic front-end.
@@
 
-In \S 2 we identify the most common audio analysis tasks and summarize several relevant \textit{open source} APIs aimed at performing such tasks.
+In \S 2 we identify the most common audio analysis tasks and summarize several relevant open source APIs aimed at performing such tasks.
 We evaluate the advantages and limitations of these APIs in order to find the requirements for SWI-DSP.
-In \S 3 we start viewing the system from the top and how it defines an interesting workflows-based architecture that characterizes input/output relationships between task-oriented modules.
+In \S 3 we start viewing the system from the top and how it defines an interesting workflow-based architecture that characterizes input/output relationships between processing tasks.
 These workflows are possible by means of common functors and a binary data standard format.
-\S 4 goes into the implementation of each task-oriented module that develops communication layers with dedicated audio processing libraries that are in charge of the actual computation and offer an open development framework.
+\S 4 goes into the implementation of the modules that develop communication layers with dedicated audio processing libraries.
+% are in charge of the actual computation and offer an open development framework.
 
-Thus, SWI-DSP extends SWI-Prolog vocabulary to perform audio analysis and processing by defining functors (capturing common datatypes) and predicates, triggering external computation performed by dedicated APIs, which are ordered in modules with different levels of abstraction.
+Thus, SWI-DSP extends SWI-Prolog for audio processing tasks by defining functors (capturing common datatypes) and predicates triggering external computation performed by dedicated APIs.
+These predicates are clustered in modules with different levels of abstraction.
 However, this architecture does not present an efficient and manageable framework as it requires specific module knowledge to create programs suitable for the concrete needs.
@@
 
-Our functional-view system can be seen now as a knowledge source that can be managed by combining procedural declarations of execution paths and declarative transactions and updates of data by means of Concurrent Transaction Logic.
+The system that was relying in a functional view of computations can be seen now as a knowledge source that can be managed by combining procedural declarations of execution paths and declarative transactions and updates of data by means of Concurrent Transaction Logic.
+By adopting a formalized mechanism to manage data we are not only achieving a higher level of abstraction but we provide the system with tools to store/export/import data.
+The system is then open to interact with other environments and applications (saving data format/type conversion issues) allowing a researcher or user to extend his workspace and data resources.
 
@@
 
+Concurrent Transaction Logic \cite{ctl} extends the predicate logic to unify in a declarative formalism the control of an underlying database (SWI-Prolog and other prolog implementations just rely on assert and retract predicates for data updating, being very limited).
+Concurrent Transaction Logic allows a declarative definition of execution paths and data updates by providing new connectives to express complex data transactions: parallel and serial conjunctions, execution in isolation and executability checking.
+Rules define possible execution paths for a goal with a higher-level meaning than just declared procedures as we are certainly dealing with actions (and their effects) instead of sequences of predicates.
+
+Therefore, we can prevent side-effects before modifying the data state, reason about the actions and their effects, control the execution of different processes and compare different paths for the same goal.
+Concurrent transaction logic applied to complex data/computational environments allows and helps to develop ``virtual laboratory workbenches'' \cite{virtual}.
+
+SWI-DSP combined with a data manager based on Concurrent Transaction Logic (CTL) develops a multimedia knowledge system to drive research with semantics and logic formalism exploiting reasoning and declarative definition of computations and providing an interactive and shareable source of knowledge.
+
 \subsection{Henry}\label{subsec:henry}
 
-Although SWI-Prolog is a well-known and used implementation of Prolog it does not represent the most suitable framework to deal with multimedia data.
-The multimedia community in general needs a standard machine-readable format with proper semantics to manage data.
-The Resource Description Framework (RDF) has been proposed as such a format and defines an Standard to represent linked data by means of triples in the form of (Subject, Predicate, Object).
+We present a related project so-called Henry which develops a real CTL-based agent aimed at creating a distributed multimedia knowledge workspace built in semantic web technologies.
+The multimedia community needs a standard machine-readable format with proper semantics to manage complex data involved in any sub-domain, so data can be exchanged and shared by different systems enriching the common knowledge and resources.
 
+The Resource Description Framework (RDF) developed as the standard format for new (data-oriented) web technologies expresses linked data by means of triples in the form of (Subject, Predicate, Object).
 This standard can be implemented with different syntaxes.
-We prefer to use Notation3 [x] whose syntax allow us to write RDF statements in a human-readable way and also to write ``rules'' as named graphs [g].
-This is a suitable framework to represent data that can be shared and published in the so-called ``web of data``.
-
+We prefer Notation3 \cite{n3} whose syntax allow us to write RDF statements in a human-readable way and also to write ``rules'' as named graphs.
 The Ontology Web Language (OWL) provides the framework to express the knowledge of a particular domain in RDF through concepts and relationships.
 Thus, descripting semantic models related to audio analysis are not restricted to a local implementation but can be shared and processed by different systems.
+These elements form a suitable framework to represent and manage data and algorithms in a ``web of (complex) data''.
 
-The project called Henry implements an N3/Prolog parser and reasoner based on Concurrent Transaction Logic.
-This agent defines an entailment module capable of reasoning over the SWI-DSP computations, which are identified through Universal Resource Identifiers (URIs), and N3 statements and rules.
-The semantic web package module for SWI-Prolog interprets SPARQL queries which are passed to Henry which produces RDF descriptions of computations by performing the entailment and triggering SWI-DSP computations.
-
-\subsection{Visualisation}\label{subsec:vis}
-
-Audio data can just be poorly interpreted in a float/binary format.
-Time and frequency domain signals and features in other domains are meaningful for an expert when they are visually represented.
-Although a logic domain does not seem to be the perfect environment to visualise audio data, we can easily save this difficulty.
-
-By controlling SWI-DSP by means of Henry we can easily convert our Prolog types and binary data into RDF (machine-readable) descriptions where each source of data is represented through a link or as a ''literal`` in a basic datatype format.
-RDF descriptions can be easily browsed with the new extensions of classical web browsers for linked data.
-Even more interesting, dedicated agents with an SPARQL point are able to query DSP through the semantic web resources as explained in \ref{subsec:henry}.
-Thus, we are able to visualise our own computations but also a scalable web of data with shared semantics which is enriched with computations from other local machines.
-
-(I dont know if i should mention SV and the new extensions here....)
+Henry develops an N3/Prolog parser and reasoner based on CTL. (Text for Yves). Talk about results and performance...
+(talk about data visualisation?)
 

### 2008-06-06 paper/paper.tex
Some more re-wording

@@
 Computations can not often be retrieved from session to session or are stored in local filesystems unlikely directed to extend any shareable source of knowledge because they are not expressed in a standard format with common semantics.
-This problem also encourages the approach presented in this work, based on a predicate calculus implementation (SWI-Prolog), in two different ways:
+This problem also encourages the logic programming approach presented in this work in two different ways:
+% FIXME - not sure about this last sentence and the two following points. Perhaps just take the second point, but put it elsewhere
 
 \begin{itemize}
- \item Concurrent Transcation Logic can be used to extend the predicate calculus to provide an unified logic framework for computations and data updating (given a database implementation).
+ \item Concurrent Transaction Logic can be used to provide a unified logic framework for computation and data updating (given an underlying database);
  \item SWI-Prolog provides an important set of interfaces to other languages (to build user applications embedding the system) and packages that allow us to interact with semantic web technologies which seem to be the best framework to provide a common source of knowledge.
@@
 
+% FIXME - in 6, make sure you cover samer's KM - cite our AES paper, for example
 Although SWI-DSP does not develop a data management system itself, we discuss this problem in \S 6 and present related works which contribute to achieve such a multimedia knowledge reasoner/manager.
@@
 
-A particular application of DSP is the study of one-dimensional, time domain signals that in real world often exist as audio signals.
-Audio signals evolve in time and have acoustic properties produced by the signal structure and shape.
-Music and speech are the main domains involving audio signals.
+A particular application of DSP is the study of one-dimensional, time domain signals.  %that in real world often exist as audio signals. 
+Audio signals evolve in time and have acoustic properties produced by the signal structure and shape. Music and speech are the main domains involving audio signals.
 
@@
 A one-dimension digital signal is therefore a processable sequence of values in a discretized time domain that represents the audio signal.
-Digital signals allow machine processing useful to modify the orignal signal or extract characterisitics from it.
-We can classify the digital signal processing paradigm in these main tasks:
+Digital signals can be automatically processed, to modify them or to extract some of their characterisitics.
+%We can classify the different digital signal processing paradigm in these three main tasks:
+We consider the following digital signal processing tasks.
 
 \begin{itemize}
- \item Effects
- \item Feature Extraction
- \item Classification
+ \item Effects;
+ \item Feature Extraction;
+ \item Classification.
 \end{itemize}
@@
 \subsection{Effects}\label{subsec:effects}
-
 A signal may be presented in a raw or non desirable way for its analysis so a basic stage of many systems is the pre-processing.

### 2008-06-06 trunk/paper/paper.tex
More corrections, and more FIXMEs

@@
+%FIXME - sentences are really too long
+%FIXME - check that with aspell
 \documentclass[runningheads]{llncs}
@@
 
-It is very common in applied systems to provide tools that modify acoustic properties of the signal by introducing specific effects.
+It is very common in musical systems to provide tools that modify acoustic properties of the signal by introducing specific effects.
 Reverberation, delay or distortion are examples of effects that can be digitally obtained reproducing natural or instrumental alterations of the signal.
 
-In practice, we can characterize this group of tasks for its input/output relationships.
-The operation outputs a signal (which may be time, frequency or multidimensional domain) when is fed with an input signal (accepting several domains as well).
-Oscillators and synthesis systems outputs signals without input.
+In practice, we characterize this group of tasks by its input/output relationships.
+The operation outputs a signal (which may be time, frequency or multidimensional domain) when it is fed with an input signal.
+% (accepting several domains as well).
+Oscillators and synthesis systems outputs signals without any inputs.
 
@@
 
-Acoustic properties which are the substract of the psychological perspecitve of audio signals are conveyed in the signal and therefore in their digital representations.
+Acoustic properties which are the substract of the psychological perspective of audio signals are conveyed in the signal and therefore in their digital representations.
 The so-called feature extraction paradigm aims at extracting different levels of audio features to characterize, classify and compare the audio signals (and any signal in general).
@@
 \begin{itemize}
- \item They do not offer a common format to exchange data or an easy way to interact with other applications.
+ \item They do not offer a common format to exchange data or an easy way to interact with other applications;
  \item They are based on a procedural system unable to reach higher-level knowledge by itself.
@@
 However, this generic API does not provide semantics to create a dataset which is the most intriguing and complex issue for classifying data properly.
-Marsyas and jMÍR integrate WEKA orienting it for music classification purporses by offering a code interface with other modules (Marsyas) or by extending the WEKA file format ARFF (jMIR).
+Marsyas and jMIR integrate WEKA orienting it for music classification purporses by offering a code interface with other modules (Marsyas) or by extending the WEKA file format ARFF (jMIR).
 Both solutions still seem to lack of effective semantics and a interactive front-end to perform classification of audio data.
@@
 \section{SWI-DSP architecture}\label{sec:architecture}
-
-From the previous comparison and review we consider that a innovating system should satisfy the following requirements:
+;
+From the previous comparison and review we consider the following requirements:
 
 \begin{itemize}
- \item Allowing workflows design
- \item Providing an interface to interact with other related systems and applications
- \item Featuring an unified, flexible, interactive and understandable front-end
- \item Providing an open-development platform
- \item Management of shareable data semantically labelled
+ \item Allowing workflow design;
+ \item Providing an interface to interact with other related systems and applications;
+ \item Featuring an unified, flexible, interactive and understandable front-end;
+ \item Providing an open-development platform;
+ \item Management of shareable data semantically labelled;
  \item Enhancing traditional procedural implementations by building a logic-based framework of declarative nature allowing reasoning and inferencing.
@@
 We develop task-oriented modules and define functors capturing the common datatypes allowing the communication between modules and hiding the internal implementation.
-In addition to this, we define a standard format for binary data ``spoken by every SWI-DSP module''.
+In addition to this, we define a standard format for binary data exchanged amongst processing modules.
 We define these utils in the \textbf{swilib} module which provides a C++ library and a SWI-Prolog module for their management.
 
+% FIXME - I am really not sure about this bit. some sentences are highly unclear, and i think that might really confuse the reader
+% If you're going to speak about functional programming, introduce it nicely.
+I am not sure "semantically oriented view of computations instead of a machine-oriented one" is really a good description of Haskell?
 This layout allow us to wrap fast and efficient implementations of algorithms (in some imperative language) to provide predicates that determine the relationship between defined types (as black box computation) giving a ``functional programming'' view of the system preserving computational efficiency.
@@
 We identify these datatypes by analysing the tasks classification.
-We also want to look at the Music Ontology specification \cite{mo} as a ``web ontology'' description of music-related concepts including audio analysis terminology.
+We also want to look at the Music Ontology specification \cite{mo} as a formal description of music-related concepts including audio analysis terminology.
 We distinguish the following datatypes:
@@
 \begin{itemize}
- \item signal: this type represents an audio signal in a physical timeline. 
- \item feature: this type represents some side of the audio signal properties
- \item parameter: this type represents the different variables of the computation context (computations are not only a one-to-one function)
- \item timestamp: this type represent the reference of a data object to a timeline
+ \item signal: this type represents an audio signal in a physical timeline;
+ \item feature: this type represents some side of the audio signal properties;
+ \item parameter: this type represents the different variables of the computation context (computations are not only a one-to-one function);
+ \item timestamp: this type represent the reference of a data object to a timeline.
 \end{itemize}
@@
 signal(Channels, SampleRate, Length, [DataIdPerChannel])).
-These datatypes (with the disadvantage of being a non-flexible model) can be interpreted by any module guaranteeing the correct instantation pattern of the predicates and ensuring a valid execution.
+These datatypes (with the disadvantage of being a non-flexible model) can be interpreted by any module guaranteeing the correct instantiation pattern of the predicates and ensuring a valid execution.
 
@@
 
+% FIXME - database? persistent?
 We have implemented a mechanism of \textit{data identifiers} that vary incrementally and are stored in a database updated in each session.
@@
 \begin{itemize}
- \item It hides awkward binary representations or inefficient SWI-Prolog lists by handy representations of data as atoms: e.g. \textbf{\_\_data\_435}
- \item It implements a look-up database whose status can be easily updated and checked
+ \item It hides the raw binary representations or inefficient SWI-Prolog lists by handy representations of data as atoms: e.g. \textbf{\_\_data\_435};
+ \item It implements a look-up database whose status can be easily updated and checked;
  \item It provides multiple ways to export/import data to the database through the identifiers.
-Data can be wrapped into or extracted from a blob term, displayed/read as prolog list (length restrictions) and dump into or load from a binary file
- \item It unifies the ``raw'' data representation allowing the communication at the lowest level between modules 
+Data can be wrapped into or extracted from a blob term, displayed/read as prolog list (length restrictions) and dump into or load from a binary file;
+ \item It unifies the ``raw'' data representation allowing the communication at the lowest level between modules.
 \end{itemize}

### 2008-06-06 trunk/paper/paper.tex
Some changes - FIXMEs, typos

@@
 %FIXME - check that with aspell
+%FIXME - make sure you use `task-oriented' only in appropriate places (eg. feature extraction)
 \documentclass[runningheads]{llncs}
@@
 
+% FIXME - refer to our AES paper here - and rely on it for the description of function tabling
 We do not develop such manager within SWI-DSP as the database implementation actually depends on the needs of the host agent interacting with SWI-DSP.
@@
 
+Such data management is handled by the framework described in \cite{UsAES2006}.
+
+% FIXME - highly unclear, especially as you introduce CTR afterwards. I think you should rewrite that.
+% Also, the export/import of data is really too quickly introduced, and probably not in the right place?
 The system that was relying in a functional view of computations can be seen now as a knowledge source that can be managed by combining procedural declarations of execution paths and declarative transactions and updates of data by means of Concurrent Transaction Logic.
@@
 
+% FIXME - perhaps just "transaction logic" would do here - be gentle for the reader:-)
 \subsection{Concurrent Transaction Logic}\label{sec:ctl}
 
-Concurrent Transaction Logic \cite{ctl} extends the predicate logic to unify in a declarative formalism the control of an underlying database (SWI-Prolog and other prolog implementations just rely on assert and retract predicates for data updating, being very limited).
+Concurrent Transaction Logic \cite{ctl}  unifies in a declarative formalism the control of an underlying database (SWI-Prolog and other prolog implementations just rely on assert and retract predicates for data updating, being very limited).
 Concurrent Transaction Logic allows a declarative definition of execution paths and data updates by providing new connectives to express complex data transactions: parallel and serial conjunctions, execution in isolation and executability checking.
@@
 
-Therefore, we can prevent side-effects before modifying the data state, reason about the actions and their effects, control the execution of different processes and compare different paths for the same goal.
-Concurrent transaction logic applied to complex data/computational environments allows and helps to develop ``virtual laboratory workbenches'' \cite{virtual}.
+Therefore, we can manage the side effects of computation on an underlying database in a consistent way, reason about the actions and their effects, control the execution of different processes and compare different paths for the same goal.
+Concurrent transaction logic applied to complex data/computational environments was succesffully used to to develop ``virtual laboratory workbenches'' \cite{virtual}.
 
-SWI-DSP combined with a data manager based on Concurrent Transaction Logic (CTL) develops a multimedia knowledge system to drive research with semantics and logic formalism exploiting reasoning and declarative definition of computations and providing an interactive and shareable source of knowledge.
+SWI-DSP combined with a data manager based on Concurrent Transaction Logic (CTR) develops a multimedia knowledge system to drive research with semantics and logic formalism exploiting reasoning and declarative definition of computations and providing an interactive and shareable source of knowledge.
 
-\subsection{Henry}\label{subsec:henry}
+We extended the data management framework described in \cite{UsAES2006} to handle transaction logic constructs.
 
-We present a related project so-called Henry which develops a real CTL-based agent aimed at creating a distributed multimedia knowledge workspace built in semantic web technologies.
-The multimedia community needs a standard machine-readable format with proper semantics to manage complex data involved in any sub-domain, so data can be exchanged and shared by different systems enriching the common knowledge and resources.
+% FIXME - Hmm... Not good in the current state.
+%\subsection{Henry}\label{subsec:henry}
+%We present a related project so-called Henry which develops a real CTR-based agent aimed at creating a distributed multimedia knowledge workspace built on semantic web technologies.
+The multimedia community needs a standard machine-readable format with proper semantics to manage complex data involved in any sub-domain, so data can be exchanged and shared by different systems enriching the common knowledge.
 
-The Resource Description Framework (RDF) developed as the standard format for new (data-oriented) web technologies expresses linked data by means of triples in the form of (Subject, Predicate, Object).
+%The Resource Description Framework (RDF) developed as the standard format for new (data-oriented) web technologies expresses linked data by means of triples in the form of (Subject, Predicate, Object).
 This standard can be implemented with different syntaxes.
@@
 
-Henry develops an N3/Prolog parser and reasoner based on CTL. (Text for Yves). Talk about results and performance...
-(talk about data visualisation?)
+%Henry develops an N3/Prolog parser and reasoner based on CTL. (Text for Yves). Talk about results and performance...
+%(talk about data visualisation?)
 
@@
 We have presented a powerful reasoner for multimedia analysis implemented as a logic-based framework which allow us to convert low-level computations into shareable and distributed knowledge.
-The system design ''breaks apart`` traditional applications architecture to introduce logic and reasoning machinery on top of the computation engines providing a more intelligent, capable and understandable DSP reasoner.
+The system design ``breaks apart'' traditional applications architecture to introduce logic and reasoning machinery on top of the computation engines providing a more intelligent, capable and understandable DSP reasoner.
 This reasoner can be extended and enriched by the open source APIs working on the foundations of the system and controlled by ''Prolog-speaking`` agents.

### 2008-06-07 paper/paper.tex

@@
+%FIXME - sentences are really too long
+%FIXME - check that with aspell
+%FIXME - make sure you use `task-oriented' only in appropriate places (eg. feature extraction)
 \documentclass[runningheads]{llncs}
@@
 \section{Introduction}\label{sec:intro}
-
 Digital signal processing (DSP) is a broad field and an important object of engineering research.
@@
 
-In this paper we address this problematic situation by proposing and developing SWI-DSP, a logic-based framework for \textit{Audio Processing} (as subset of DSP) that extends SWI-Prolog \cite{swi}.
+In this paper we address this problematic situation by proposing and developing SWI-DSP, a logic-based framework for \textit{Audio Processing} (as subset of DSP) built on a predicate calculus implementation: SWI-Prolog \cite{swi}.
 We rely on a logic programming language with aggregated audio analysis capabilities getting rid of the application-dependency paradigm and offering a single logic front-end.
@@
 In \S 3 we start viewing the system from the top and how it defines an interesting workflow-based architecture that characterizes input/output relationships between processing tasks.
-These workflows are possible by means of common functors and a binary data standard format.
+These workflows are possible by means of common datatypes and a binary data standard format.
 \S 4 goes into the implementation of the modules that develop communication layers with dedicated audio processing libraries.
-% are in charge of the actual computation and offer an open development framework.
 
-Thus, SWI-DSP extends SWI-Prolog for audio processing tasks by defining functors (capturing common datatypes) and predicates triggering external computation performed by dedicated APIs.
+Thus, SWI-DSP develops predicates, triggering external computation performed by dedicated APIs, and datatypes that allow and constrain their communication and instantiation.
 These predicates are clustered in modules with different levels of abstraction.
@@
 Data management is an important paradigm involved in many domains and also in the multimedia research domain.
-Computations can not often be retrieved from session to session or are stored in local filesystems unlikely directed to extend any shareable source of knowledge because they are not expressed in a standard format with common semantics.
-This problem also encourages the logic programming approach presented in this work in two different ways:
-% FIXME - not sure about this last sentence and the two following points. Perhaps just take the second point, but put it elsewhere
+Computations can not often be retrieved from session to session or are stored in local filesystems.
+Local data repositories are unlikely directed to extend any shareable source of knowledge because they are not expressed in a standard format with common semantics.
 
-\begin{itemize}
- \item Concurrent Transaction Logic can be used to provide a unified logic framework for computation and data updating (given an underlying database);
- \item SWI-Prolog provides an important set of interfaces to other languages (to build user applications embedding the system) and packages that allow us to interact with semantic web technologies which seem to be the best framework to provide a common source of knowledge.
-\end{itemize}
-
-% FIXME - in 6, make sure you cover samer's KM - cite our AES paper, for example
+% FIXME - in 6, make sure you cover samer's KM - cite our AES paper, for example (Yves)
+% Samer, you could include some text there and mention the matlab interface either there or in the modules section
 Although SWI-DSP does not develop a data management system itself, we discuss this problem in \S 6 and present related works which contribute to achieve such a multimedia knowledge reasoner/manager.
@@
 
-A particular application of DSP is the study of one-dimensional, time domain signals.  %that in real world often exist as audio signals. 
-Audio signals evolve in time and have acoustic properties produced by the signal structure and shape. Music and speech are the main domains involving audio signals.
+A particular application of DSP is the study of audio signals: one-dimensional, time domain signals which also allow a representation in other domains (e.g.
+frequency).
+Audio signals evolve in time and have acoustic properties produced by the signal structure and shape.
+Music and speech are the main domains involving audio signals.
 
@@
 Digital signals can be automatically processed, to modify them or to extract some of their characterisitics.
-%We can classify the different digital signal processing paradigm in these three main tasks:
+
 We consider the following digital signal processing tasks.
@@
 
-It is very common in applied systems to provide tools that modify acoustic properties of the signal by introducing specific effects.
+It is very common in musical systems to provide tools that modify acoustic properties of the signal by introducing specific effects.
 Reverberation, delay or distortion are examples of effects that can be digitally obtained reproducing natural or instrumental alterations of the signal.
 
-In practice, we can characterize this group of tasks for its input/output relationships.
-The operation outputs a signal (which may be time, frequency or multidimensional domain) when is fed with an input signal (accepting several domains as well).
-Oscillators and synthesis systems outputs signals without input.
+In practice, we characterize this group of tasks by its input/output relationships.
+The operation outputs a signal (which may be time, frequency or multidimensional domain) when it is fed with an input signal.
+Oscillators and synthesis systems outputs signals without any inputs.
 
@@
 
-Acoustic properties which are the substract of the psychological perspecitve of audio signals are conveyed in the signal and therefore in their digital representations.
+Acoustic properties which are the substract of the psychological perspective of audio signals are conveyed in the signal and therefore in their digital representations.
 The so-called feature extraction paradigm aims at extracting different levels of audio features to characterize, classify and compare the audio signals (and any signal in general).
 The domain of the feature represents a qualitative side of the signal (tempo, chromatism, timbre...), so features do not convey all the signal information.
-This is interesting as distinct signals may have some very close features and differ much in others.
+This is interesting as distinct signals may have some very close features and differ significantly in others.
 
@@
 We host Vamp plugins as basis for feature extraction in SWI-DSP.
-In general plugin libraries are characterized for being suitable for development but application-dependent.
+In general, plugin libraries are characterized for being suitable for development but application-dependent.
 
@@
 
-With the same motivation of this work there are some integrative APIs which combines different tasks to perform low and hig-level operations in a closed implementation.
+With the same motivation of this work there are some integrative APIs which combines different tasks to perform low and high-level operations in a closed implementation.
 CLAM \cite{clam}, jMIR \cite{jmir} and Marsyas \cite{marsyas} present a complex framework collecting tools to perform different DSP tasks.
-The most interesting feature of these frameworks is the possibility to create DSP networks combining specific algorithms and modules thanks to a workflow-based architecture (Marsyas and CLAM) or with a descriptive task format (jMIR).
+The most interesting feature of these frameworks is the possibility to create DSP networks combining specific algorithms and modules thanks to a workflow-based architecture (Marsyas and CLAM) or with a descriptive data format (jMIR).
 
@@
 WEKA (Waikato Environment for Knowledge Analysis) \cite{weka} collects different machine-learning algorithms and provides several user interfaces to manage the analysis.
-It aslo provdes a formal interface (in Java) for development and interaction with other systems.
+It also provides a formal interface (in Java) for development and interaction with other systems.
 
 However, this generic API does not provide semantics to create a dataset which is the most intriguing and complex issue for classifying data properly.
-Marsyas and jMÍR integrate WEKA orienting it for music classification purporses by offering a code interface with other modules (Marsyas) or by extending the WEKA file format ARFF (jMIR).
+Marsyas and jMIR integrate WEKA orienting it for music classification purporses by offering a code interface with other modules (Marsyas) or by extending the WEKA file format ARFF (jMIR).
 Both solutions still seem to lack of effective semantics and a interactive front-end to perform classification of audio data.
@@
 
-From the previous comparison and review we consider that a innovating system should satisfy the following requirements:
+From the previous comparison and review we consider the following requirements:
 
 \begin{itemize}
- \item Allowing workflows design
- \item Providing an interface to interact with other related systems and applications
- \item Featuring an unified, flexible, interactive and understandable front-end
- \item Providing an open-development platform
- \item Management of shareable data semantically labelled
+ \item Allowing workflow design;
+ \item Providing an interface to interact with other related systems and applications;
+ \item Featuring an unified, flexible, interactive and understandable front-end;
+ \item Providing an open-development platform;
+ \item Management of shareable data semantically labelled;
  \item Enhancing traditional procedural implementations by building a logic-based framework of declarative nature allowing reasoning and inferencing.
@@
 We develop task-oriented modules and define functors capturing the common datatypes allowing the communication between modules and hiding the internal implementation.
-In addition to this, we define a standard format for binary data ``spoken by every SWI-DSP module''.
+In addition to this, we define a standard format for binary data exchanged among processing modules.
 We define these utils in the \textbf{swilib} module which provides a C++ library and a SWI-Prolog module for their management.
 
-This layout allow us to wrap fast and efficient implementations of algorithms (in some imperative language) to provide predicates that determine the relationship between defined types (as black box computation) giving a ``functional programming'' view of the system preserving computational efficiency.
-This functional programming approach (which allow us to create workflows as rules with a procedural meaning) is motivated by the advantages of functional languages like Haskell \cite{haskell} that present a semantically oriented view of computations instead of a machine-oriented one.
-There are also several works that develop such a typed-functional language on a logic framework like lambda-Prolog or Mercury \cite{mercury}.
-However, our approach results to be more suitable to develop an open platform for research and development.
+Processing libraries and analysis applications are normally implemented in fast and efficient code using some ``imperative language'' (C/C++) which is machine-operation oriented instead of focused on task knowledge.
+Functional programming languages like Haskell \cite{haskell} develop a different programming paradigm.
+Functional languages try to capture the meaning of computational task by providing a framework based on datatypes and functional relationships.
+They express ``what'' is processed instead of ``how'' it is processed.
+This seems to be a better approach to define high-level audio processing functions but the efficiency of imperative languages is strongly required in multimedia analysis.
+
+Logic programming languages have also a high-level and declarative nature.
+They combine knowledge and rules to reason about the knowledge and infer new insights.
+This is an important feature to consider a logic framework as underlying mechanism to relate the hierarchy of data involded in multimedia domains.
+However, they are not powerful as a computational machine even when it is possible to declare procedures with an execution sequence.
+There are several works that develop a typed-functional language on a logic framework like lambda-Prolog or Mercury \cite{mercury}.
+Mercury embed computational skills in a purely declarative language within a strong typed and moded logic framework.
+Mercury types and predicate modes provide the compiler with the information to generate efficient code but they restrict the interactivity of the system.
+Furthermore, even when Mercury has shown to be fast as prolog language it is still slower than code in imperative languages.
+
+Our approach results to be more suitable to develop an open platform for research and development.
 External engines based on open source APIs provide faster computations than any functional or logic/strong-typed language and a suitable developers interface.
-Furthermore, SWI-Prolog features a very powerful, interactive and flexible front-end which is also a requirement for the proposed system.
+In addition to this, SWI-Prolog features a very powerful, interactive and flexible front-end which is also a requirement for the proposed system.
 
@@
 We identify these datatypes by analysing the tasks classification.
-We also want to look at the Music Ontology specification \cite{mo} as a ``web ontology'' description of music-related concepts including audio analysis terminology.
-We distinguish the following datatypes:
+We also want to look at the Music Ontology specification \cite{mo} as a formal description of music-related concepts including audio analysis terminology.
+We distinguish the following elemental datatypes:
 
 \begin{itemize}
- \item signal: this type represents an audio signal in a physical timeline. 
- \item feature: this type represents some side of the audio signal properties
- \item parameter: this type represents the different variables of the computation context (computations are not only a one-to-one function)
- \item timestamp: this type represent the reference of a data object to a timeline
+ \item signal: this type represents an audio signal in a physical timeline;
+ \item feature: this type represents some side of the audio signal properties;
+ \item parameter: this type represents the different variables of the computation context (computations are not only a one-to-one function);
+ \item timestamp: this type represent the reference of a data object to a timeline.
 \end{itemize}
@@
 
-We have implemented a mechanism of \textit{data identifiers} that vary incrementally and are stored in a database updated in each session.
+We have implemented a mechanism of \textit{data identifiers} that vary incrementally and are stored in a structure updated at running-time (it does not develop a persistency mechanism itself).
 Each identifier holds the binary representation of an arbitrary large sequence of floating point format data.
@@
 \begin{itemize}
- \item It hides awkward binary representations or inefficient SWI-Prolog lists by handy representations of data as atoms: e.g. \textbf{\_\_data\_435}
- \item It implements a look-up database whose status can be easily updated and checked
- \item It provides multiple ways to export/import data to the database through the identifiers.
-Data can be wrapped into or extracted from a blob term, displayed/read as prolog list (length restrictions) and dump into or load from a binary file
- \item It unifies the ``raw'' data representation allowing the communication at the lowest level between modules 
+ \item It hides the raw binary representations or inefficient SWI-Prolog lists by handy representations of data as atoms: e.g. \textbf{\_\_data\_435};
+ \item It implements a look-up database whose status can be easily updated and checked;
+ \item It provides multiple ways to export/import data from the structure through the identifiers.
+Data can be wrapped into or extracted from a blob term and displayed/read as prolog list (length restrictions).
+It allows dumping data into or loading it from a binary file (using the identifiers to keep track of the data);
+ \item It unifies the ``raw'' data representation allowing the communication at the lowest level between modules.
 \end{itemize}
@@
 
-?-data_out('__data_0', '../myfilech1').
-?-data_out('__data_1', '../mufilech2').
+?-data_out('__data_0', '../__data_0').
+?-data_out('__data_1', '../__data_1').
 
@@
 
-Each task (functional) module is composed by lower level modules which offer communication with an external API or library in charge of the computations.
+Each task module (functional predicates) is composed by lower level modules which offer communication with an external API or library in charge of the computations.
 By hosting ``plugin libraries'' we can extend easily the SWI-DSP vocabulary (new plugins can be added without code modifications).
+These low level modules are not constrained by datatypes (every API deals with their own concepts) so they require a precise understanding of the API and the interface.
 
@@
 The former is a powerful module for decoding which interacts with C decoding libraries (mad, soundfile, fishsound, oggz and faad) to extract a \textit{Signal} term given and audio input file supported by the libraries.
-This process is triggered by a simple call to the predicate \textit{aspl\_decode(+InputFile, -Signal)} which recognizes the file extension and calls a specific library allowing a fast decoding process and even more important an independent term representing the encoded data in a standard way for any format.
+This process is triggered by a simple call to the predicate \textit{aspl\_decode(+InputFile, -Signal)} which recognizes the file extension and calls a specific library allowing a fast decoding process and even more important an independent term representing the encoded data in a standard way for any encoded format.
 
@@
 This predicate calculates the number of possible frames using the length of the Signal and the StepSize and creates a list of natural indexes which are passed to the low-level framing predicate in a non-deterministic way using member/2.
-The rest of predicates are mainly oriented to extract information from a Signal or Frame term.
+The rest of predicates are mainly oriented to extract information from a Signal or a Frame term.
 
@@
 
-This module defines a standard task-oriented set of predicates for feature extraction.
-As we have seen in \S 2 the module is characterized for a clear input/output relationship.
+This module defines a set of predicates and rules (specific API routines for the computations) for feature extraction.
+As we have seen in \S 2, the module is characterized for a clear input/output relationship.
 So we can declare a simple predicate to extract the tempo of an audio signal: tempo(+Input, -Tempo) being Input an instantation of the Signal datatype and Tempo a list of instantations of the Feature datatype.
-This predicate that returns a tempo-related list of features given an input signal is indeed triggering a specific computation tool, so we could select the specific tool by extending the predicate: \textit{tempo(+Input, +Tool, -Tempo)}.
+This predicate, that returns a tempo-related list of features given an input signal, is indeed triggering a specific computation tool, so we could select the specific tool by extending the predicate: \textit{tempo(+Input, +Tool, -Tempo)}.
 
 We rely on the Vamp Plugin API as it offers an efficient plugin and host implementation.
-Certainly, there is a limitation in the sense that any feature extraction algorithm we may want to use has to be implemented as Vamp Plugin, but as counterpart it is easier to write an algorithm as Vamp Plugin than to interface many APIs or particular algorithm implementations.
+Certainly, there is a limitation in the sense that any feature extraction algorithm we may want to use has to be implemented as Vamp Plugin.
+As a counterpart, it is easier to write an algorithm as Vamp Plugin than to interface many APIs or particular algorithm implementations.
 
-The sub-module \textbf{swivamp} is a communication layer based on the SWI-Prolog foreign interface which allow us to query the plugin information, set up the plugin context for the transform, execute the plugin given a Signal term and retrieve the output features for the plugin according to the lifecycle specification.
-Plugins are wrapped in a similar fashion than the audio data so we can use a plugin instance representation which allow us to have concurrent plugins working at the same time.
+The sub-module \textbf{swivamp} is a communication layer based on the SWI-Prolog foreign interface allowing us to query the plugin information, set up the plugin context for the transform, execute the plugin given a Signal term and retrieve the output features for the plugin according to the lifecycle specification.
+Plugins are wrapped in a similar fashion than the audio data so we can use a plugin instance representation (a plugin id pointing to a plugin isntance existing in memory), so that we can have concurrent plugins working at the same time.
 The previous tempo/3 predicate is declared as this set of rules:
@@
 The WEKA API is interfaced through the package JPL available for SWI-Prolog defining a complex Java-Prolog interface to read/write ARFF files, load ARFF files, instantiate classifiers, execute classification and retrieve the classification scores of the dataset.
-This communication layer offers a really cumbersome and API-dependent module which requires semantics and interpretation rules to become into a useful logic-framework for audio data classification as described in the next section.
+This communication layer offers by itself a really cumbersome and API-dependent module.
+It requires semantics and interpretation rules to become into a useful logic-framework for audio data classification as described in the next section.
 
@@
 We have organized the modules in communication layers and task-oriented modules, but there is an important trade-off between understandability and specification.
-The communication layer even offering an application-independent access to the computation libraries still requires a high knowledge of such libraries besides a SWI-Prolog familiarization which is not expected in most part of the potential users.
+The communication layer even offering an application-independent access to the computation libraries still requires a high knowledge of such libraries besides a SWI-Prolog familiarization.
+However, these requirements are not expected in most part of the potential users.
 The datatypes are not the only necessary semantic resource to provide an actual workbench for DSP.
@@
 This scenario specifies the classification domain, the categories (names, types and ranges), the classification schema and the training set.
-It also provides a pattern for the input data set which is interpreted along with the classification environment to return the resulting
+It also provides a pattern for the input data set which is interpreted along with the classification environment to return the output distribution.
 \end{itemize}
@@
 More complex scenarios can be created in order to achieve higher or broader semantics for research activity.
-This interactive logical framework can be interfaced, tranlasted or embedded into other environments by means of the wide range of SWI-Prolog tools (see \ref{subsec:henry}).
+This interactive logical framework can be interfaced, tranlasted or embedded into other environments by means of the wide range of SWI-Prolog tools as explained in the next section.
 
@@
 We still need to address, probably, the most important issue to accomplish the proposed architecture.
-We could say that such a system is indeed aimed at managing complex and large data, which is related through computational ``blocks'', and allowing reasoning about it.
+We could say that such a system is indeed aimed at managing complex and large data, which is related through functional ``blocks'', and allowing reasoning about it.
 This goal requires the following achievements: an efficient storage of data objects and the relationships (predicates) linking them, mechanisms to control the concurrency, consistency and side-effects of the computations and a high-level interface to interact with them.
@@
 We do not develop such manager within SWI-DSP as the database implementation actually depends on the needs of the host agent interacting with SWI-DSP.
-However, supposed a database we can describe the architecture in detail.
-Data is stored in the database by ``tabling'' predicates (which link the input, the context and the output).
-Tabled predicates (we table predicates that comprise a big load of computations) are stored in the database the first time they are called so data can be retrieved automatically without running the external engines simulating not functional but semantic relationships of complex (audio) data.
-For this purpose, the data identifiers mechanism seems to be ideal as data ids are linked (and binary data is dumped and loaded from external files relieving memory) and stored wraping complex data as a declarative-oriented units of knowledge.
-An example of a tabled predicate is:
+However, supposed an external database we can describe the architecture in detail.
+Data is stored in the database by ``tabling'' predicates \cite{aes2006}.
+Tabled predicates (we table predicates that comprise a big load of computations) can be seen as `` virtual tables'' of n-tuples.
+After tabling a predicate, the computation is not longer processed and the data is retrieved as in relational database query.
+This database is seen from a deductive model (facts and rules in prolog), so inferences over relational knowledge are also allowed.
+For this purpose, the data identifiers mechanism seems to be ideal as data ids are relationally linked (and binary data is dumped and loaded from external files relieving memory) and stored wrapping complex data as units of knowledge.
+The persistency of a tabled predicate is described as follows:
 
@@
 ?-table(decode('../myfile', signal(2, 44100, 1200000, ['__data_345', '__data_346']))).
+
+The underlying process in pseudo-code:
+
+table(predicate(File, signal(Ch, Sr, Length, Data)):-
+	to_database('predicate', File, Ch, Sr, Length, Data).
+to_database('predicate', File, Ch, Sr, Length, (DataID|List)):-
+	specific_implementation('predicate', File, Ch, Sr, Length, DataID),
+	data_out(DataID), 
+	%DataID is the name of the file in a predefined path of binary data
+	clean_data_from_memory(DataID),
+	to_database('predicate', File, Ch, Sr, Length, List).	
 \end{verbatim}
 
-The system that was relying in a functional view of computations can be seen now as a knowledge source that can be managed by combining procedural declarations of execution paths and declarative transactions and updates of data by means of Concurrent Transaction Logic.
-By adopting a formalized mechanism to manage data we are not only achieving a higher level of abstraction but we provide the system with tools to store/export/import data.
-The system is then open to interact with other environments and applications (saving data format/type conversion issues) allowing a researcher or user to extend his workspace and data resources.
+This data management framework described as in \cite{aes2006} can be extended by handling transaction logic constructs.
+Transaction logic provides the formalism to combine, with a declarative meaning, functional predicates and data transactions.
 
-\subsection{Concurrent Transaction Logic}\label{sec:ctl}
+\subsection{Transaction Logic}\label{sec:ctl}
 
-Concurrent Transaction Logic \cite{ctl} extends the predicate logic to unify in a declarative formalism the control of an underlying database (SWI-Prolog and other prolog implementations just rely on assert and retract predicates for data updating, being very limited).
-Concurrent Transaction Logic allows a declarative definition of execution paths and data updates by providing new connectives to express complex data transactions: parallel and serial conjunctions, execution in isolation and executability checking.
-Rules define possible execution paths for a goal with a higher-level meaning than just declared procedures as we are certainly dealing with actions (and their effects) instead of sequences of predicates.
+Transaction Logic \cite{ctl}  unifies in a declarative formalism the control of an underlying database (SWI-Prolog and other prolog implementations just rely on assert and retract predicates for data updating, being very limited).
+Transaction Logic allows a declarative definition of execution paths and data updates by providing new connectives to express complex data transactions: parallel and serial conjunctions, execution in isolation and executability checking.
+Rules define possible execution paths for a goal with a higher-level meaning than just declared procedures.
+Thus, we are certainly dealing with actions (and their effects) instead of sequences of predicates.
 
-Therefore, we can prevent side-effects before modifying the data state, reason about the actions and their effects, control the execution of different processes and compare different paths for the same goal.
-Concurrent transaction logic applied to complex data/computational environments allows and helps to develop ``virtual laboratory workbenches'' \cite{virtual}.
+We can manage the side effects of computation on an underlying database in a consistent way, reason about the actions and their effects, control the execution of different processes and compare different paths for the same goal.
+Transaction logic applied to complex data/computational environments has been successfully applied to develop ``virtual laboratory workbenches'' \cite{virtual}.
 
-SWI-DSP combined with a data manager based on Concurrent Transaction Logic (CTL) develops a multimedia knowledge system to drive research with semantics and logic formalism exploiting reasoning and declarative definition of computations and providing an interactive and shareable source of knowledge.
+SWI-DSP combined with a data manager based on Concurrent Transaction Logic (CTR), which also controls concurrent processes, develops a multimedia knowledge system to drive research with semantics and a logic formalism exploiting reasoning and declarative definition of computations and providing an interactive source of knowledge.
 
-\subsection{Henry}\label{subsec:henry}
+%Samer's KM and Henry...
 
-We present a related project so-called Henry which develops a real CTL-based agent aimed at creating a distributed multimedia knowledge workspace built in semantic web technologies.
-The multimedia community needs a standard machine-readable format with proper semantics to manage complex data involved in any sub-domain, so data can be exchanged and shared by different systems enriching the common knowledge and resources.
+% FIXME - Hmm... Not good in the current state.
+%\subsection{Henry}\label{subsec:henry}
+%We present a related project so-called Henry which develops a real CTR-based agent aimed at creating a distributed multimedia knowledge workspace built on semantic web technologies.
+The multimedia community needs a standard machine-readable format with proper semantics to manage complex data involved in any sub-domain, so data can be exchanged and shared by different systems enriching the common knowledge.
 
-The Resource Description Framework (RDF) developed as the standard format for new (data-oriented) web technologies expresses linked data by means of triples in the form of (Subject, Predicate, Object).
+%The Resource Description Framework (RDF) developed as the standard format for new (data-oriented) web technologies expresses linked data by means of triples in the form of (Subject, Predicate, Object).
 This standard can be implemented with different syntaxes.
@@
 
+\bibitem{aes2006} Abdallah, S., Raimond, Y., Sandler, M.: An ontology-based approach to information management for music analysis systems. In: 120th Audio Engineering Society. Paris (2006).
+
 \bibitem{clam} CLAM, \url{http://www.clam.iua.upf.edu/}

### 2008-06-09 trunk/paper/paper.tex

@@
-%FIXME - sentences are really too long
-%FIXME - check that with aspell
-%FIXME - make sure you use `task-oriented' only in appropriate places (eg. feature extraction)
 \documentclass[runningheads]{llncs}
@@
 \begin{abstract}
-In this paper we describe a logic-based framework for Digital Signal Processing.
+In this paper we describe a logic-based framework for Digital Signal Processing (DSP).
 This framework is implemented as a set of SWI-Prolog modules developing interfaces to heterogeneous signal processing libraries.
@@
 
-\keywords{Digital Signal Processing, SWI-prolog modules, signal processing workflows, processing-enabled reasoner, interfaced open source engines}
+\keywords{Digital Signal Processing, SWI-Prolog modules, signal processing workflows, processing-enabled reasoner, interfaced open source engines}
 \end{abstract}
@@
 
-In this paper we address this problematic situation by proposing and developing SWI-DSP, a logic-based framework for \textit{Audio Processing} (as subset of DSP) that extends SWI-Prolog \cite{swi}.
+In this paper we address this problematic situation by proposing and developing SWI-DSP, a logic-based framework for \textit{Audio Processing} (as subset of DSP) built on a predicate calculus implementation: SWI-Prolog \cite{swi}.
 We rely on a logic programming language with aggregated audio analysis capabilities getting rid of the application-dependency paradigm and offering a single logic front-end.
-Logic resources are also used to achieve a more capable DSP engine.
+Logic resources are also used to achieve a more capable DSP engine: a signal processing-enabled reasoner.
 The main and primary motivation for a logic-based implementation of such a system is to turn our conception of traditional procedural applications for computation of data towards semantic frameworks to manage knowledge involving specific computations.
 
-In \S 2 we identify the most common audio analysis tasks and summarize several relevant open source APIs aimed at performing such tasks.
+In \S 2 we identify the most common audio analysis tasks and summarise several relevant open source APIs aimed at performing such tasks.
 We evaluate the advantages and limitations of these APIs in order to find the requirements for SWI-DSP.
-In \S 3 we start viewing the system from the top and how it defines an interesting workflow-based architecture that characterizes input/output relationships between processing tasks.
-These workflows are possible by means of common functors and a binary data standard format.
-\S 4 goes into the implementation of the modules that develop communication layers with dedicated audio processing libraries.
-% are in charge of the actual computation and offer an open development framework.
+In \S 3 we start viewing the system from the top and how it defines an interesting workflow-based architecture that characterises input/output relationships between processing tasks.
+These workflows are possible by means of common datatypes and a binary data standard format.
+\S 4 goes into the implementation of the modules that develop a communication layer with dedicated audio processing libraries.
 
-Thus, SWI-DSP extends SWI-Prolog for audio processing tasks by defining functors (capturing common datatypes) and predicates triggering external computation performed by dedicated APIs.
+Thus, SWI-DSP develops predicates triggering external computation performed by dedicated APIs.
+The datatypes allow and constrain the predicates communication and instantiation.
 These predicates are clustered in modules with different levels of abstraction.
@@
 Data management is an important paradigm involved in many domains and also in the multimedia research domain.
-Computations can not often be retrieved from session to session or are stored in local filesystems unlikely directed to extend any shareable source of knowledge because they are not expressed in a standard format with common semantics.
-This problem also encourages the logic programming approach presented in this work in two different ways:
-% FIXME - not sure about this last sentence and the two following points. Perhaps just take the second point, but put it elsewhere
+Computations can not often be retrieved from session to session or are stored in local filesystems.
+Local data repositories are unlikely directed to extend any shareable source of knowledge because they are not expressed in a standard format with common semantics.
 
-\begin{itemize}
- \item Concurrent Transaction Logic can be used to provide a unified logic framework for computation and data updating (given an underlying database);
- \item SWI-Prolog provides an important set of interfaces to other languages (to build user applications embedding the system) and packages that allow us to interact with semantic web technologies which seem to be the best framework to provide a common source of knowledge.
-\end{itemize}
-
-% FIXME - in 6, make sure you cover samer's KM - cite our AES paper, for example
 Although SWI-DSP does not develop a data management system itself, we discuss this problem in \S 6 and present related works which contribute to achieve such a multimedia knowledge reasoner/manager.
@@
 
-A particular application of DSP is the study of one-dimensional, time domain signals.  %that in real world often exist as audio signals. 
-Audio signals evolve in time and have acoustic properties produced by the signal structure and shape. Music and speech are the main domains involving audio signals.
+A particular application of DSP is the study of audio signals: one-dimensional, time domain signals which also allow representations in other domains (e.g.
+frequency).
+Audio signals evolve in time and have acoustic properties produced by the signal structure and shape.
+Music and speech are the main fields involving audio signals.
 
@@
 A one-dimension digital signal is therefore a processable sequence of values in a discretized time domain that represents the audio signal.
-Digital signals can be automatically processed, to modify them or to extract some of their characterisitics.
-%We can classify the different digital signal processing paradigm in these three main tasks:
+Digital signals can be automatically processed, to modify them or to extract some of their characteristics.
+
 We consider the following digital signal processing tasks.
@@
 \subsection{Effects}\label{subsec:effects}
-A signal may be presented in a raw or non desirable way for its analysis so a basic stage of many systems is the pre-processing.
-The signal is processed to obtain a satisfying version of it by different techniques (filtering, normalization, amplification...).
+A signal may be presented in a raw or non desirable way for its analysis, so a basic stage of many systems is the pre-processing.
+The signal is processed to obtain a satisfying version of it by different techniques (filtering, normalisation, amplification...).
 
@@
 The operation outputs a signal (which may be time, frequency or multidimensional domain) when it is fed with an input signal.
-% (accepting several domains as well).
 Oscillators and synthesis systems outputs signals without any inputs.
 
-One important dedicated API is the Linux Audio Developer's Simple Plugin API (LADSPA) \cite{ladspa}  which has been succesfully used in many systems and applications and is a widely spread implementation for algorithms development.
+One important dedicated API is the Linux Audio Developer's Simple Plugin API (LADSPA) \cite{ladspa}  which has been successfully used in many systems and applications and is a widely spread implementation for algorithms development.
 We host LADSPA in SWI-DSP but there are other interesting libraries like DSSI plugins for effects and LV2 (new generation of LADSPA) which are targets for future extensions.
@@
 
-Acoustic properties which are the substract of the psychological perspective of audio signals are conveyed in the signal and therefore in their digital representations.
+Acoustic properties which are the substrate of the psychological perspective of audio signals are conveyed in the signal and therefore in their digital representations.
 The so-called feature extraction paradigm aims at extracting different levels of audio features to characterize, classify and compare the audio signals (and any signal in general).
-The domain of the feature represents a qualitative side of the signal (tempo, chromatism, timbre...), so features do not convey all the signal information.
-This is interesting as distinct signals may have some very close features and differ much in others.
+The domain of the feature represents a qualitative facet of the signal (tempo, chromatics, timbre...), so features do not convey all the signal information.
+This is interesting as distinct signals may have some very close features and differ significantly in others.
 
@@
 
-With the same motivation of this work there are some integrative APIs which combines different tasks to perform low and hig-level operations in a closed implementation.
+With the same motivation of this work there are some integrative APIs which combines different tasks to perform low and high-level operations in a closed implementation.
 CLAM \cite{clam}, jMIR \cite{jmir} and Marsyas \cite{marsyas} present a complex framework collecting tools to perform different DSP tasks.
-The most interesting feature of these frameworks is the possibility to create DSP networks combining specific algorithms and modules thanks to a workflow-based architecture (Marsyas and CLAM) or with a descriptive task format (jMIR).
+The most interesting feature of these frameworks is the possibility to create DSP networks combining specific algorithms and modules thanks to a workflow-based architecture (Marsyas and CLAM) or with a descriptive data format (jMIR).
 
-CLAM also offers a metamodel of DSP processes wich provides semantics to design the networks properly presenting a higher and more manageable level of abstraction.
+CLAM also offers a metamodel of DSP processes which provides semantics to design the networks properly presenting a higher and more manageable level of abstraction.
 However, these APIs still present two important limitations that our work is aimed at solving:
@@
 WEKA (Waikato Environment for Knowledge Analysis) \cite{weka} collects different machine-learning algorithms and provides several user interfaces to manage the analysis.
-It aslo provdes a formal interface (in Java) for development and interaction with other systems.
+It also provides a formal interface (in Java) for development and interaction with other systems.
 
 However, this generic API does not provide semantics to create a dataset which is the most intriguing and complex issue for classifying data properly.
-Marsyas and jMIR integrate WEKA orienting it for music classification purporses by offering a code interface with other modules (Marsyas) or by extending the WEKA file format ARFF (jMIR).
+Marsyas and jMIR integrate WEKA orienting it for music classification purposes by offering a code interface with other modules (Marsyas) or by extending the WEKA file format ARFF (jMIR).
 Both solutions still seem to lack of effective semantics and a interactive front-end to perform classification of audio data.
@@
 \section{SWI-DSP architecture}\label{sec:architecture}
-;
+
 From the previous comparison and review we consider the following requirements:
@@
  \item Management of shareable data semantically labelled;
- \item Enhancing traditional procedural implementations by building a logic-based framework of declarative nature allowing reasoning and inferencing.
+ \item Enhancing traditional procedural implementations by building a logic-based framework of declarative nature allowing reasoning and inferring.
 \end{itemize}
@@
 We develop task-oriented modules and define functors capturing the common datatypes allowing the communication between modules and hiding the internal implementation.
-In addition to this, we define a standard format for binary data exchanged amongst processing modules.
+In addition to this, we define a standard format for binary data exchanged among processing modules.
 We define these utils in the \textbf{swilib} module which provides a C++ library and a SWI-Prolog module for their management.
 
-% FIXME - I am really not sure about this bit. some sentences are highly unclear, and i think that might really confuse the reader
-% If you're going to speak about functional programming, introduce it nicely.
-I am not sure "semantically oriented view of computations instead of a machine-oriented one" is really a good description of Haskell?
-This layout allow us to wrap fast and efficient implementations of algorithms (in some imperative language) to provide predicates that determine the relationship between defined types (as black box computation) giving a ``functional programming'' view of the system preserving computational efficiency.
-This functional programming approach (which allow us to create workflows as rules with a procedural meaning) is motivated by the advantages of functional languages like Haskell \cite{haskell} that present a semantically oriented view of computations instead of a machine-oriented one.
-There are also several works that develop such a typed-functional language on a logic framework like lambda-Prolog or Mercury \cite{mercury}.
-However, our approach results to be more suitable to develop an open platform for research and development.
+Processing libraries and analysis applications are normally implemented in fast and efficient code using an ``imperative language'' (e.g.
+C/C++) which is machine-operation oriented instead of focused on task knowledge.
+Functional programming languages like Haskell \cite{haskell} develop a different programming paradigm.
+Functional languages try to capture the meaning of computational task by providing a framework based on datatypes and functional relationships.
+They express ``what'' is processed instead of ``how'' it is processed.
+This seems to be a better approach to define high-level audio processing functions but the efficiency of imperative languages is strongly required in multimedia analysis.
+
+Logic programming languages have also a high-level and declarative nature.
+They combine knowledge and rules to reason about the knowledge and infer new insights.
+This is an important feature to consider a logic framework as underlying mechanism to relate the hierarchy of data involved in multimedia domains.
+However, they are not as powerful as a computational machine even when it is possible to declare procedures with an execution sequence.
+There are several works that develop a typed-functional language on a logic framework like lambda-Prolog or Mercury \cite{mercury}.
+Mercury embed computational skills in a purely declarative language within a strong typed and moded logic framework.
+Mercury types and predicate modes provide the compiler with the information to generate efficient code but they restrict the interactivity of the system.
+Furthermore, even when Mercury has shown to develop faster code than any logic languange, it is still slower than code in imperative languages.
+
+Our approach results to be more suitable to develop an open platform for research and development.
 External engines based on open source APIs provide faster computations than any functional or logic/strong-typed language and a suitable developers interface.
-Furthermore, SWI-Prolog features a very powerful, interactive and flexible front-end which is also a requirement for the proposed system.
+In addition to this, SWI-Prolog features a very powerful, interactive and flexible front-end which is also a requirement for the proposed system.
 
@@
 We also want to look at the Music Ontology specification \cite{mo} as a formal description of music-related concepts including audio analysis terminology.
-We distinguish the following datatypes:
+We distinguish the following elemental datatypes:
 
@@
  \item signal: this type represents an audio signal in a physical timeline;
- \item feature: this type represents some side of the audio signal properties;
+ \item feature: this type represents some facet of the audio signal properties;
  \item parameter: this type represents the different variables of the computation context (computations are not only a one-to-one function);
@@
 
-% FIXME - database? persistent?
-We have implemented a mechanism of \textit{data identifiers} that vary incrementally and are stored in a database updated in each session.
+We have implemented a mechanism of \textit{data identifiers} that vary incrementally and are stored in a structure updated at running-time (it does not develop a persistency mechanism itself).
 Each identifier holds the binary representation of an arbitrary large sequence of floating point format data.
@@
  \item It implements a look-up database whose status can be easily updated and checked;
- \item It provides multiple ways to export/import data to the database through the identifiers.
-Data can be wrapped into or extracted from a blob term, displayed/read as prolog list (length restrictions) and dump into or load from a binary file;
+ \item It provides multiple ways to export/import data from the structure through the identifiers.
+Data can be wrapped into or extracted from a blob term and displayed/read as Prolog list (length restrictions).
+It allows dumping data into or loading it from a binary file (using the identifiers to keep track of the data);
  \item It unifies the ``raw'' data representation allowing the communication at the lowest level between modules.
@@
 The following example shows how a decoded audio file can be easily represented by the data identifiers that make easy to work with such a big sequence of data.
+A data identifier can be used in the same way as the name of a person to identify a data object which is related to other entities within a predicate.
 \begin{verbatim}
@@
 \centerline{\framebox{\includegraphics[width=\columnwidth]{workflow.png}}}
-\caption{Example of how to create a workflow: compare the tempo of a signal and a reverbered version of it}
+\caption{Example of how to create a workflow: compare the tempo of a signal and a reverberated version of it}
 \label{fig:workflow}
@@
 
-Each task (functional) module is composed by lower level modules which offer communication with an external API or library in charge of the computations.
+Each task module (functional predicates) is composed by lower level modules which offer communication with an external API or library in charge of the computations.
 By hosting ``plugin libraries'' we can extend easily the SWI-DSP vocabulary (new plugins can be added without code modifications).
+These low level modules are not constrained by datatypes (every API deals with its own concepts) so they require a precise understanding of the API and the interface.
 
@@
 The former is a powerful module for decoding which interacts with C decoding libraries (mad, soundfile, fishsound, oggz and faad) to extract a \textit{Signal} term given and audio input file supported by the libraries.
-This process is triggered by a simple call to the predicate \textit{aspl\_decode(+InputFile, -Signal)} which recognizes the file extension and calls a specific library allowing a fast decoding process and even more important an independent term representing the encoded data in a standard way for any format.
+This process is triggered by a simple call to the predicate \textit{aspl\_decode(+InputFile, -Signal)} which recognizes the file extension and calls a specific library allowing a fast decoding process and even more important an independent term representing the encoded data in a standard way for the supported formats: mp3, ogg, mp4, aac, wav and equivalents.
 
@@
 This predicate calculates the number of possible frames using the length of the Signal and the StepSize and creates a list of natural indexes which are passed to the low-level framing predicate in a non-deterministic way using member/2.
-The rest of predicates are mainly oriented to extract information from a Signal or Frame term.
+The lowest framing predicate ``chunks'' the data vector identified by the id.
+The rest of predicates are mainly oriented to extract information from a Signal or a Frame term.
 
@@
 
-This module defines a standard task-oriented set of predicates for feature extraction.
-As we have seen in \S 2 the module is characterized for a clear input/output relationship.
-So we can declare a simple predicate to extract the tempo of an audio signal: tempo(+Input, -Tempo) being Input an instantation of the Signal datatype and Tempo a list of instantations of the Feature datatype.
-This predicate that returns a tempo-related list of features given an input signal is indeed triggering a specific computation tool, so we could select the specific tool by extending the predicate: \textit{tempo(+Input, +Tool, -Tempo)}.
+This module defines a set of predicates and rules (specific API routines for the computations) for feature extraction.
+As we have seen in \S 2, the module is characterized for a clear input/output relationship.
+So we can declare a simple predicate to extract the tempo of an audio signal: tempo(+Input, -Tempo) being Input an instantiation of the Signal datatype and Tempo a list of instantiations of the Feature datatype.
+This predicate, that returns a tempo-related list of features given an input signal, is indeed triggering a specific computation tool, so we could select the specific tool by extending the predicate: \textit{tempo(+Input, +Tool, -Tempo)}.
 
 We rely on the Vamp Plugin API as it offers an efficient plugin and host implementation.
-Certainly, there is a limitation in the sense that any feature extraction algorithm we may want to use has to be implemented as Vamp Plugin, but as counterpart it is easier to write an algorithm as Vamp Plugin than to interface many APIs or particular algorithm implementations.
+Certainly, there is a limitation in the sense that any feature extraction algorithm we may want to use has to be implemented as Vamp Plugin.
+As a counterpart, it is easier to write an algorithm as Vamp Plugin than to interface many APIs or particular algorithm implementations.
 
-The sub-module \textbf{swivamp} is a communication layer based on the SWI-Prolog foreign interface which allow us to query the plugin information, set up the plugin context for the transform, execute the plugin given a Signal term and retrieve the output features for the plugin according to the lifecycle specification.
-Plugins are wrapped in a similar fashion than the audio data so we can use a plugin instance representation which allow us to have concurrent plugins working at the same time.
+The sub-module \textbf{swivamp} is a communication layer based on the SWI-Prolog foreign interface allowing us to query the plugin information, set up the plugin context for the transform, execute the plugin given a Signal term and retrieve the output features for the plugin according to the lifecycle specification.
+Plugins are wrapped in a similar fashion than the audio data so we can use a plugin instance representation (a plugin id pointing to a plugin instance existing in memory), so that we can have concurrent plugins working at the same time.
 The previous tempo/3 predicate is declared as this set of rules:
@@
 \medskip
-
 \noindent
-
 \begin{verbatim}
-
 tempo(Input, Tool, Tempo):-
@@
 
+The first rule selects the tool for the queried feature.
+The following one identifies automatically the specific plugin within the Vamp API in charge of extraction the ``tempo'' (and passes a term to unify the returned feature).
+The third predicate sets up the plugin for the processing (in this case with default parameters).
+The predicate vamp\_compute\_features acts as host retrieving and organizing the extracted features which are indeed computed by the last predicate.
+This last predicate configures the framing and calls the swivamp interface that rapidly communicates and executes the plugin (the efficiency of the extraction is preserverd).
+The features are returned as a list of terms represented using the feature datatype: feature(type, timestamp(start, duration), '\_\_data\_id').
+
 \subsection{audio processing}\label{subsec:swilasdpa}
@@
 
-So far we have presented how SWI-DSP provides a vocabulary of logic predicates and datatypes to perform audio analysis imitating a functional language, but this aproach is still inappropiate for such tasks.
+So far we have presented how SWI-DSP provides a vocabulary of logic predicates and datatypes to perform audio analysis imitating a functional language, but this approach is still inappropriate for such tasks.
 We have organized the modules in communication layers and task-oriented modules, but there is an important trade-off between understandability and specification.
-The communication layer even offering an application-independent access to the computation libraries still requires a high knowledge of such libraries besides a SWI-Prolog familiarization which is not expected in most part of the potential users.
+The communication layer even offering an application-independent access to the computation libraries still requires a high knowledge of such libraries besides a SWI-Prolog familiarization.
+However, these requirements are not expected in most part of the potential users.
 The datatypes are not the only necessary semantic resource to provide an actual workbench for DSP.
@@
  \item Transform: This scenario provides a pattern to apply any computation over an audio signal.
-It specifices by a datatype the framing parameters, the running engine, the engine configuration and parameters, the sample rate (if variable), the type of transform and the domain specifications.
+It specifies by a datatype the framing parameters, the running engine, the engine configuration and parameters, the sample rate (if variable), the type of transform and the domain specifications.
 These factors are captured with a functor used to link input/output (with full flexibility) in transform(+Input, +Transform, -Output)
@@
 More complex scenarios can be created in order to achieve higher or broader semantics for research activity.
-This interactive logical framework can be interfaced, tranlasted or embedded into other environments by means of the wide range of SWI-Prolog tools (see \ref{subsec:henry}).
 
-\section{Data management and computation reasoning}\label{subsec:datamanage}
+%Reword this
+This logic-based framework for DSP (Fig.
+\ref{fig:dspdrawing}) develops a powerful and interactive computational engine that exploits logic resources.
+Such an engine can be embedded in other environments (by SWI-Prolog interfaces to other languages) developing graphical interfaces, visualisation applications or data management systems.
+In section \ref{sec:datamanage}, we present related works that extend the predicate calculus to provide a declarative framework for reasoning.
+
+\begin{figure}
+\centerline{\framebox{\includegraphics[width=\columnwidth]{swidsp.png}}}
+\caption{Arquitecture of SWI-DSP}
+\label{fig:dspdrawing}
+\end{figure}
+
+\section{Data management and reasoning}\label{sec:datamanage}
 
 We still need to address, probably, the most important issue to accomplish the proposed architecture.
-We could say that such a system is indeed aimed at managing complex and large data, which is related through computational ``blocks'', and allowing reasoning about it.
+We could say that such a system is indeed aimed at managing complex and large data, which is related through functional ``blocks'', and allowing reasoning about it.
 This goal requires the following achievements: an efficient storage of data objects and the relationships (predicates) linking them, mechanisms to control the concurrency, consistency and side-effects of the computations and a high-level interface to interact with them.
 
-% FIXME - refer to our AES paper here - and rely on it for the description of function tabling
 We do not develop such manager within SWI-DSP as the database implementation actually depends on the needs of the host agent interacting with SWI-DSP.
-However, supposed a database we can describe the architecture in detail.
-Data is stored in the database by ``tabling'' predicates (which link the input, the context and the output).
-Tabled predicates (we table predicates that comprise a big load of computations) are stored in the database the first time they are called so data can be retrieved automatically without running the external engines simulating not functional but semantic relationships of complex (audio) data.
-For this purpose, the data identifiers mechanism seems to be ideal as data ids are linked (and binary data is dumped and loaded from external files relieving memory) and stored wraping complex data as a declarative-oriented units of knowledge.
-An example of a tabled predicate is:
+However, supposed an external database we can describe the architecture in detail.
+Data is stored in the database by ``tabling'' predicates \cite{aes2006}.
+Tabled predicates (we table predicates that comprise a big load of computations) can be seen as `` virtual tables'' of n-tuples.
+After tabling a predicate, the computation is not longer processed and the data is retrieved as in relational database query.
+This database is seen from a deductive model (facts and rules in prolog), so inferences over relational knowledge are also allowed.
+For this purpose, the data identifiers mechanism seems to be ideal as data ids are relationally linked (and binary data is dumped and loaded from external files relieving memory) and stored wrapping complex data as units of knowledge.
+The persistency of a tabled predicate is described as follows:
 
@@
 ?-table(decode('../myfile', signal(2, 44100, 1200000, ['__data_345', '__data_346']))).
+
+The underlying process in pseudo-code:
+
+table(predicate(File, signal(Ch, Sr, Length, Data)):-
+	to_database('predicate', File, Ch, Sr, Length, Data).
+to_database('predicate', File, Ch, Sr, Length, (DataID|List)):-
+	specific_implementation('predicate', File, Ch, Sr, Length, DataID),
+	data_out(DataID), 
+	%DataID is the name of the file in a predefined path of binary data
+	clean_data_from_memory(DataID),
+	to_database('predicate', File, Ch, Sr, Length, List).	
 \end{verbatim}
 
-Such data management is handled by the framework described in \cite{UsAES2006}.
+This data management framework described as in \cite{aes2006} can be extended by handling transaction logic constructs.
+Transaction logic provides the formalism to combine, with a declarative meaning, functional predicates and data transactions.
 
-% FIXME - highly unclear, especially as you introduce CTR afterwards. I think you should rewrite that.
-% Also, the export/import of data is really too quickly introduced, and probably not in the right place?
-The system that was relying in a functional view of computations can be seen now as a knowledge source that can be managed by combining procedural declarations of execution paths and declarative transactions and updates of data by means of Concurrent Transaction Logic.
-By adopting a formalized mechanism to manage data we are not only achieving a higher level of abstraction but we provide the system with tools to store/export/import data.
-The system is then open to interact with other environments and applications (saving data format/type conversion issues) allowing a researcher or user to extend his workspace and data resources.
+\subsection{Transaction Logic}\label{subsec:ctl}
 
-% FIXME - perhaps just "transaction logic" would do here - be gentle for the reader:-)
-\subsection{Concurrent Transaction Logic}\label{sec:ctl}
+Transaction Logic \cite{ctl}  unifies in a declarative formalism the control of an underlying database (SWI-Prolog and other prolog implementations just rely on assert and retract predicates for data updating, being very limited).
+Transaction Logic allows a declarative definition of execution paths and data updates by providing new connectives to express complex data transactions: parallel and serial conjunctions, execution in isolation and executability checking.
+Rules define possible execution paths for a goal with a higher-level meaning than just declared procedures.
+Thus, we are certainly dealing with actions (and their effects) instead of sequences of predicates.
 
-Concurrent Transaction Logic \cite{ctl}  unifies in a declarative formalism the control of an underlying database (SWI-Prolog and other prolog implementations just rely on assert and retract predicates for data updating, being very limited).
-Concurrent Transaction Logic allows a declarative definition of execution paths and data updates by providing new connectives to express complex data transactions: parallel and serial conjunctions, execution in isolation and executability checking.
-Rules define possible execution paths for a goal with a higher-level meaning than just declared procedures as we are certainly dealing with actions (and their effects) instead of sequences of predicates.
+We can manage the side effects of computation on an underlying database in a consistent way, reason about the actions and their effects, control the execution of different processes and compare different paths for the same goal.
+Transaction logic applied to complex data/computational environments has been successfully applied to develop ``virtual laboratory workbenches'' \cite{virtual}.
 
-Therefore, we can manage the side effects of computation on an underlying database in a consistent way, reason about the actions and their effects, control the execution of different processes and compare different paths for the same goal.
-Concurrent transaction logic applied to complex data/computational environments was succesffully used to to develop ``virtual laboratory workbenches'' \cite{virtual}.
+SWI-DSP combined with a data manager based on Concurrent Transaction Logic (CTR), which also controls concurrent processes, develops a multimedia knowledge system to drive research with semantics and a logic formalism exploiting reasoning and declarative definition of computations and providing an interactive source of knowledge.
 
-SWI-DSP combined with a data manager based on Concurrent Transaction Logic (CTR) develops a multimedia knowledge system to drive research with semantics and logic formalism exploiting reasoning and declarative definition of computations and providing an interactive and shareable source of knowledge.
-
-We extended the data management framework described in \cite{UsAES2006} to handle transaction logic constructs.
+%Samer's KM and Henry...
 
@@
 
-SWI-DSP vocabulary and cabalities can be extended in several ways.
+SWI-DSP vocabulary and capabilities can be extended in several ways.
 First, by interfacing more open source APIs to cover the implementation of any DSP task (not necessarily restricted to audio domain).
@@
 there are studies that use Inductive Logic Programming systems to extract harmonic rules or rhythm patterns given certain input descriptors [ref]).
-By providing semantics (OWL ontologies cleary point to this goal) to interconnect properly each logical module we can build a hierarchy of modules which help us to extract and share different levels of knowledge from digital signals.
+By providing semantics (OWL ontologies clearly point to this goal) to interconnect properly each logical module we can build a hierarchy of modules which help us to extract and share different levels of knowledge from digital signals.
 Such a system could be used either to extract low level cepstral coefficients from an input signal (mathematics level) or to obtain patterns of music (musicology level).
@@
 Avignon (1996).
+
+\bibitem{aes2006} Abdallah, S., Raimond, Y., Sandler, M.: An ontology-based approach to information management for music analysis systems. In: 120th Audio Engineering Society. Paris (2006).
 

### 2008-06-09 paper/paper.tex

@@
 \begin{abstract}
-In this paper we describe a logic-based framework for Digital Signal Processing.
+In this paper we describe a logic-based framework for Digital Signal Processing (DSP).
 This framework is implemented as a set of SWI-Prolog modules developing interfaces to heterogeneous signal processing libraries.
-A set of functors capturing common datatypes and a standard format for binary data are used to mediate our different analysis components.
+Prolog terms representing common datatypes and a standard format for binary data are used to mediate between the different components of the system.
 The system provides the user with a unified and flexible front-end to define signal processing workflows.
-The framework develops a signal processing-enabled reasoner capable of performing the primary tasks through the following interfaces: decoding (specific libraries for each format), feature extraction (Vamp plugins), audio processing and effects (LV2/LADSPA plugins) and classification (WEKA machine learning API).
+Using this framework, we develop a signal processing-enabled reasoner capable of performing the primary tasks through the following interfaces: decoding (specific libraries for each format), feature extraction (Vamp plugins), audio processing and effects (LV2/LADSPA plugins) and classification (WEKA machine learning API).
 
-\keywords{Digital Signal Processing, SWI-prolog modules, signal processing workflows, processing-enabled reasoner, interfaced open source engines}
+\keywords{Digital Signal Processing, SWI-Prolog modules, signal processing workflows, processing-enabled reasoner, interfaced open source engines}
 \end{abstract}
@@
 In this paper we address this problematic situation by proposing and developing SWI-DSP, a logic-based framework for \textit{Audio Processing} (as subset of DSP) built on a predicate calculus implementation: SWI-Prolog \cite{swi}.
-We rely on a logic programming language with aggregated audio analysis capabilities getting rid of the application-dependency paradigm and offering a single logic front-end.
-Logic resources are also used to achieve a more capable DSP engine.
-The main and primary motivation for a logic-based implementation of such a system is to turn our conception of traditional procedural applications for computation of data towards semantic frameworks to manage knowledge involving specific computations.
+We rely on a logic programming language with foreign language modules providing audio analysis capabilities to get rid of the application-dependency paradigm and offering a single logic-based front-end.
+Logic resources are also used to achieve a more capable DSP engine: a signal processing-enabled reasoner.
+The main and primary motivation for a logic-based implementation of such a system is to turn our conception of traditional procedural applications for computation of data towards declarative semantic frameworks to manage knowledge involving specific computations.
 
-In \S 2 we identify the most common audio analysis tasks and summarize several relevant open source APIs aimed at performing such tasks.
+In \S 2 we identify the most common audio analysis tasks and summarise several relevant open source APIs aimed at performing such tasks.
 We evaluate the advantages and limitations of these APIs in order to find the requirements for SWI-DSP.
-In \S 3 we start viewing the system from the top and how it defines an interesting workflow-based architecture that characterizes input/output relationships between processing tasks.
+In \S 3 we start viewing the system from the top and how it defines an interesting workflow-based architecture that characterises input/output relationships between processing tasks.
 These workflows are possible by means of common datatypes and a binary data standard format.
-\S 4 goes into the implementation of the modules that develop communication layers with dedicated audio processing libraries.
+\S 4 goes into the implementation of the modules that develop a communication layer with dedicated audio processing libraries.
 
-Thus, SWI-DSP develops predicates, triggering external computation performed by dedicated APIs, and datatypes that allow and constrain their communication and instantiation.
-These predicates are clustered in modules with different levels of abstraction.
-However, this architecture does not present an efficient and manageable framework as it requires specific module knowledge to create programs suitable for the concrete needs.
+Thus, SWI-DSP develops predicates triggering external computation performed by dedicated APIs.
+The datatypes allow and constrain the predicates communication and instantiation.
+These predicates are organised into modules with different levels of abstraction.
+However, this architecture lacks of an uniform interface to create programs suitable for the concrete needs.
 We tackle this problem in \S 5 by defining an unified and flexible front-end, composed by DSP ``scenarios'', which allows the user to define workflows with understandable semantics.
@@
 
-% FIXME - in 6, make sure you cover samer's KM - cite our AES paper, for example (Yves)
-% Samer, you could include some text there and mention the matlab interface either there or in the modules section
 Although SWI-DSP does not develop a data management system itself, we discuss this problem in \S 6 and present related works which contribute to achieve such a multimedia knowledge reasoner/manager.
@@
 
-A particular application of DSP is the study of audio signals: one-dimensional, time domain signals which also allow a representation in other domains (e.g.
-frequency).
+A particular application of DSP is the study of audio signals: one-dimensional, time domain signals, however, they also have representations in other domains such as frequency, time-frequency, etc.
 Audio signals evolve in time and have acoustic properties produced by the signal structure and shape.
-Music and speech are the main domains involving audio signals.
+Music and speech are the main fields involving audio signals.
 
-The properties of an analog and continuous signal (acoustic signals) can be studied by analysing their digital (sampled and quantified) versions.
-A one-dimension digital signal is therefore a processable sequence of values in a discretized time domain that represents the audio signal.
-Digital signals can be automatically processed, to modify them or to extract some of their characterisitics.
+The properties of an analog and continuous signal (acoustic signals) can be studied by analysing their digital (sampled and quantised) versions.
+A one-dimension digital signal is therefore a mapping from a discrete time domain (usually finite) to a discrete value domain and is often represented as a sequence of values associated with successive time points.
+Digital signals can be processed to make new derived signals or extract some of their characteristics.
 
@@
 \subsection{Effects}\label{subsec:effects}
-A signal may be presented in a raw or non desirable way for its analysis so a basic stage of many systems is the pre-processing.
-The signal is processed to obtain a satisfying version of it by different techniques (filtering, normalization, amplification...).
+A first step in many signal processing is a normalisation, conditioning or re-representation of the signal, which may involve several techniques such as filtering, amplification, Fourier transforming, etc.
 
@@
 
-In practice, we characterize this group of tasks by its input/output relationships.
+In practise, we characterise this group of tasks by its input/output relationships.
 The operation outputs a signal (which may be time, frequency or multidimensional domain) when it is fed with an input signal.
@@
 
-One important dedicated API is the Linux Audio Developer's Simple Plugin API (LADSPA) \cite{ladspa}  which has been succesfully used in many systems and applications and is a widely spread implementation for algorithms development.
+One important dedicated API is the Linux Audio Developer's Simple Plugin API (LADSPA) \cite{ladspa}  which has been successfully used in many systems and applications and is a widely spread implementation for algorithms development.
 We host LADSPA in SWI-DSP but there are other interesting libraries like DSSI plugins for effects and LV2 (new generation of LADSPA) which are targets for future extensions.
@@
 
-Acoustic properties which are the substract of the psychological perspective of audio signals are conveyed in the signal and therefore in their digital representations.
-The so-called feature extraction paradigm aims at extracting different levels of audio features to characterize, classify and compare the audio signals (and any signal in general).
-The domain of the feature represents a qualitative side of the signal (tempo, chromatism, timbre...), so features do not convey all the signal information.
+Acoustic properties which are the substrate of the psychological perspective of audio signals are conveyed in the signal and therefore in their digital representations.
+The so-called feature extraction paradigm aims at extracting different levels of audio features to characterise, classify and compare the audio signals (and any signal in general).
+The domain of the feature represents a qualitative facet of the signal (tempo, chromatics, timbre...), so features do not convey all the signal information.
 This is interesting as distinct signals may have some very close features and differ significantly in others.
@@
 We host Vamp plugins as basis for feature extraction in SWI-DSP.
-In general, plugin libraries are characterized for being suitable for development but application-dependent.
+In general, plugin libraries are characterised for being suitable for development but application-dependent.
 
@@
 
-CLAM also offers a metamodel of DSP processes wich provides semantics to design the networks properly presenting a higher and more manageable level of abstraction.
+CLAM also offers a metamodel of DSP processes which provides semantics to design the networks properly presenting a higher and more manageable level of abstraction.
 However, these APIs still present two important limitations that our work is aimed at solving:
@@
  \item They do not offer a common format to exchange data or an easy way to interact with other applications;
- \item They are based on a procedural system unable to reach higher-level knowledge by itself.
+ \item They are based on a procedural system unable to collect and integrate knowledge from different sources to obtain new knowledge.
 \end{itemize}
@@
 However, this generic API does not provide semantics to create a dataset which is the most intriguing and complex issue for classifying data properly.
-Marsyas and jMIR integrate WEKA orienting it for music classification purporses by offering a code interface with other modules (Marsyas) or by extending the WEKA file format ARFF (jMIR).
+Marsyas and jMIR integrate WEKA orienting it for music classification purposes by offering a code interface with other modules (Marsyas) or by extending the WEKA file format ARFF (jMIR).
 Both solutions still seem to lack of effective semantics and a interactive front-end to perform classification of audio data.
@@
 
-Processing libraries and analysis applications are normally implemented in fast and efficient code using some ``imperative language'' (C/C++) which is machine-operation oriented instead of focused on task knowledge.
-Functional programming languages like Haskell \cite{haskell} develop a different programming paradigm.
-Functional languages try to capture the meaning of computational task by providing a framework based on datatypes and functional relationships.
-They express ``what'' is processed instead of ``how'' it is processed.
-This seems to be a better approach to define high-level audio processing functions but the efficiency of imperative languages is strongly required in multimedia analysis.
+Processing libraries and analysis applications are normally implemented in fast and efficient code using an imperative language (e.g.
+C/C++) which is machine-operation oriented instead of focused on task knowledge.
+Functional programming offers a different paradigm.
+Functional languages try to capture the meaning of computational tasks in a more declarative way, using functional relationships to describe ``what'' the computation should be rather than ``how'' to do it.
+Modern functional languages like Haskell \cite{haskell} provide rich and powerful type systems describing and constraining the set of legal values a function can take as input or return as output.
+This seems to be a better approach to define high-level audio processing functions, but C/C++ is still the language for the most important open source resources.
 
@@
 They combine knowledge and rules to reason about the knowledge and infer new insights.
-This is an important feature to consider a logic framework as underlying mechanism to relate the hierarchy of data involded in multimedia domains.
-However, they are not powerful as a computational machine even when it is possible to declare procedures with an execution sequence.
+This is an important feature to consider a logic framework as underlying mechanism to relate the hierarchy of data involved in multimedia domains.
+However, they are not as powerful as a computational machine even when it is possible to declare procedures with an execution sequence.
 There are several works that develop a typed-functional language on a logic framework like lambda-Prolog or Mercury \cite{mercury}.
-Mercury embed computational skills in a purely declarative language within a strong typed and moded logic framework.
-Mercury types and predicate modes provide the compiler with the information to generate efficient code but they restrict the interactivity of the system.
-Furthermore, even when Mercury has shown to be fast as prolog language it is still slower than code in imperative languages.
+Mercury embeds computational capabilities in a purely declarative language within a strong typed and moded logic framework.
+Mercury types and predicate modes provide the compiler with the information to generate efficient code but as environment for declaring and interactive prototype, Mercury is ``less forgiving'' than Prolog.
+SWI-Prolog in particular has a very comfortable development environment and hence we use SWI-Prolog.
 
@@
 External engines based on open source APIs provide faster computations than any functional or logic/strong-typed language and a suitable developers interface.
-In addition to this, SWI-Prolog features a very powerful, interactive and flexible front-end which is also a requirement for the proposed system.
 
@@
 
-In section \ref{sec:tasks} we have seen how tasks are strongly characterized by their input/output relationship.
+In section \ref{sec:tasks} we have seen how tasks are strongly characterised by their input/output relationship.
 The system requires datatypes that guarantee the correct input and output matching.
@@
 We also want to look at the Music Ontology specification \cite{mo} as a formal description of music-related concepts including audio analysis terminology.
-We distinguish the following elemental datatypes:
+We distinguish the following datatypes to any signal processing system:
 
@@
 
-We have implemented a mechanism of \textit{data identifiers} that vary incrementally and are stored in a structure updated at running-time (it does not develop a persistency mechanism itself).
+We have implemented a mechanism of unique \textit{data identifiers} that are stored in a structure updated at runtime.
 Each identifier holds the binary representation of an arbitrary large sequence of floating point format data.
@@
  \item It provides multiple ways to export/import data from the structure through the identifiers.
-Data can be wrapped into or extracted from a blob term and displayed/read as prolog list (length restrictions).
+Data can be wrapped into or extracted from a blob term and displayed/read as Prolog list (length restrictions).
 It allows dumping data into or loading it from a binary file (using the identifiers to keep track of the data);
- \item It unifies the ``raw'' data representation allowing the communication at the lowest level between modules.
+ \item It unifies the ``raw'' data representation allowing the communication at the lowest level between modules;
+ \item One identifier can be used to retrieve data from a previous session as long as we make the data object persist under its identifier and reserve the identifier in the new session.
 \end{itemize}
@@
 
-?-data_out('__data_0', '../__data_0').
-?-data_out('__data_1', '../__data_1').
+%we can relieve memory by dumping the data into files 
+%and loading it later on. Files will have the corresponding id
+?-data_out('__data_0').
+?-data_out('__data_1').
 
-?-get_frame(signal(2, 44100, 1200000, ['__data_0', '__data_1']), 200, 5,
+?-data_in('__data_0'), data_in('__data_1'),
+get_frame(signal(2, 44100, 1200000, ['__data_0', '__data_1']), 200, 5,
 frame(Ch, Sr, L, [Ch1D, Ch2D])),
@@
 \centerline{\framebox{\includegraphics[width=\columnwidth]{workflow.png}}}
-\caption{Example of how to create a workflow: compare the tempo of a signal and a reverbered version of it}
+\caption{Example of how to create a workflow: extract the tempo of a signal and a reverberated version of it}
 \label{fig:workflow}
@@
 By hosting ``plugin libraries'' we can extend easily the SWI-DSP vocabulary (new plugins can be added without code modifications).
-These low level modules are not constrained by datatypes (every API deals with their own concepts) so they require a precise understanding of the API and the interface.
+These low level modules are not constrained by datatypes (every API deals with its own concepts) so they require a precise understanding of the API and the interface.
 
@@
 This module is splitted in two sub-modules: \textbf{swiaudiosource} and \textbf{swiaudiodata}.
-The former is a powerful module for decoding which interacts with C decoding libraries (mad, soundfile, fishsound, oggz and faad) to extract a \textit{Signal} term given and audio input file supported by the libraries.
-This process is triggered by a simple call to the predicate \textit{aspl\_decode(+InputFile, -Signal)} which recognizes the file extension and calls a specific library allowing a fast decoding process and even more important an independent term representing the encoded data in a standard way for any encoded format.
+The former is a module for decoding which interacts with C decoding libraries (mad, soundfile, fishsound, oggz and faad) to extract a \textit{Signal} term given and audio input file supported by the libraries.
+This process is triggered by a simple call to the predicate \textit{aspl\_decode(+InputFile, -Signal)} which recognises the file extension and calls a specific library allowing a fast decoding process and even more important an independent term representing the encoded data in a standard way for the supported formats: mp3, ogg, mp4, aac and wav.
 
@@
 The most important set of predicates is dedicated to framing.
-The original signal is decomposed in overlapping blocks of data outputting frame terms through the non-deterministic predicate \textit{framing(+Signal, +StepSize, +BlockSize, -Frame)}.
-This predicate calculates the number of possible frames using the length of the Signal and the StepSize and creates a list of natural indexes which are passed to the low-level framing predicate in a non-deterministic way using member/2.
 The rest of predicates are mainly oriented to extract information from a Signal or a Frame term.
@@
 This module defines a set of predicates and rules (specific API routines for the computations) for feature extraction.
-As we have seen in \S 2, the module is characterized for a clear input/output relationship.
-So we can declare a simple predicate to extract the tempo of an audio signal: tempo(+Input, -Tempo) being Input an instantation of the Signal datatype and Tempo a list of instantations of the Feature datatype.
+As we have seen in \S 2, the module is characterised for a clear input/output relationship.
+So we can declare a simple predicate to extract the tempo of an audio signal: tempo(+Input, -Tempo) being Input an instantiation of the Signal datatype and Tempo a list of instantiations of the Feature datatype.
 This predicate, that returns a tempo-related list of features given an input signal, is indeed triggering a specific computation tool, so we could select the specific tool by extending the predicate: \textit{tempo(+Input, +Tool, -Tempo)}.
@@
 The sub-module \textbf{swivamp} is a communication layer based on the SWI-Prolog foreign interface allowing us to query the plugin information, set up the plugin context for the transform, execute the plugin given a Signal term and retrieve the output features for the plugin according to the lifecycle specification.
-Plugins are wrapped in a similar fashion than the audio data so we can use a plugin instance representation (a plugin id pointing to a plugin isntance existing in memory), so that we can have concurrent plugins working at the same time.
-The previous tempo/3 predicate is declared as this set of rules:
+Plugins are wrapped in a similar fashion than the audio data so we can use a plugin instance representation (a plugin id pointing to a plugin instance existing in memory), so that we can have concurrent plugins working at the same time.
+The previous tempo/3 predicate is declared as this set of rules (defined with a procedural meaning):
 
 \medskip
-
 \noindent
-
 \begin{verbatim}
-
-tempo(Input, Tool, Tempo):-
+tempo(Input, Tool, [feature(Type, Timestamp, Feature)|...]):-
      Input = 'Signal'(Ch, Sr, L, PCMs),
      Tool = 'Vamp',
-     vamp_feature_of('tempo', Signal, WholeFeature).
+     vamp_feature_of(tempo, Input, Tempo).
+
 vamp_feature_of(Type, Signal, WholeFeature):-
@@
      flatten(Features, WholeFeature).
+
 vamp_outputs_for(Signal, PluginKey, Output, Feature):-
@@
      vamp_compute_feature(Signal, Step, Block, Outputs, Plugin, WholeFeature).
+
 vamp_compute_feature(Signal, Step, Block, Outputs, Plugin, Features):-
@@
      delete(RawFeatures, [], Features).
+
 vamp_process_signal(Signal, L, StepSize, Block, Out, Plugin, Features):-
@@
 \noindent
+
+The first rule selects the tool for the queried feature.
+The following one identifies automatically the specific plugin within the Vamp API in charge of extraction the ``tempo'' (and passes a term to unify the returned feature).
+The third predicate sets up the plugin for the processing (in this case with default parameters).
+The predicate vamp\_compute\_features acts as host retrieving and organising the extracted features which are indeed computed by the last predicate.
+This last predicate configures the framing and calls the swivamp interface that rapidly communicates and executes the plugin (the efficiency of the extraction is preserved).
+The features are returned as a list of terms represented using the feature datatype: feature(type, timestamp(start, duration), '\_\_data\_id').
 
@@
 
-So far we have presented how SWI-DSP provides a vocabulary of logic predicates and datatypes to perform audio analysis imitating a functional language, but this aproach is still inappropiate for such tasks.
-We have organized the modules in communication layers and task-oriented modules, but there is an important trade-off between understandability and specification.
-The communication layer even offering an application-independent access to the computation libraries still requires a high knowledge of such libraries besides a SWI-Prolog familiarization.
+So far we have presented how SWI-DSP provides a vocabulary of logic predicates and datatypes to perform audio analysis imitating a functional language, but this approach is still inappropriate for such tasks.
+We have organised the modules in communication layers and task-oriented modules, but there is an important trade-off between understandability and specification.
+The communication layer even offering an application-independent access to the computation libraries still requires a high knowledge of such libraries besides a SWI-Prolog familiarisation.
 However, these requirements are not expected in most part of the potential users.
@@
 This scenarios define the necessary patterns and datatypes (implemented as predicates and functors) to allow the user to actually perform high-level research preserving the necessary flexibility.
-The elemental scenarios are:
+The elementary scenarios are:
 
@@
 More complex scenarios can be created in order to achieve higher or broader semantics for research activity.
-This interactive logical framework can be interfaced, tranlasted or embedded into other environments by means of the wide range of SWI-Prolog tools as explained in the next section.
 
-\section{Data management and computation reasoning}\label{subsec:datamanage}
+This logic-based framework for DSP (Fig.
+\ref{fig:dspdrawing}) develops an interactive computational engine that exploits logic resources and their declarative nature (but still using procedural meaning).
+Such an engine can be embedded in other environments (by SWI-Prolog interfaces to other languages) developing graphical interfaces, visualisation applications or data management systems.
+In section \ref{sec:datamanage}, we introduce the elements to attain a declarative framework to reason about computations and processed data (which requires a management system).
+
+\begin{figure}
+\centerline{\framebox{\includegraphics[width=\columnwidth]{swidsp.png}}}
+\caption{Arquitecture of SWI-DSP}
+\label{fig:dspdrawing}
+\end{figure}
+
+\section{Data management and computation reasoning}\label{sec:datamanage}
 
@@
 We could say that such a system is indeed aimed at managing complex and large data, which is related through functional ``blocks'', and allowing reasoning about it.
-This goal requires the following achievements: an efficient storage of data objects and the relationships (predicates) linking them, mechanisms to control the concurrency, consistency and side-effects of the computations and a high-level interface to interact with them.
+This goal requires the following capabilities: an efficient storage of data objects and the relationships (predicates) linking them, mechanisms to control the concurrency, consistency and side-effects of the computations and a high-level interface to interact with them.
 
-We do not develop such manager within SWI-DSP as the database implementation actually depends on the needs of the host agent interacting with SWI-DSP.
-However, supposed an external database we can describe the architecture in detail.
-Data is stored in the database by ``tabling'' predicates \cite{aes2006}.
-Tabled predicates (we table predicates that comprise a big load of computations) can be seen as `` virtual tables'' of n-tuples.
-After tabling a predicate, the computation is not longer processed and the data is retrieved as in relational database query.
+We do not develop such manager within SWI-DSP, but we describe a Prolog-based framework that provide these services and to which SWI-DSP can provide computational services.
+Functional predicates can be externally stored in a database as tabled predicates.
+Those predicates can be seen as a \textit{relation} and each \textit{record} is formed of the terms of a particular instantiation of that predicate.
+By tabling functional predicates we can avoid re-computations of the same predicate treating it as another ``fact of the world''.
+
 This database is seen from a deductive model (facts and rules in prolog), so inferences over relational knowledge are also allowed.
-For this purpose, the data identifiers mechanism seems to be ideal as data ids are relationally linked (and binary data is dumped and loaded from external files relieving memory) and stored wrapping complex data as units of knowledge.
-The persistency of a tabled predicate is described as follows:
-
-\begin{verbatim}
-?-table(decode('../myfile', signal(2, 44100, 1200000, ['__data_345', '__data_346']))).
-
-The underlying process in pseudo-code:
-
-table(predicate(File, signal(Ch, Sr, Length, Data)):-
-	to_database('predicate', File, Ch, Sr, Length, Data).
-to_database('predicate', File, Ch, Sr, Length, (DataID|List)):-
-	specific_implementation('predicate', File, Ch, Sr, Length, DataID),
-	data_out(DataID), 
-	%DataID is the name of the file in a predefined path of binary data
-	clean_data_from_memory(DataID),
-	to_database('predicate', File, Ch, Sr, Length, List).	
-\end{verbatim}
+For this purpose, the data identifiers mechanism seems to be ideal as data ids are relationally linked and stored wrapping complex data as units of knowledge.
+The data objects referred by the ids are made persistent between sessions by dumping the data to an external file which can be retrieved through the data identifier (directly or through a mapping).
+The data object is loaded at runtime any time we use a query involving a fact referring to its identifier.
 
 This data management framework described as in \cite{aes2006} can be extended by handling transaction logic constructs.
-Transaction logic provides the formalism to combine, with a declarative meaning, functional predicates and data transactions.
+Transaction logic provides the formalism to combine, with a declarative meaning, functional predicates and actions that read or change the ``state of the world'', such as database updates, creation of files on disk or interaction with a user.
 
-\subsection{Transaction Logic}\label{sec:ctl}
+\subsection{Transaction Logic}\label{subsec:ctl}
 
 Transaction Logic \cite{ctl}  unifies in a declarative formalism the control of an underlying database (SWI-Prolog and other prolog implementations just rely on assert and retract predicates for data updating, being very limited).
-Transaction Logic allows a declarative definition of execution paths and data updates by providing new connectives to express complex data transactions: parallel and serial conjunctions, execution in isolation and executability checking.
+Transaction Logic allows a declarative definition of execution paths and database updates by providing new connectives to express complex data transactions: parallel and serial conjunctions, execution in isolation and executability checking.
 Rules define possible execution paths for a goal with a higher-level meaning than just declared procedures.
@@
 
-SWI-DSP combined with a data manager based on Concurrent Transaction Logic (CTR), which also controls concurrent processes, develops a multimedia knowledge system to drive research with semantics and a logic formalism exploiting reasoning and declarative definition of computations and providing an interactive source of knowledge.
+SWI-DSP combined with a database manager based on Concurrent Transaction Logic (CTR), which also controls concurrent processes, develops a multimedia knowledge system to drive research with semantics and a logic formalism exploiting reasoning and declarative definition of computations and providing an interactive source of knowledge.
+
+\subsection{Implementation}
+
+\begin{figure}
+\centerline{\framebox{\includegraphics[width=\columnwidth]{reasoner.png}}}
+\caption{Collaboration diagram between SWI-DSP, the Matlab interface and the reasoner implementations}
+\label{fig:reasoners}
+\end{figure}
 

### 2008-06-11 trunk/vamp-n3/generator/rdf_vamp.pl
a rdf-vamp plugins description generator

@@ -1,0 +1,54 @@
+/**
+	Bit module to describe vamp plugins as rdf
+	based on Chris Sutton's template_generator
+	David Pastor 2008, c4dm
+	TODO:
+		Pair plugin-library automatically (maybe from swivamp or just analysing the name...)
+		Specifies the type of plugin if it is not ok with the default one
+		Dump into a file rather than on the screen
+		
+*/
+
+:-module(rdf_vamp, [available_libraries/1
+			,	rdf_description/2
+			,	rdf_description/3
+			]).
+
+:- style_check(-discontiguous).
+:- load_foreign_library(template_generator).
+
+:- use_module('../../swivamp/vamp').
+
+/**Define your namespace for vamp plugins**/
+
+pluginlib(qm).
+pluginlib(examples).
+namespace(qm, 'http://purl.org/ontology/vamp/qm-plugins/').
+namespace(examples, 'http://purl.org/ontology/vamp/examples/').
+
+/**
+	Returns the available libraries on the system. This is hardcoded, so requires manual updating
+
+	available_libraries(-LibID)
+*/
+available_libraries(X):-
+	pluginlib(X).
+
+/**
+	RDF descriptions:
+
+	rdf_description(?PluginKey, ?PluginLibrary, +YourURI)
+
+	You should introduce your plugin library your plugin key and your URI. You can get some information from swivamp :)
+	eg. vamp_plugin_system(?PluginKey).
+
+**/
+rdf_description(PluginKey, PluginLibrary):-
+	rdf_description(PluginKey, PluginLibrary, 'http://www.vamp-plugins.org/').
+
+rdf_description(PluginKey, PluginLibrary, YourURI):-
+	vamp_plugin_system(PluginKey),
+	pluginlib(PluginLibrary), namespace(PluginLibrary, Namespace),
+	create_rdf_description(PluginKey, Namespace, YourURI).
+	
+	

### 2008-06-11 trunk/vamp-n3/generator/template_generator.cpp

@@ -1,0 +1,290 @@
+/* -*- c-basic-offset: 4 indent-tabs-mode: nil -*-  vi:set ts=8 sts=4 sw=4: */
+
+/** RDF description generator for Vamp Plugins matching the Vamp Ontology **/
+
+#ifndef _FLI_H_INCLUDED
+#include <SWI-cpp.h>
+#endif
+
+#include "vamp-sdk/PluginHostAdapter.h"
+#include "vamp-sdk/hostext/PluginChannelAdapter.h"
+#include "vamp-sdk/hostext/PluginInputDomainAdapter.h"
+#include "vamp-sdk/hostext/PluginLoader.h"
+#include "vamp/vamp.h"
+
+#include <iostream>
+#include <fstream>
+#include <sstream>
+#include <sndfile.h>
+
+#include <cmath>
+
+using std::cout;
+using std::cin;
+using std::cerr;
+using std::getline;
+using std::endl;
+using std::string;
+using std::vector;
+using std::ofstream;
+using std::ios;
+
+using Vamp::HostExt::PluginLoader;
+using Vamp::Plugin;
+
+/*
+  
+   usage:
+
+   template-generator vamp:aubioonset:onsets
+
+*/
+
+string programURI = "http://www.vamp-plugins.org/doap.rdf#template-generator";
+
+void usage()
+{
+    cerr << "usage: template-generator [PLUGIN_BASE_URI YOUR_URI] vamp:soname:plugin[:output]" << endl;
+    exit(2);
+}
+
+template <class T>
+inline string to_string (const T& t)
+{
+    std::stringstream ss;
+    ss << t;
+    return ss.str();
+}
+
+string describe_namespaces(Plugin* plugin, string pluginBundleBaseURI)
+{
+	string res=\
+"@prefix rdfs:     <http://www.w3.org/2000/01/rdf-schema#> .\n\
+@prefix xsd:      <http://www.w3.org/2001/XMLSchema#> .\n\
+@prefix vamp:     <http://www.purl.org/ontology/vamp/> .\n\
+@prefix vampex:   <http://www.purl.org/ontology/vamp/examples/> .\n\
+@prefix plugbase: <"+pluginBundleBaseURI+"> .\n\
+@prefix owl:      <http://www.w3.org/2002/07/owl#> .\n\
+@prefix dc:       <http://purl.org/dc/elements/1.1/> .\n\
[…]

### 2008-06-11 vamp-n3/generator/rdf_vamp.pl
Automatic writing of plugin descriptions.

@@ -12,19 +12,23 @@
 :-module(rdf_vamp, [available_libraries/1
 			,	rdf_description/2
 			,	rdf_description/3
+			,	rdf_description/1
 			]).
 
 :- style_check(-discontiguous).
 :- load_foreign_library(template_generator).
 
 :- use_module('../../swivamp/vamp').
+%:- consult('vamp_af').
 
-/**Define your namespace for vamp plugins**/
+/**Define your namespace for vamp plugins. This should be automated**/
 
 pluginlib(qm).
 pluginlib(examples).
+pluginlib('').
 namespace(qm, 'http://purl.org/ontology/vamp/qm-plugins/').
 namespace(examples, 'http://purl.org/ontology/vamp/examples/').
+namespaces('', 'http://purl.org/ontology/vamp/plugins').
 
 /**
 	Returns the available libraries on the system. This is hardcoded, so requires manual updating
@@ -45,6 +49,11 @@ available_libraries(X):-
 **/
 rdf_description(PluginKey, PluginLibrary):-
 	rdf_description(PluginKey, PluginLibrary, 'http://www.vamp-plugins.org/').
+
+rdf_description(PluginKey):-
+	vamp_plugin_system(PluginKey),
+	get_library(PluginKey, Lib),
+	rdf_description(PluginKey, Lib, 'http://www.vamp-plugins.org/').
 
 rdf_description(PluginKey, PluginLibrary, YourURI):-
 	vamp_plugin_system(PluginKey),

### 2008-06-11 trunk/vamp-n3/generator/template_generator.cpp
Automatic writing of plugin descriptions.

@@ -146,22 +146,22 @@ string describe_output(Plugin::OutputDescriptor o)
         if (o.sampleType == Plugin::OutputDescriptor::VariableSampleRate)
 	{
 
-                	res=\
-                        "thisplug:output_"+o.identifier+" a  vamp:SparseOutput ;\n\
-                          vamp:identifier     \""+o.identifier+"\" ;\n\
-                          dc:title            \""+o.name+"\" ;\n\
-                          dc:description      \""+o.description+"\"  ;\n\
-                          vamp:fixed_bin_count \""+(o.hasFixedBinCount == 1 ? "true" : "false")+"\" ;\n\
-                          vamp:is_quantized     \""+(o.isQuantized == 1 ? "true" : "false")+"\"  ;\n\
-                          vamp:unit             \""+(o.unit)+"\" ;\n";
+                res=\
+"thisplug:output_"+o.identifier+" a  vamp:SparseOutput ;\n\
+    vamp:identifier       \""+o.identifier+"\" ;\n\
+    dc:title              \""+o.name+"\" ;\n\
+    dc:description        \""+o.description+"\"  ;\n\
+    vamp:fixed_bin_count  \""+(o.hasFixedBinCount == 1 ? "true" : "false")+"\" ;\n\
+    vamp:is_quantized     \""+(o.isQuantized == 1 ? "true" : "false")+"\"  ;\n\
+    vamp:unit             \""+(o.unit)+"\" ;\n";
                           
                         
 
                 // FIXME ? Bin names may vary based on plugin setup, so including them here might be misleading...
                 if (o.hasFixedBinCount)
                 {
-		res+="    vamp:bin_count      "+to_string(o.binCount)+" ;\n";
-		res+="    vamp:bin_names      (";
+		res+="    vamp:bin_count          "+to_string(o.binCount)+" ;\n";
+		res+="    vamp:bin_names          (";
 
 		unsigned int i;
 		for (i=0; i+1 < o.binNames.size(); i++)
@@ -173,10 +173,10 @@ string describe_output(Plugin::OutputDescriptor o)
         
                 if (o.isQuantized)
                 {
-                res+="  vamp:quantize_step     "+to_string(o.quantizeStep)+"  ;\n";
+                res+="   vamp:quantize_step        "+to_string(o.quantizeStep)+"  ;\n";
                 }
 
[…]

### 2008-06-15 paper/logicdsp_iclp.tex
submitted version of the iclp paper added

--- 2008-06-09 paper/paper.tex
@@
 
-So far we have presented how SWI-DSP provides a vocabulary of logic predicates and datatypes to perform audio analysis imitating a functional language, but this approach is still inappropriate for such tasks.
-We have organised the modules in communication layers and task-oriented modules, but there is an important trade-off between understandability and specification.
-The communication layer even offering an application-independent access to the computation libraries still requires a high knowledge of such libraries besides a SWI-Prolog familiarisation.
-However, these requirements are not expected in most part of the potential users.
-The datatypes are not the only necessary semantic resource to provide an actual workbench for DSP.
-We require a front-end that reflects the experimental domain providing an interactive ``laboratory''.
+So far we have described how SWI-DSP provides a vocabulary of logic predicates and datatypes for performing audio analysis, but this aproach is still inappropiate for some tasks.
+We have organised the modules in communication layers and task-oriented modules, but there is a trade-off between understandability and rich functionality.
+Use of the computational facilities provided by the various modules requires quite detailed knowledge of their capabilities as well as familiarity with Prolog.
+We do not expect this type of expertise from most of our potential users.
+Hence, we require a front-end that is closer to the application domain, providing an interactive ``laboratory'' at a level of abstraction closer to that expected by the signal processing practitioner.
 
@@
 This logic-based framework for DSP (Fig.
-\ref{fig:dspdrawing}) develops an interactive computational engine that exploits logic resources and their declarative nature (but still using procedural meaning).
+\ref{fig:dspdrawing}) develops an interactive computational engine that exploits logic resources and their declarative nature (but also relies on procedural declarations).
 Such an engine can be embedded in other environments (by SWI-Prolog interfaces to other languages) developing graphical interfaces, visualisation applications or data management systems.
-In section \ref{sec:datamanage}, we introduce the elements to attain a declarative framework to reason about computations and processed data (which requires a management system).
+Further knowledge can be achieved by combining different sources of data under a single declarative framework as explained in section \ref{sec:datamanage}.
 
@@
 \centerline{\framebox{\includegraphics[width=\columnwidth]{swidsp.png}}}
-\caption{Arquitecture of SWI-DSP}
+\caption{Architecture of SWI-DSP}
 \label{fig:dspdrawing}
@@
 
-We still need to address, probably, the most important issue to accomplish the proposed architecture.
-We could say that such a system is indeed aimed at managing complex and large data, which is related through functional ``blocks'', and allowing reasoning about it.
-This goal requires the following capabilities: an efficient storage of data objects and the relationships (predicates) linking them, mechanisms to control the concurrency, consistency and side-effects of the computations and a high-level interface to interact with them.
+We still need to address what is probably the most important issue in constructing the proposed ``laboratory workbench''.
+A primary aim of such a system is to manage complex relationships between many items of data, some of which may be large and temporally structured (e.g.
+audio signals).
+Some of the relationships are functional, in the sense that some items are computed from others using signal processing functions provided through the foreign language interfaces described above.
+Some relationships are not functional, eg, several human experts might label a set of signals by musical genre, resulting in many possible labellings for each signal.
+This type of knowledge management requires mechanisms for persistent storage of data objects and the procedures that were used to compute them, as well as effective management of a local database, side-effects of computations, information from other sources such as manual annotations or external databases, and a high-level interface for specifying and scheduling potentially complex workflows involving many computations and combining data from different sources.
 
-We do not develop such manager within SWI-DSP, but we describe a Prolog-based framework that provide these services and to which SWI-DSP can provide computational services.
-Functional predicates can be externally stored in a database as tabled predicates.
-Those predicates can be seen as a \textit{relation} and each \textit{record} is formed of the terms of a particular instantiation of that predicate.
-By tabling functional predicates we can avoid re-computations of the same predicate treating it as another ``fact of the world''.
+We do not develop such manager within SWI-DSP, but we are developing a Prolog-based system that does address most of these concerns \cite{aes2006}.
+Information pertaining to many types of object, including signals, computed results, musical scores, composers etc.
+is represented uniformly using a system of predicates, some of which are backed by a relational database.
+When time-consuming computations are involved in satisfying a particular predicate, the results can be persistently stored along with a record of the bindings for that predicate.
+If, at a later time, the same computation is requested using the same input parameters, the result is retrieved from the database instead of being recomputed.
+It is also possible to look up the input parameters that resulted in a given value being computed, and
+to list all previously recorded computations of a given predicate.
 
-This database is seen from a deductive model (facts and rules in prolog), so inferences over relational knowledge are also allowed.
-For this purpose, the data identifiers mechanism seems to be ideal as data ids are relationally linked and stored wrapping complex data as units of knowledge.
-The data objects referred by the ids are made persistent between sessions by dumping the data to an external file which can be retrieved through the data identifier (directly or through a mapping).
-The data object is loaded at runtime any time we use a query involving a fact referring to its identifier.
+This is in some ways similar to the ``tabling'' mechanism provided by some logic languages, notably XSB \cite{xsb}.
+The previous implementation of this knowledge management system included a module that allowed the use of MATLAB to implement computationally intensive functions.
+SWI-DSP can now provide a similar capability to use VAMP plugins, LADSPA plugins, and WEKA code.
+The data identifier mechanism is compatible with the overall system as the identifiers can be stored in a database and associated with files containing the actual binary data.
 
-This data management framework described as in \cite{aes2006} can be extended by handling transaction logic constructs.
-Transaction logic provides the formalism to combine, with a declarative meaning, functional predicates and actions that read or change the ``state of the world'', such as database updates, creation of files on disk or interaction with a user.
-
-\subsection{Transaction Logic}\label{subsec:ctl}
-
-Transaction Logic \cite{ctl}  unifies in a declarative formalism the control of an underlying database (SWI-Prolog and other prolog implementations just rely on assert and retract predicates for data updating, being very limited).
-Transaction Logic allows a declarative definition of execution paths and database updates by providing new connectives to express complex data transactions: parallel and serial conjunctions, execution in isolation and executability checking.
-Rules define possible execution paths for a goal with a higher-level meaning than just declared procedures.
-Thus, we are certainly dealing with actions (and their effects) instead of sequences of predicates.
-
-We can manage the side effects of computation on an underlying database in a consistent way, reason about the actions and their effects, control the execution of different processes and compare different paths for the same goal.
+The other half of the problem, that of workflow management, is addressed using a implementation of \emph{transaction logic} \cite{ctl}.
+Transaction logic allows the specification, in a declarative way, of sequential or parallel compositions of actions, possibly involving database update, side-effects, and interaction with a user or other external device.
+The primitives offered by the transaction logic formalise the notion of calls with side-effects and
+provide an alternative to some of the more procedural styles of Prolog usage, such as failure driven loops, making it easier to specify complex workflows without sacrificing a clear declarative semantics.
 Transaction logic applied to complex data/computational environments has been successfully applied to develop ``virtual laboratory workbenches'' \cite{virtual}.
 
-SWI-DSP combined with a database manager based on Concurrent Transaction Logic (CTR), which also controls concurrent processes, develops a multimedia knowledge system to drive research with semantics and a logic formalism exploiting reasoning and declarative definition of computations and providing an interactive source of knowledge.
+SWI-DSP combined with a data manager based on Concurrent Transaction Logic (CTR), which also controls concurrent processes, develops a multimedia knowledge system to drive research with semantics and a logic formalism exploiting reasoning and declarative definition of computations and providing an interactive source of knowledge.
 
-\subsection{Implementation}
+\subsection{Towards a distributed space of knowledge}
+
+Finally, we address a main problem pointed out at the beginning of the exposition.
+Local sources of data managed by closed systems are not the right approach in order to create a scalable source of shared knowledge.
+We find in the semantic web technologies the support to open the depicted system and contribute to a distributed space of knowledge as described in \cite{distributed}.
+A suitable machine-processable format for data, Resource Description Framework (RDF), and a language to express knowledge domains in such a format, Ontology Web Language (OWL), allow us to link data resources by relationships and identify them all with Uniform Resource Identifiers (URIs) belonging to a shareable namespace.
+
+Notation3 (N3) \cite{n3} is a syntax for RDF extended to represent rules. Therefore, Notation3 can be used to reproduce a deductive model: facts as triples (Subject, Predicate, Object) and rules.
+
+The ongoing work so-called Henry develops a similar transaction logic-based reasoner but N3-driven.
+Henry translates Prolog predicates into RDF triples and reasons about N3 rules by N3/Prolog parser and entailment modules.
+Functional predicates provided by SWI-DSP are interpreted by Henry as builtins that trigger computations.
+Thus, Henry is capable of integrating and reasoning about heterogeneous data belonging to different sources and knowledge domains and is enhanced by the processing services provided by SWI-DSP.
 
@@
 \centerline{\framebox{\includegraphics[width=\columnwidth]{reasoner.png}}}
-\caption{Collaboration diagram between SWI-DSP, the Matlab interface and the reasoner implementations}
+\caption{Collaboration diagram between SWI-DSP, the Matlab interface and the high-level reasoning implementations}
 \label{fig:reasoners}
@@
 
-%Samer's KM and Henry...
-
-% FIXME - Hmm... Not good in the current state.
-%\subsection{Henry}\label{subsec:henry}
-%We present a related project so-called Henry which develops a real CTR-based agent aimed at creating a distributed multimedia knowledge workspace built on semantic web technologies.
-The multimedia community needs a standard machine-readable format with proper semantics to manage complex data involved in any sub-domain, so data can be exchanged and shared by different systems enriching the common knowledge.
-
-%The Resource Description Framework (RDF) developed as the standard format for new (data-oriented) web technologies expresses linked data by means of triples in the form of (Subject, Predicate, Object).
-This standard can be implemented with different syntaxes.
-We prefer Notation3 \cite{n3} whose syntax allow us to write RDF statements in a human-readable way and also to write ``rules'' as named graphs.
-The Ontology Web Language (OWL) provides the framework to express the knowledge of a particular domain in RDF through concepts and relationships.
-Thus, descripting semantic models related to audio analysis are not restricted to a local implementation but can be shared and processed by different systems.
-These elements form a suitable framework to represent and manage data and algorithms in a ``web of (complex) data''.
-
-%Henry develops an N3/Prolog parser and reasoner based on CTL. (Text for Yves). Talk about results and performance...
-%(talk about data visualisation?)
-
 \section{Conclusion and Future work}\label{sec:conclusion}
 
-(may have to reword this after all the changes)
-We have presented a powerful reasoner for multimedia analysis implemented as a logic-based framework which allow us to convert low-level computations into shareable and distributed knowledge.
-The system design ``breaks apart'' traditional applications architecture to introduce logic and reasoning machinery on top of the computation engines providing a more intelligent, capable and understandable DSP reasoner.
-This reasoner can be extended and enriched by the open source APIs working on the foundations of the system and controlled by ''Prolog-speaking`` agents.
-SWI-DSP is fairly well positioned for the new web technologies and actually it finds in them the perfect semantic and storing support to overwhelm traditional and closed DSP systems.
+We have presented a logic-based engine that provides digital signal processing services under an interactive and declarative framework which is able to make data inferences.
+Such an engine solves many of the problems presented by current approaches to develop an integrated platform for signal processing and multimedia analysis in general.
+This system is aimed at providing services to high-level reasoners capable of integrating and reasoning about heterogeneous data available in different data systems with the same declarative nature.
 
-Future work and development is oriented to:
+We expect of SWI-DSP to be a useful workbench for DSP research and also a platform for development enriching the community resources.
+The proposed system and future extensions are expected to become a powerful workspace to integrate and analyse heterogeneous multimedia data providing enhanced services to a wide range of users and contributing to extend a common source of knowledge.
 
-\begin{itemize}
- \item Testing and comparison of performance with other rival systems
- \item Extending the SWI-DSP vocabulary and capabilities
-\end{itemize}
+Near future work and development on SWI-DSP are oriented to extend its vocabulary and capabilities.
+First, by interfacing more open source APIs to cover the implementation of more DSP tasks (not necessarily restricted to the audio domain).
+Also, by supporting symbolic data (MIDI) input which is a really important source of audio knowledge and providing analysis modules for such an input.
 
-SWI-DSP vocabulary and capabilities can be extended in several ways.
-First, by interfacing more open source APIs to cover the implementation of any DSP task (not necessarily restricted to audio domain).
-Secondly, by supporting symbolic data (MIDI) input which is a really important source of audio knowledge and providing analysis modules for such an input.
-Finally, other logic paradigms can contribute to enhance SWI-DSP (e.g.
-there are studies that use Inductive Logic Programming systems to extract harmonic rules or rhythm patterns given certain input descriptors [ref]).
-By providing semantics (OWL ontologies clearly point to this goal) to interconnect properly each logical module we can build a hierarchy of modules which help us to extract and share different levels of knowledge from digital signals.
-Such a system could be used either to extract low level cepstral coefficients from an input signal (mathematics level) or to obtain patterns of music (musicology level).
+In general, any source of knowledge, processing tool or logic paradigm which allow us to integrate and infer new knowledge is a target in order to reach higher-levels of abstraction to understand the nature and effects of the signals.
 
@@
 
-\bibitem{tlprog} Bonner, A.J., Kifer, M.: Transaction Logic Programming. In: 10th International Conference on Logic Programming. Budapest (1993)
-
 \bibitem{virtual} Bonner, A.J., Shrufi, A., Rozen, S.: LabFlow-1: a Database Benchmark for High-Throughput Workflow Management.
@@
 \bibitem{aes2006} Abdallah, S., Raimond, Y., Sandler, M.: An ontology-based approach to information management for music analysis systems. In: 120th Audio Engineering Society. Paris (2006).
+
+\bibitem{xsb} Sagonas, K., Swift, T., Warren, D. S.: {XSB} as an efficient deductive database engine. In: ACM SIGMOD international conference on management of data, pp. 442-453. New York (1994).
 
@@
 
-\bibitem{kmrdf} Source code project of KM-RDF. \url{http://code.google.com/p/km-rdf/w/list}
+\bibitem{henry} Henry. \url{http://dbtune.org/henry/}
 

### 2008-06-15 henry/skolemize.pl
Some second-order magic moving away from existential variables

@@ -1,0 +1,32 @@
+:- module(skolemize,[skolemize/3]).
+
+/**
+ * Some unit tests
+ */
+:- begin_tests(skolemize).
+test(skolemize) :-
+	skolemize(
+	[rdf(A,has,bnode(Nose)),rdf(bnode(Nose),colour,bnode(Colour)),rdf(bnode(Colour),label,pink)],
+	[rdf(A,isa,man)],
+	[rdf(A, has, individual(Nose, [A])), rdf(individual(Nose, [A]), colour, bnode(Colour)), rdf(individual(Colour, [A]), label, pink)]).
+:- end_tests(skolemize).
+/**
+ * End of unit tests
+ */
+
+
+skolemize([],_,[]).
+skolemize([rdf(S,P,O)|T1],Body,[rdf(individual(B,Vars),P,O)|T2]) :-
+	S=@=bnode(B),S=bnode(B),
+	free_variables(Body,Vars),!,
+	skolemize(T1,Body,T2).
+skolemize([rdf(S,P,O)|T1],Body,[rdf(S,P,individual(B,Vars))|T2]) :-
+        O=@=bnode(B),O=bnode(B),
+	free_variables(Body,Vars),!,
+        skolemize(T1,Body,T2).
+
+
+
+
+
+

### 2008-06-15 trunk/paper/logicdsp_iclp.tex
Final version

--- 2008-06-15 paper/logicdsp_iclp.tex
@@
 
-\subsection{Towards a distributed space of knowledge}
-
-Finally, we address a main problem pointed out at the beginning of the exposition.
-Local sources of data managed by closed systems are not the right approach in order to create a scalable source of shared knowledge.
-We find in the semantic web technologies the support to open the depicted system and contribute to a distributed space of knowledge as described in \cite{distributed}.
-A suitable machine-processable format for data, Resource Description Framework (RDF), and a language to express knowledge domains in such a format, Ontology Web Language (OWL), allow us to link data resources by relationships and identify them all with Uniform Resource Identifiers (URIs) belonging to a shareable namespace.
-
-Notation3 (N3) \cite{n3} is a syntax for RDF extended to represent rules. Therefore, Notation3 can be used to reproduce a deductive model: facts as triples (Subject, Predicate, Object) and rules.
-
-The ongoing work so-called Henry develops a similar transaction logic-based reasoner but N3-driven.
-Henry translates Prolog predicates into RDF triples and reasons about N3 rules by N3/Prolog parser and entailment modules.
-Functional predicates provided by SWI-DSP are interpreted by Henry as builtins that trigger computations.
-Thus, Henry is capable of integrating and reasoning about heterogeneous data belonging to different sources and knowledge domains and is enhanced by the processing services provided by SWI-DSP.
+Currently, we have another implementation of such a manager oriented to interact with the \textit{Semantic Web}.
+This project, so called Henry\footnote{Source code project at \url{http://code.google.com/p/km-rdf/}.}, aggregates processing workflows and structured data available on the Semantic Web and publishes back the results it is able to derive.
+It can interpret queries and express data through a Notation3 \cite{n3} parser/entailment module.
+Notation3 is a syntax for the Semantic Web format which provides a way to reproduce a deductive model based on triples (Subject, Predicate, Object) and rules.
+SWI-DSP provides processing services to Henry.
+There is a working instance at \url{http://dbtune.org/henry/}.
 
@@
 \begin{thebibliography}{4}
-
-\bibitem{distributed} Raimond, Y., Sutton, C., Sandler, M.: A distributed data space for music-related information.
-In: Workshop on the Many Faces of Multimedia Semantic.
-ACM (Multimedia), Augsburg (2007).
 

### 2008-06-16 henry/entailment.pl
Refactoring Henry's code

@@ -1,0 +1,50 @@
+:- module(entailment,[rdf/3]).
+
+:- use_module(rdf_e).
+
+:- begin_tests(entailment).
+test(list_construction,[]) :-
+	rdf_l(A,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',a),
+	rdf_l(A,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest',B),
+	rdf_l(B,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',b),
+	rdf_l(B,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest',C),
+	rdf_l(C,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',c),
+	rdf_l(C,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest','http://www.w3.org/1999/02/22-rdf-syntax-ns#nil'),!,
+	A = [a,b,c],
+	B = [b,c],
+	C = [c].
+test(list_in1,[]) :-
+	rdf_l(A,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',a),
+        rdf_l(A,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest',B),
+        rdf_l(B,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',b),
+        rdf_l(B,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest',C),
+        rdf_l(C,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',c),
+        rdf_l(C,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest','http://www.w3.org/1999/02/22-rdf-syntax-ns#nil'),
+	rdf_l(D,'http://www.w3.org/2000/10/swap/list#in',A),!,
+	memberchk(D,[a,b,c]).
+test(list_in2,[blocked('The first member/2 creates an infinity of lists => infinite loop. This test should succeed as list_in1.')]) :-
+	rdf_l(D,'http://www.w3.org/2000/10/swap/list#in',A),
+	rdf_l(A,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',a),
+        rdf_l(A,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest',B),
[…]

### 2008-06-16 henry/n3_to_prolog.pl

@@ -1,0 +1,110 @@
+:- module(n3_to_prolog,[n3_prolog_clause/1,compile_all/0]).
+
+:- use_module(skolemize).
+:- use_module(utils).
+:- consult(namespaces).
+
+:- use_module(library('semweb/rdf_db'),
+              [ rdf_global_id/2,
+                rdf_reachable/3,
+                rdf_has/3,
+                rdf_bnode/1,
+                rdf_is_bnode/1,
+                rdf_subject/1,
+                rdf_equal/2
+              ]).
+:- use_module(library('semweb/rdfs'),
+              [ rdfs_list_to_prolog_list/2 ]).
+
+
+/**
+ * Some unit tests
+ */
+
+/**
+ *
+ */
+
+
+compile_all :-
+	forall(n3_prolog_clause(Clause),(format(user_error,'~w\n',[Clause]),assert(entailment:Clause))).
+
+n3_prolog_clause(':-'(H,B)) :-
+	implies(BodyGraph,HeadGraph),
+	n3_pl(HeadGraph,PLH,B1),
+	n3_pl(BodyGraph,PLB,B2),
+	skolemize(PLH,PLB,Skolem),
+	append(B1,B2,Bindings),
+	merge_bindings(Bindings),
+	member(H,Skolem),
+	list_to_conj(PLB,B).
+	
+n3_pl(Head,PredList,Bindings) :-
+        findall(rdf_e(S,P,O),rdf_db:rdf(S,P,O,Head),Triples),
+        n3_pl2(Triples,PredList,Bindings).
+n3_pl2([],[],[]).
+n3_pl2([rdf_e(_,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',_)|T],T2,B) :-
+        !,n3_pl2(T,T2,B).
+n3_pl2([rdf_e(_,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest',_)|T],T2,B) :-
+        !,n3_pl2(T,T2,B).
+n3_pl2([rdf_e(_,'http://www.w3.org/1999/02/22-rdf-syntax-ns#type','http://www.w3.org/1999/02/22-rdf-syntax-ns#List')|T],T2,B) :-
+        !,n3_pl2(T,T2,B).
+n3_pl2([rdf_e(S,P,O)|T],[rdf_e(SS,PP,OO)|T2],Bindings) :-
+        convert_n(S,SS,B1),
+        convert_n(P,PP,B2),
[…]

### 2008-06-16 trunk/henry/henry.pl
Refactoring Henry's code

@@ -1,0 +1,7 @@
+:- module(henry,[]).
+
+:- [n3_load].
+:- [n3_to_prolog].
+:- [entailment].
+
+

### 2008-06-16 trunk/henry/n3_load.pl

@@ -1,0 +1,11 @@
+:- module(n3_load,[n3_load/1]).
+
+:- use_module(n3_dcg).
+
+n3_load(File) :-
+        tokenise(File,Tokens),
+        absolute_file_name(File,Absolute),
+        format(atom(Base),'file://~w',[Absolute]), % That's probably not right
+        phrase(n3_dcg:document(Base,Doc),Tokens),
+        forall(member(rdf(A,B,C,D),Doc),rdf_db:rdf_assert(A,B,C,D)).
+

### 2008-06-16 trunk/henry/rdf_e.pl

@@ -1,0 +1,56 @@
+:- module(rdf_e,[rdf_e/3,rdf_b/3]).
+
+:- use_module(library('semweb/rdf_db'),
+              [ rdf_global_id/2,
+                rdf_reachable/3,
+                rdf_has/3,
+                rdf_bnode/1,
+                rdf_is_bnode/1,
+                rdf_subject/1,
+                rdf_equal/2
+              ]).
+
+
+:- dynamic rdf_e/3.
+
+
+:- begin_tests(rdf_e).
+test(list_to_query) :-
+	rdf_query(N,[a,b,c],(rdf(N, 'http://www.w3.org/1999/02/22-rdf-syntax-ns#first', a), rdf(N, 'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest', G498), rdf(G498, 'http://www.w3.org/1999/02/22-rdf-syntax-ns#first', b), rdf(G498, 'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest', G512), rdf(G512, 'http://www.w3.org/1999/02/22-rdf-syntax-ns#first', c), rdf(G512, 'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest', 'http://www.w3.org/1999/02/22-rdf-syntax-ns#nil'), true)),!.
+test(query_list) :-
+	rdf_db:rdf_assert(l1,isa,list),
+	rdf_db:rdf_assert(l1,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',a),
+	rdf_db:rdf_assert(l1,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest',l2),
+	rdf_db:rdf_assert(l2,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',b),
+	rdf_db:rdf_assert(l2,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest','http://www.w3.org/1999/02/22-rdf-syntax-ns#nil'),!,
+	rdf_e([a,b],isa,list),!,
+	rdf_db:rdf_retractall(l1,isa,list),
+	rdf_db:rdf_retractall(l1,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',a),
[…]

### 2008-06-16 trunk/henry/skolemize.pl

@@ -6,9 +6,9 @@
 :- begin_tests(skolemize).
 test(skolemize) :-
 	skolemize(
-	[rdf(A,has,bnode(Nose)),rdf(bnode(Nose),colour,bnode(Colour)),rdf(bnode(Colour),label,pink)],
-	[rdf(A,isa,man)],
-	[rdf(A, has, individual(Nose, [A])), rdf(individual(Nose, [A]), colour, bnode(Colour)), rdf(individual(Colour, [A]), label, pink)]).
+	[rdf_e(A,has,bnode(Nose)),rdf_e(bnode(Nose),colour,bnode(Colour)),rdf_e(bnode(Colour),label,pink)],
+	[rdf_e(A,isa,man)],
+	[rdf_e(A, has, individual(Nose, [A])), rdf_e(individual(Nose, [A]), colour, bnode(Colour)), rdf_e(individual(Colour, [A]), label, pink)]).
 :- end_tests(skolemize).
 /**
  * End of unit tests
@@ -16,16 +16,16 @@ test(skolemize) :-
 
 
 skolemize([],_,[]).
-skolemize([rdf(S,P,O)|T1],Body,[rdf(individual(B,Vars),P,O)|T2]) :-
+skolemize([rdf_e(S,P,O)|T1],Body,[rdf_e(individual(B,Vars),P,O)|T2]) :-
 	S=@=bnode(B),S=bnode(B),
 	free_variables(Body,Vars),!,
 	skolemize(T1,Body,T2).
-skolemize([rdf(S,P,O)|T1],Body,[rdf(S,P,individual(B,Vars))|T2]) :-
+skolemize([rdf_e(S,P,O)|T1],Body,[rdf_e(S,P,individual(B,Vars))|T2]) :-
         O=@=bnode(B),O=bnode(B),
 	free_variables(Body,Vars),!,
         skolemize(T1,Body,T2).
-
-
+skolemize([H|T1],Body,[H|T2]) :-
+	skolemize(T1,Body,T2).
 
 
 

### 2008-06-16 trunk/henry/utils.pl

@@ -1,0 +1,8 @@
+:- module(utils,[list_to_conj/2]).
+
+
+list_to_conj([H],H) :- !.
+list_to_conj([H,T],(H,T)) :- !.
+list_to_conj([H|T],(H,T2)) :-
+        list_to_conj(T,T2).
+

### 2008-06-16 henry/entailment.pl
Handling list transforms (+ associated unit tests)

@@ -35,11 +35,10 @@ test(list_in2,[blocked('The first member/2 creates an infinity of lists => infin
 
 
 rdf(S,P,O) :-
+	ground(P),
 	rdf_l(S,P,O).
 rdf(S,P,O) :-
 	rdf_e(S,P,O).
-rdf(S,P,O) :-
-	rdf_b(S,P,O).
 
 
 rdf_l([_],'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest','http://www.w3.org/1999/02/22-rdf-syntax-ns#nil').

### 2008-06-16 trunk/henry/rdf_e.pl
Handling list transforms (+ associated unit tests)

@@ -1,4 +1,4 @@
-:- module(rdf_e,[rdf_e/3,rdf_b/3]).
+:- module(rdf_e,[rdf_e/3]).
 
 :- use_module(library('semweb/rdf_db'),
               [ rdf_global_id/2,
@@ -9,7 +9,8 @@
                 rdf_subject/1,
                 rdf_equal/2
               ]).
-
+:- use_module(library('semweb/rdfs'),
+              [ rdfs_list_to_prolog_list/2 ]).
 
 :- dynamic rdf_e/3.
 
@@ -37,6 +38,12 @@ rdf_e(S,P,O) :-
 	rdf_query(OO,O,Q2),
 	rdf_db:(Q1,Q2,rdf(SS,P,OO,G)),
 	\+rdf_db:rdf(G,_,_),\+rdf_db:rdf(_,_,G),P\='http://www.w3.org/2000/10/swap/log#implies'.
+rdf_e(S,P,O) :-
+        builtin(P,Pred),
+        wrap_list(S,SS),
+	wrap_list(O,OO),
+	call(Pred,SS,OO).
+
 
 %individual_to_atom(individual(_,Vars),NN) :-
 %	concat_atom(['__bnode_'|Vars],'_',NN),!.
@@ -47,10 +54,11 @@ rdf_query('http://www.w3.org/1999/02/22-rdf-syntax-ns#nil',[],true) :-!.
 rdf_query(LH,[H|T],(rdf(LH,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',H),rdf(LH,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest',LT),QT)) :-
 	rdf_query(LT,T,QT).
 
+wrap_list(S,SS) :-
+	atomic(S),
+	rdfs_list_to_prolog_list(S,SS),!.
+wrap_list(S,S).
 
-%rdf_b(S,P,O) :-
-%	builtin(P,Pred),
-%	call(Pred,S,O).
-%builtin('http://www.w3.org/2000/10/swap/list#in',rdfmember).
-%rdfmember(M,L) :- \+var(L),member(M,L).
+builtin('http://www.w3.org/2000/10/swap/list#in',rdf_e:rdfmember).
+rdfmember(M,L) :- \+var(L),member(M,L).
 

### 2008-06-16 trunk/henry/builtins.pl
Solved an annoying bug in skolemisation Now includes a n3_retract, to retract a loaded n3 file

@@ -50,6 +50,7 @@ getnext(X, X, L) :-
 	last(L, X).
 
 
+
 %Just illustrating a bug
 %builtin('http://purl.org/ontology/swi/append',builtin:append).
 %builtin:append([[a,b],[c]],[a,b,c]).

### 2008-06-16 trunk/henry/n3_to_prolog.pl

@@ -33,9 +33,9 @@ n3_prolog_clause(':-'(H,B)) :-
 	implies(BodyGraph,HeadGraph),
 	n3_pl(HeadGraph,PLH,B1),
 	n3_pl(BodyGraph,PLB,B2),
-	skolemize(PLH,PLB,Skolem),
 	append(B1,B2,Bindings),
 	merge_bindings(Bindings),
+	skolemize(PLH,PLB,Skolem),
 	member(H,Skolem),
 	list_to_conj(PLB,B).
 	

### 2008-06-16 trunk/henry/rdf_e.pl

@@ -11,6 +11,7 @@
               ]).
 :- use_module(library('semweb/rdfs'),
               [ rdfs_list_to_prolog_list/2 ]).
+:- use_module(builtins).
 
 :- dynamic rdf_e/3.
 
@@ -45,9 +46,6 @@ rdf_e(S,P,O) :-
 	call(Pred,SS,OO).
 
 
-%individual_to_atom(individual(_,Vars),NN) :-
-%	concat_atom(['__bnode_'|Vars],'_',NN),!.
-%individual_to_atom(N,N).
 
 rdf_query(N,N,true) :- \+is_list(N).
 rdf_query('http://www.w3.org/1999/02/22-rdf-syntax-ns#nil',[],true) :-!.
@@ -59,6 +57,4 @@ wrap_list(S,SS) :-
 	rdfs_list_to_prolog_list(S,SS),!.
 wrap_list(S,S).
 
-builtin('http://www.w3.org/2000/10/swap/list#in',rdf_e:rdfmember).
-rdfmember(M,L) :- \+var(L),member(M,L).
 

### 2008-06-16 henry/entailment.pl
Solved an annoying bug in skolemisation Now includes a n3_retract, to retract a loaded n3 file

@@ -2,6 +2,8 @@
 
 :- use_module(rdf_e).
 
+
+
 :- begin_tests(entailment).
 test(list_construction,[]) :-
 	rdf_l(A,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',a),
@@ -33,11 +35,29 @@ test(list_in2,[blocked('The first member/2 creates an infinity of lists => infin
         memberchk(D,[a,b,c]).
 :- end_tests(entailment).
 
-
 rdf(S,P,O) :-
+	copy_term((S,O),(SS,OO)),
+	entail(rdf(SS,P,OO),rdf(S,P,O)).
+
+entail(rdf(S,P,O),rdf(SS,P,OO)) :-
+	rdf_2(S,P,O),
+	% some filtering, to not make SeRQL crash with lists-as-nodes and bnodes-as-terms
+	clean(S,SS),
+	clean(O,OO).
+
+clean(individual(B,Vars),NN) :-
+       concat_atom([B|Vars],'_',NN),!.
+%clean(L,N) :-
+%	is_list(L),!,
+%	rdf_
+clean(N,N).
+
+
+
+rdf_2(S,P,O) :-
 	ground(P),
 	rdf_l(S,P,O).
-rdf(S,P,O) :-
+rdf_2(S,P,O) :-
 	rdf_e(S,P,O).
 
 

### 2008-06-16 henry/n3_load.pl

@@ -9,3 +9,20 @@ n3_load(File) :-
         phrase(n3_dcg:document(Base,Doc),Tokens),
         forall(member(rdf(A,B,C,D),Doc),rdf_db:rdf_assert(A,B,C,D)).
 
+n3_retract(File) :-
+	graphs_in_file(File,Graphs),
+	forall(member(Graph,Graphs),rdf_db:rdf_retractall(_,_,_,Graph)).
+
+graphs_in_file(File,T) :-
+	absolute_file_name(File,Absolute),
+	format(atom(Base),'file://~w',[Absolute]),
+	graphs_in(Base,Gs),list_to_set(Gs,T).
+
+
+graphs_in(Graph,[Graph|Graphs]) :-
+	findall([N|T],((rdf_db:rdf(N,_,_,Graph);(rdf_db:rdf(_,_,N,Graph),atomic(N))),graph(N),graphs_in(N,T)),Ts),
+	flatten(Ts,Graphs).
+
+graph(G) :-
+	atom_concat('__bnode_graph',_,G).
+

### 2008-06-16 henry/skolemize.pl

@@ -1,5 +1,7 @@
 :- module(skolemize,[skolemize/3]).
 
+:- use_module(library('semweb/rdf_db'),[rdf_bnode/1]).
+
 /**
  * Some unit tests
  */
@@ -8,7 +10,8 @@ test(skolemize) :-
 	skolemize(
 	[rdf_e(A,has,bnode(Nose)),rdf_e(bnode(Nose),colour,bnode(Colour)),rdf_e(bnode(Colour),label,pink)],
 	[rdf_e(A,isa,man)],
-	[rdf_e(A, has, individual(Nose, [A])), rdf_e(individual(Nose, [A]), colour, bnode(Colour)), rdf_e(individual(Colour, [A]), label, pink)]).
+	Skolem),!,
+	Skolem==[rdf_e(A, has, individual(Nose, [A])), rdf_e(individual(Nose, [A]), colour, individual(Colour, [A])), rdf_e(individual(Colour, [A]), label, pink)].
 :- end_tests(skolemize).
 /**
  * End of unit tests
@@ -16,12 +19,17 @@ test(skolemize) :-
 
 
 skolemize([],_,[]).
+skolemize([rdf_e(S,P,O)|T1],Body,[rdf_e(individual(B,Vars),P,individual(C,Vars))|T2]) :-
+	compound(S),S=bnode(B),((var(B),rdf_bnode(B));true),
+	compound(O),O=bnode(C),((var(C),rdf_bnode(C));true),
+	free_variables(Body,Vars),!,
+	skolemize(T1,Body,T2).
 skolemize([rdf_e(S,P,O)|T1],Body,[rdf_e(individual(B,Vars),P,O)|T2]) :-
-	S=@=bnode(B),S=bnode(B),
+	compound(S),S=bnode(B),((var(B),rdf_bnode(B));true),
 	free_variables(Body,Vars),!,
 	skolemize(T1,Body,T2).
 skolemize([rdf_e(S,P,O)|T1],Body,[rdf_e(S,P,individual(B,Vars))|T2]) :-
-        O=@=bnode(B),O=bnode(B),
+        compound(O),O=bnode(B),((var(B),rdf_bnode(B));true),
 	free_variables(Body,Vars),!,
         skolemize(T1,Body,T2).
 skolemize([H|T1],Body,[H|T2]) :-

### 2008-06-16 henry/n3_load.pl
Some unit testing for load/retract

@@ -1,6 +1,14 @@
-:- module(n3_load,[n3_load/1]).
+:- module(n3_load,[n3_load/1,n3_retract/1]).
 
 :- use_module(n3_dcg).
+
+:- begin_tests(n3_load).
+test(load_and_retract) :-
+	n3_load('examples/uncle.n3'),
+	n3_retract('examples/uncle.n3'),
+	\+rdf_db:rdf(_,_,_).
+:- end_tests(n3_load).
+
 
 n3_load(File) :-
         tokenise(File,Tokens),

### 2008-06-17 trunk/henry/entailment.pl
Fixing publication of lists - still buggy, but much smaller implementation

@@ -1,7 +1,8 @@
 :- module(entailment,[rdf/3]).
 
 :- use_module(rdf_e).
-
+:- use_module(utils).
+:- use_module(library('semweb/rdf_db'),[rdf_bnode/1]).
 
 
 :- begin_tests(entailment).
@@ -33,11 +34,45 @@ test(list_in2,[blocked('The first member/2 creates an infinity of lists => infin
         rdf_l(C,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',c),
         rdf_l(C,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest','http://www.w3.org/1999/02/22-rdf-syntax-ns#nil'),!,
         memberchk(D,[a,b,c]).
+test(describe_list,[]) :-
+	findall(Triple,describe_list([[a],[b]],list,Triple),Triples), % Should perhaps test with rdf_l, but bnodes make it complicated
+	Triples = [
+		rdf(list,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',B)
+	,	rdf(B,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',a)
+	,	rdf(B,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest','http://www.w3.org/1999/02/22-rdf-syntax-ns#nil')
+	,	rdf(list,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest',C)
+	,       rdf(C,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest','http://www.w3.org/1999/02/22-rdf-syntax-ns#nil')
+	,	rdf(C,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',D)
+	,	rdf(D,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',b)
+	,	rdf(D,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest','http://www.w3.org/1999/02/22-rdf-syntax-ns#nil')
+	].
 :- end_tests(entailment).
 
 rdf(S,P,O) :-
-	copy_term((S,O),(SS,OO)),
-	entail(rdf(SS,P,OO),rdf(S,P,O)).
+	copy_term((S,P,O),(SS,PP,OO)),
[…]

### 2008-06-18 henry/entailment.pl
List RDF description now ok

@@ -51,10 +51,12 @@ test(describe_list,[]) :-
 rdf(S,P,O) :-
 	copy_term((S,P,O),(SS,PP,OO)),
 	entail(rdf(SS,PP,OO),rdf(SSS,PPP,OOO)),
+	(is_list(SSS) -> rdf_bnode(SSSS) ; SSS=SSSS),
+	(is_list(OOO) -> rdf_bnode(OOOO) ; OOOO=OOO),
 	(
-		(\+is_list(SSS),\+is_list(OOO),S=SSS,P=PPP,O=OOO);
-		(is_list(SSS),rdf_bnode(List),describe_list(SSS,List,rdf(S,P,O)));
-		(is_list(OOO),rdf_bnode(List),describe_list(OOO,List,rdf(S,P,O)))
+		(S=SSSS,P=PPP,O=OOOO);
+		(is_list(SSS),describe_list(SSS,SSSS,rdf(S,P,O)));
+		(is_list(OOO),describe_list(OOO,OOOO,rdf(S,P,O)))
 	).
 
 describe_list([H|_],S,rdf(S,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',H)) :-

### 2008-06-18 henry/builtins.pl
Oops, forgot to remove the statement illustrating the bug

@@ -52,6 +52,6 @@ getnext(X, X, L) :-
 
 
 %Just illustrating a bug
-builtin('http://purl.org/ontology/swi/append',builtin:append).
-builtin:append([[a,b],[c]],[a,b,c]).
+%builtin('http://purl.org/ontology/swi/append',builtin:append).
+%builtin:append([[a,b],[c]],[a,b,c]).
 

### 2008-06-18 henry/entailment.pl
Cleaning useless stuff

@@ -85,18 +85,8 @@ entail(rdf(S,P,O),rdf(SS,P,OO)) :-
 clean(Bnode,NN) :-
        ground(Bnode), Bnode=individual(B,Vars),
 	concat_atom([B|Vars],'_',NN),!.
-%clean(L,N) :-
-%	is_list(L),!,
-%	associate(N,L).
 clean(N,N).
 
-
-:- dynamic association/2.
-associate(N,L) :-
-	association(N,L),!.
-associate(N,L) :-
-	var(N),rdf_bnode(N),
-	assert(association(N,L)).
 
 rdf_2(S,P,O) :-
 	ground(P),

### 2008-06-18 henry/entailment.pl
SPARQL end-point, corrected a small bug on lists is_list([a|B]) fails...

@@ -23,8 +23,14 @@ test(list_in1,[]) :-
         rdf_l(B,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest',C),
         rdf_l(C,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',c),
         rdf_l(C,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest','http://www.w3.org/1999/02/22-rdf-syntax-ns#nil'),
-	rdf_l(D,'http://www.w3.org/2000/10/swap/list#in',A),!,
+	rdf_e(D,'http://www.w3.org/2000/10/swap/list#in',A),!,
 	memberchk(D,[a,b,c]).
+test(list_in2,[]) :-
+	rdf(E,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',literal(a)),
+	rdf(E,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest','http://www.w3.org/1999/02/22-rdf-syntax-ns#nil'),
+	rdf(A,'http://www.w3.org/2000/10/swap/list#in',E),!,
+	E = [literal(a)],
+	A = literal(a).
 test(list_in2,[blocked('The first member/2 creates an infinity of lists => infinite loop. This test should succeed as list_in1.')]) :-
 	rdf_l(D,'http://www.w3.org/2000/10/swap/list#in',A),
 	rdf_l(A,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',a),
@@ -48,12 +54,14 @@ test(describe_list,[]) :-
 	].
 :- end_tests(entailment).
 
+
 rdf(S,P,O) :-
 	copy_term((S,P,O),(SS,PP,OO)),
 	entail(rdf(SS,PP,OO),rdf(SSS,PPP,OOO)),
 	(is_list(SSS) -> rdf_bnode(SSSS) ; SSS=SSSS),
-	(is_list(OOO) -> rdf_bnode(OOOO) ; OOOO=OOO),
+	(is_list(OOO) -> rdf_bnode(OOOO) ; OOO=OOOO),
 	(
+		((is_list(SSS);is_list(OOO)),S=SSS,P=PPP,O=OOO);
 		(S=SSSS,P=PPP,O=OOOO);
 		(is_list(SSS),describe_list(SSS,SSSS,rdf(S,P,O)));
[…]

### 2008-06-18 henry/rdf_e.pl

@@ -47,7 +47,7 @@ rdf_e(S,P,O) :-
 
 
 
-rdf_query(N,N,true) :- \+is_list(N).
+rdf_query(N,N,true) :- \+N=[_|_].
 rdf_query('http://www.w3.org/1999/02/22-rdf-syntax-ns#nil',[],true) :-!.
 rdf_query(LH,[H|T],(rdf(LH,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',H),rdf(LH,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest',LT),QT)) :-
 	rdf_query(LT,T,QT).

### 2008-06-18 trunk/henry/builtins.pl
SPARQL end-point, corrected a small bug on lists is_list([a|B]) fails...

@@ -52,6 +52,6 @@ getnext(X, X, L) :-
 
 
 %Just illustrating a bug
-%builtin('http://purl.org/ontology/swi/append',builtin:append).
-%builtin:append([[a,b],[c]],[a,b,c]).
+builtin('http://purl.org/ontology/swi/append',builtin:append).
+builtin:append([[a,b],[c]],[a,b,c]).
 

### 2008-06-18 trunk/henry/henry.pl

@@ -1,7 +1,22 @@
-:- module(henry,[]).
+#!/usr/local/bin/pl -g serql_welcome -L128m -G128m -T128m -s
+
+:- load_files([ 'SeRQL/load'
+              ],
+              [ silent(true)
+              ]).
+
+:- use_module(entailment).
 
 :- [n3_load].
 :- [n3_to_prolog].
 :- [entailment].
+
+server :-
+        serql_server([]).
+
+server(Port) :-
+        serql_server([port(Port)]).
+
+:- server(1212).
 
 

### 2008-06-18 henry/n3_to_prolog.pl
Test for RDF nodes was buggy

@@ -27,7 +27,7 @@
 
 
 compile_all :-
-	forall(n3_prolog_clause(Clause),(format(user_error,'~w\n',[Clause]),assert(entailment:Clause))).
+	forall(n3_prolog_clause(Clause),(format(user_error,'~w\n',[Clause]),assert(rdf_e:Clause))).
 
 n3_prolog_clause(':-'(H,B)) :-
 	implies(BodyGraph,HeadGraph),

### 2008-06-18 henry/rdf_e.pl

@@ -37,7 +37,7 @@ test(query_list) :-
 rdf_e(S,P,O) :-
 	rdf_query(SS,S,Q1),
 	rdf_query(OO,O,Q2),
-	rdf_db:(Q1,Q2,rdf(SS,P,OO,G)),
+	rdf_db:(Q1,Q2,rdf(SS,P,OO,G)),G\=_:_,
 	\+rdf_db:rdf(G,_,_),\+rdf_db:rdf(_,_,G),P\='http://www.w3.org/2000/10/swap/log#implies'.
 rdf_e(S,P,O) :-
         builtin(P,Pred),
@@ -47,7 +47,7 @@ rdf_e(S,P,O) :-
 
 
 
-rdf_query(N,N,true) :- \+N=[_|_].
+rdf_query(N,N,true) :- atomic(N);var(N);N=@=literal(_). % is a RDF node?
 rdf_query('http://www.w3.org/1999/02/22-rdf-syntax-ns#nil',[],true) :-!.
 rdf_query(LH,[H|T],(rdf(LH,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',H),rdf(LH,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest',LT),QT)) :-
 	rdf_query(LT,T,QT).

### 2008-06-18 front-end/transform.pl
Transform interpreation rules to automate plugin execution

@@ -1,61 +1,186 @@
 /**
-	General view of signal transforms
-	David Pastor Escuredo, c4dm
+	General view of signal transforms (so far LADSPA effects and VAMP features over audio input)
+	David Pastor Escuredo, c4dm, Queen Mary University of Lonfon
 	May 2008
 
-	Contains hardcoded rules for interpretation of a transform predicate
+	Contains hardcoded rules for interpretation of a transform predicate. 
 */
 :-module(transform, [api/1
-		,	plugin_engine/1
-			]).
+		,	transform_type/1
+		, 	transform/2
+		,	apply_transform/3
+		,	engine/1	
+		]).
 
 
-:-use_module(../feature-extractor/fe).
-:-use_module(../swivamp/vamp).
-:-use_module(../swiladspa/ladspa).
-//load effects
+:-use_module('../feature-extractor/fe').
+:-use_module('../swivamp/vamp').
+:-use_module('../swiladspa/ladspa').
 :-style_check(-discontiguous).
 
-/*'Transform'(TransformType, 'Engine'(API, PluginName, Version), SampleRate, StepSize, BlockSize, ListOfParameters, Program).*/
+/*transform(TransformType, engine(API, PluginName, Version), SampleRate, StepSize, BlockSize, ListOfParameters, Program).*/
 
 
 /* Hardcoded information */
-transform_type('feature_extraction').
+transform_type('feature').
 transform_type('effect').
 api('vamp').
 api('ladspa').
+/** running time database of transforms transform_id(ID, TransformFunctor) **/
+:- dynamic transform_id/2.
+
+/**
+	Available transforms (=outputs from plugins), returns the number engine for it. New APIs should be added here.
+	transform(?Transform, ?Engine)
+**/
+transform(Transform, engine(API, PluginName, Version)):-	
+	vamp_feature_system(Transform), 
+	API = 'vamp',
+	vamp_plugin_for(PluginName, Transform, _Index), 
+	vamp_plugin_version(PluginName, Version).
+transform(Transform, engine(API, PluginName, _Version)):-
+	ladspa_effect_system(Transform),
+	API = 'ladspa',
+	ladspa_plugin_for(PluginName, Transform, _Port).
+
+
+/**
+	APPLY TRANSFORM
+
+**/
[…]

### 2008-06-18 trunk/swiladspa/ladspa.pl
Transform interpreation rules to automate plugin execution

@@ -50,7 +50,7 @@
 
 	ldpl_input_audio(-PluginName, -ListOfIndexesOfInputAudioPorts) Eg: ldpl_input_audio('AmplifierStereo', [0, 1])
 
-	ldpl_output_audio(-PluginName, -ListOfIndexesOfOutputudioPorts)
+	ldpl_output_audio(-PluginName, -ListOfIndexesOfOutputAudioPorts)
 
 **/
 
@@ -99,6 +99,8 @@
 /**
 	Non-deterministic scan of plugins in system
 	ldpl_plugin_system(?Plugin)
+
+	IS FAILING AGAIN!!!!
 	*/
 ldpl_plugin_system(Plugin):-
 	ldpl_plugins(List),
@@ -110,6 +112,7 @@ ldpl_plugin_system(Plugin):-
 	ldpl_port_description(?PluginName, 'LadpsaPort'(?Type, ?Index, -Name))
 */
 ldpl_port_description(Name, 'LadspaPort'(T, I, N)):-
+	var(N),
 	ldpl_plugin_system(Name),
 	ldpl_input_audio(Name, InAu),
 	ldpl_output_audio(Name, OutAu),
@@ -129,7 +132,36 @@ is_inC(T, InC, I):-
 is_outC(T, OutC, I):-
 	member(I, OutC),
 	T = 'OutputControl'.
+ldpl_port_descriptor(Name, 'LasdpaPort'(T, I, N)):-
+	nonvar(N),
+	ldpl_port_descriptor(Name, 'LasdpaPort'(T, I, N2)),
+	N=N2.
+	
 
+/**
+	LADSPSA EFFECTS
+*/
+ladspa_plugin_effects(PluginName, Effects):-
+	findall(Effect,	ldpl_port_description(PluginName, 'LadspaPort'(_, _, Effect)), Effects).
+
+ladspa_effect_system(Effect):-
+	ldpl_plugin_system(PluginName),
+	ladspa_plugin_effects(PluginName, Effects),
+	member(Effect, Effects).
+
+ladspa_plugin_for(PluginKey, Effect, Index):-
+	var(Effect),
+	var(Index),
+	vamp_plugin_system(PluginKey),
+	vmpl_port_descriptor(PluginKey, 'LadspaPort'(_, Index, Effect)).
+ladspa_plugin_for(PluginKey, Effect, Index):-
+	nonvar(Effect),
+	var(PluginKey),
[…]

### 2008-06-18 trunk/swivamp/vamp.pl

@@ -101,7 +101,8 @@
 	*/
 
 /** 
-	vamp_plugins_sytem(?PluginKey) This predicates checks if the Plugin is on the system (PluginKey nomenclature description is library:pluginName)
+	vamp_plugins_sytem(?PluginKey) This predicates checks if the
+	Plugin is on the system (PluginKey nomenclature description is library:pluginName)
 	This predicate is non-deterministic
 	*/
 
@@ -119,15 +120,21 @@ vamp_feature_system(FeatureType):-
 	member(FeatureType, AvailableFeatures).
 
 /**
-	vamp_plugin_for(?PluginKey, -FeatureType, ?Index): Non deterministic relationship among a plugin, the features it ouptus and the index of
+	vamp_plugin_for(?PluginKey, ?FeatureType, ?Index): Non deterministic relationship among a plugin, the features it ouptus and the index of
 	each of them in the descriptor
 	*/
 
 vamp_plugin_for(PluginKey, FeatureType, Index):-
+	var(FeatureType),
 	vamp_plugin_system(PluginKey),
 	vmpl_load_plugin(PluginKey, 44100, Plugin),
 	vmpl_plugin_numberOutputs(Plugin, Size),
 	vamp_plugin_features(Plugin, Size, FeatureType, Index).
+vamp_plugin_for(PluginKey, FeatureType, Index):-
+	nonvar(FeatureType),
+	var(PluginKey),
+	findall(FT, vamp_plugin_for(PluginKey, Output, _), Outputs),
+	member(FeatureType, Outputs).
 
 vamp_plugin_features(Plugin, Size, FeatureType, Index):-
 	Limit is Size-1,

### 2008-06-18 front-end/transform.pl

@@ -56,15 +56,27 @@ apply_transform(Input, transform(Type, engine(API, PluginName, Version), Sr, Ste
 	%check_sample_rate(Sr),
 	load_engine(engine(API, PluginName, Version), Plugin, Sr)
 	set_engine(API, Plugin, ListOfParameters, Configuration, Program),
+	restrict_output(API, Output),
 	engine_process(API, Plugin, Step, Block, Indexes, Outputs).
 /**should run in one**/
 	
+/*output syntax*/
+restrict_output('vamp', Output):-
+	Output = [['Feature'(Type, Timestamp, Data)|_SameType]|_MoreOutputsOfThePlugin]
+
+restrict_output('ladspa', Output):-
+	Output = signal(_Ch, _Sr , _L, _Data); Output=frequencySignal(_Ch, _Length, _Data2).
 
 /**
-	Processing+framing
+	Processing+framing. Indexes as list better i think
 */
 engine_process('vamp', Plugin, Step, Block, Indexes, Output):-
-	
+	restrict_output
+	set_blockSize(Plugin, BlockSize),
+	set_stepSize(Plugin, StepSize),
+	get_channels(Signal, Channels),
+	vmpl_initialize_plugin(Plugin, Channels, StepSize, BlockSize),
+	vamp_compute_feature(Signal, StepSize, BlockSize, Indexes, Plugin, WholeFeature).
 
 /**
 	setting up

### 2008-06-18 henry/rdf_e.pl
Trying to fix a bug with literal-as-nodes, but failed Falling back to in-memory caching - I need a second persistent store for avoiding deadlocks (tweaking sparql.pl with transactions doesn't work)

@@ -12,6 +12,8 @@
 :- use_module(library('semweb/rdfs'),
               [ rdfs_list_to_prolog_list/2 ]).
 :- use_module(builtins).
+:- use_module(namespaces).
+:- use_module(utils).
 
 :- dynamic rdf_e/3.
 
@@ -43,18 +45,48 @@ rdf_e(S,P,O) :-
         builtin(P,Pred),
         wrap_list(S,SS),
 	wrap_list(O,OO),
+	call_builtin(P,Pred,SS,OO).
+% Use rdf_reachable for handling owl:sameAs
+
+call_builtin(P,Pred,SS,OO) :-
+	\+tabled(P),
 	call(Pred,SS,OO).
+call_builtin(P,Pred,SS,OO) :-
+	tabled(P), \+cache(SS,P,OO),!,
+	call(Pred,SS,OO),
+	assert(cache(SS,P,OO)).
+	%rdf_assert_g(SSS,SS,G1),
+	%rdf_assert_g(OOO,OO,G2),
+	%literal_to_node(SSS,SSSS),
+	%rdf_db:(rdf_assert(SSSS,P,OOO),G1,G2).
+call_builtin(P,_,SS,OO) :-
+	tabled(P), writeln('looking up in cache'),
+	cache(SS,P,OO).
+
+:- dynamic cache/3.
+cache(S,P,O) :-
+	rdf_db:rdf(SS,P,OO,cache),
+	%literal_to_node(SSS,SS),
+	wrap_list(SS,S),wrap_list(OO,O).
+	
 
 
-
-rdf_query(N,N,true) :- atomic(N);var(N);N=@=literal(_). % is a RDF node?
+rdf_query(N,N,true) :- \+is_list(N),(atomic(N);var(N);(compound(N),N=literal(_))). % is a RDF node?
 rdf_query('http://www.w3.org/1999/02/22-rdf-syntax-ns#nil',[],true) :-!.
 rdf_query(LH,[H|T],(rdf(LH,'http://www.w3.org/1999/02/22-rdf-syntax-ns#first',H),rdf(LH,'http://www.w3.org/1999/02/22-rdf-syntax-ns#rest',LT),QT)) :-
 	rdf_query(LT,T,QT).
+rdf_assert_g(N,N,true) :- \+is_list(N),(atomic(N);var(N);(compound(N),N=literal(_))). % is a RDF node?
[…]

### 2008-06-18 henry/utils.pl

@@ -1,8 +1,13 @@
-:- module(utils,[list_to_conj/2]).
+:- module(utils,[list_to_conj/2,literal_to_node/2]).
 
 
 list_to_conj([H],H) :- !.
 list_to_conj([H,T],(H,T)) :- !.
 list_to_conj([H|T],(H,T2)) :-
         list_to_conj(T,T2).
+
+% subject as literals hack...
+literal_to_node(Literal,Node) :- var(Literal), atom_concat('__literal',_,Node),atom_concat('__',At,Node),term_to_atom(Literal,At),!.
+literal_to_node(At,At) :- atomic(At).
+literal_to_node(literal(Lit),Node) :- var(Node), term_to_atom(literal(Lit),At), atom_concat('__',At,Node).
 

### 2008-06-18 trunk/henry/n3_load.pl
Trying to fix a bug with literal-as-nodes, but failed Falling back to in-memory caching - I need a second persistent store for avoiding deadlocks (tweaking sparql.pl with transactions doesn't work)

@@ -1,9 +1,11 @@
 :- module(n3_load,[n3_load/1,n3_retract/1]).
 
 :- use_module(n3_dcg).
+:- use_module(utils).
 
 :- begin_tests(n3_load).
 test(load_and_retract) :-
+	rdf_db:rdf_reset_db, % hem... must be careful about that
 	n3_load('examples/uncle.n3'),
 	n3_retract('examples/uncle.n3'),
 	\+rdf_db:rdf(_,_,_).
@@ -15,7 +17,7 @@ n3_load(File) :-
         absolute_file_name(File,Absolute),
         format(atom(Base),'file://~w',[Absolute]), % That's probably not right
         phrase(n3_dcg:document(Base,Doc),Tokens),
-        forall(member(rdf(A,B,C,D),Doc),rdf_db:rdf_assert(A,B,C,D)).
+        forall(member(rdf(A,B,C,D),Doc),(literal_to_node(A,AA),rdf_db:rdf_assert(AA,B,C,D))).
 
 n3_retract(File) :-
 	graphs_in_file(File,Graphs),

### 2008-06-18 trunk/henry/namespaces.pl

@@ -14,5 +14,5 @@
 :- rdf_register_ns(to,'http://purl.org/ontology/tonality/').
 :- rdf_register_ns(tl,'http://purl.org/NET/c4dm/timeline.owl#').
 :- rdf_register_ns(event,'http://purl.org/NET/c4dm/event.owl#').
-
+:- rdf_register_ns(t,'http://purl.org/ontology/tabling/').
 

### 2008-06-19 front-end/transform.pl
Transform interpretation rules working for VAMP. Strong declarative meaning of plugin-outputs relationships. tempotracker example

@@ -3,12 +3,14 @@
 	David Pastor Escuredo, c4dm, Queen Mary University of Lonfon
 	May 2008
 
-	Contains hardcoded rules for interpretation of a transform predicate. 
+	Contains hardcoded rules for interpretation of a transform predicate. These rules allow the user to call the transform in *many* ways without crashing pointing to a declarative meaning of such a predicate. 
 */
 :-module(transform, [api/1
 		,	transform_type/1
 		, 	transform/2
 		,	apply_transform/3
+		,	apply_store_transform/4
+		,	my_transform/2
 		,	engine/1	
 		]).
 
@@ -16,21 +18,27 @@
 :-use_module('../feature-extractor/fe').
 :-use_module('../swivamp/vamp').
 :-use_module('../swiladspa/ladspa').
+:-use_module('../swiaudiosource/audiosource').
 :-style_check(-discontiguous).
 
-/*transform(TransformType, engine(API, PluginName, Version), SampleRate, StepSize, BlockSize, ListOfParameters, Program).*/
+
+/*************************************************************
+	  QUERYING PREVIOUS INFORMATION
+*************************************************************/
 
 
-/* Hardcoded information */
+/** Hardcoded information **/
 transform_type('feature').
 transform_type('effect').
 api('vamp').
 api('ladspa').
+
 /** running time database of transforms transform_id(ID, TransformFunctor) **/
 :- dynamic transform_id/2.
+:- dynamic computation_id/4.
 
 /**
-	Available transforms (=outputs from plugins), returns the number engine for it. New APIs should be added here.
+	AVAILABLE TRANSFORMS (=outputs from plugins), returns the engine for it. New APIs should be added here.
 	transform(?Transform, ?Engine)
 **/
 transform(Transform, engine(API, PluginName, Version)):-	
@@ -43,64 +51,124 @@ transform(Transform, engine(API, PluginName, _Version)):-
 	API = 'ladspa',
 	ladspa_plugin_for(PluginName, Transform, _Port).
 
+/**Querying the possible engines**/
+engine(engine(API, PluginName, Version)):-
+	(vamp_plugin_system(PluginName), API ='vamp', vamp_plugin_version(PluginName, Version));
[…]

### 2008-06-19 swivamp/vamp.pl

@@ -33,6 +33,7 @@
 	,	vamp_feature_system/1
 	,	vamp_plugin_for/3
 	, 	vamp_plugin_numberOutputs/2
+	,	vamp_plugin_features/2
 
 		/** 2nd step lifecycle: loading. We need the sample rate*/			
 	,	vmpl_load_plugin/3
@@ -121,9 +122,8 @@ vamp_feature_system(FeatureType):-
 
 /**
 	vamp_plugin_for(?PluginKey, ?FeatureType, ?Index): Non deterministic relationship among a plugin, the features it ouptus and the index of
-	each of them in the descriptor
+	each of them in the descriptor. Quite powerful.
 	*/
-
 vamp_plugin_for(PluginKey, FeatureType, Index):-
 	var(FeatureType),
 	vamp_plugin_system(PluginKey),
@@ -131,11 +131,26 @@ vamp_plugin_for(PluginKey, FeatureType, Index):-
 	vmpl_plugin_numberOutputs(Plugin, Size),
 	vamp_plugin_features(Plugin, Size, FeatureType, Index).
 vamp_plugin_for(PluginKey, FeatureType, Index):-
+	nonvar(PluginKey),
 	nonvar(FeatureType),
+	is_feature_of_plugin(PluginKey, FeatureType, Index).
+vamp_plugin_for(PluginKey, FeatureType, Index):-
 	var(PluginKey),
-	findall(FT, vamp_plugin_for(PluginKey, Output, _), Outputs),
-	member(FeatureType, Outputs).
-
+	var(Index),
+	nonvar(FeatureType),
+	vamp_plugin_system(Plugin),
+	is_feature_of_plugin(Plugin, FeatureType, Index),
+	PluginKey = Plugin.
+/** 
+	is_feature_of_plugin(+Plugin, ?Feature, ?Index)  and predicates helping for vamp_plugin_for/3 
+**/
+is_feature_of_plugin(Plugin, Feature, Index):-
+	nonvar(Plugin),
+	vamp_plugin_features(Plugin, List),
+	member([Feature,  Index], List).
+vamp_plugin_features(Plugin, OutIn):-
+	nonvar(Plugin),
+	findall([Output, In], vamp_plugin_for(Plugin, Output, In), OutIn).
 vamp_plugin_features(Plugin, Size, FeatureType, Index):-
 	Limit is Size-1,
 	between(0, Limit, Index),

### 2008-06-20 sandpit/n3.pl
Some code I wrote a year ago for a n3 repl

@@ -1,0 +1,219 @@
+:- module(n3,[
+		parse/3
+	,	load/2
+	,	interp/0
+	,	interp/1
+	]).
+
+/**
+ * Start of a "N3 interpreter"
+ *
+ * Yves Raimond, C4DM, QMUL, (c) 2007
+ */
+
+
+:- use_module(library('semweb/rdf_db')).
+:- use_module('../n3/n3_dcg').
+:- use_module('namespaces').
+:- use_module(library('semweb/rdfs')).
+
+/**
+ * Some Prolog wrappers for common N3 constructs
+ */
+
+:- op(1110,xfy,'<=').
+
+HeadGraph <= TailGraph :- 
+	rdf(If,owl:inverseOf,log:implies),
+	rdf(HeadGraph,If,TailGraph).
+HeadGraph <= TailGraph :-
+	rdf(TailGraph,log:implies,HeadGraph).
+
+uqv(UQVar) :-
+	atomic(UQVar),
+	atom_concat(_,'_uqvar',UQVar).
+	
+
+/**
+ * N3 interpreter - top-level predicates
+ */
+
+repl :- interp('').
+interp :-
+	interp('').
+interp(BaseURI) :- 
+	n3_dcg:turtle_tokens(user_input,Tokens),
+	phrase(document(BaseURI,N3),Tokens),
+	writeln(N3),
+	prove(N3,Binding),(writeln(Binding)).
+
+/**
+ * Proof mechanims, 2nd attempt
+ */
+graph(Graph) :-
+	rdf(_,_,_,Graph),!.
+
+pl_clause(Clause,Bindings) :-
+	(A <= B),
+	n3_clause_to_pl_clause((A <=B),Clause,Bindings).
+n3_clause_to_pl_clause('<='(Graph1,Graph2), clause(graph(TriplesHead),graph(TriplesBody)),Bindings) :-
+	graph_triples(Graph1,Triples1),
+	graph_triples(Graph2,Triples2),
+	clean_vars(Triples1,TriplesHead,Bindings1),
+	clean_vars(Triples2,TriplesBody,Bindings1,Bindings).
+	%clean_lists(T1,TriplesHead),
+	%clean_lists(T2,TriplesBody).
+
+
+graph_triples(GraphURI,Triples) :-
+	bagof(triple(S,P,O),rdf(S,P,O,GraphURI),Triples).
+
[…]

### 2008-06-20 trunk/henry/persist.pl
New tests, and new persistency mechanism

@@ -1,0 +1,158 @@
+:- module(persist,[init_persistency/0,commit/0,mem_load/1,bin_db/1,assert_tabled/1,just_tabled/1,cached/1,tabled/1,persist/1,s_table/1]).
+
+
+:- use_module(library('semweb/rdf_db')).
+:- use_module('../swiaudiodata/audiodata').
+:- use_module(rdf_e).
+:- use_module(namespaces).
+
+:- begin_tests(persist).
+test(commit_triple) :-
+	assert(rdf_e:cache(a,b,c)),
+	commit,!,
+	rdf_db:rdf(a,b,c),!,
+	rdf_db:rdf_reset_db. % Ouch
+:- end_tests(persist).
+
+bin_db(bindb).
+
+:- 
+	bin_db(DB),
+	(exists_directory(DB) -> true; make_directory(DB)).
+
+
+tabled(Predicate) :-
+	rdf_db:rdf(Predicate,rdf:type,t:'TabledPredicate');
+	rdf_db:rdf(Predicate,rdf:type,t:'CachedPredicate').
+just_tabled(Predicate) :-
+	rdf_db:rdf(Predicate,rdf:type,t:'TabledPredicate').
+cached(Predicate) :-
+	rdf_db:rdf(Predicate,rdf:type,t:'CachedPredicate').
+:- rdf_meta assert_tabled(r).
+assert_tabled(Predicate) :-
+	rdf_db:rdf_assert(Predicate,rdf:type,t:'TabledPredicate').
+
+
+persist_l([]).
+persist_l([H|T]) :-
+	persist_l(H),
+	persist_l(T).
+persist_l(H) :- persist_node(H).
+persist_l(_).
+clean_l([]).
+clean_l([H|T]) :-
+        clean_l(H),
+        clean_l(T).
+clean_l(H) :- clean_node(H).
+clean_l(_).
+
+
+:- multifile list/2.
+
+commit :- 
+	forall(cache(S,P,O),
+		(
+			persist(rdf(S,P,O)),
+			((pl_list_to_rdf_list(S,Triples1,SS),list(SS,S),!);(S=SS,Triples1=[])),
+			((pl_list_to_rdf_list(O,Triples2,OO),list(OO,O),!);(O=OO,Triples2=[])),
+			append(Triples1,Triples2,Triples),
+			free_variables([Triples,SS,OO],Vars),bnode_list(Vars),
+			!,rdf_assert(SS,P,OO,persist),
[…]

### 2008-06-20 henry/rdf_e.pl
New tests, and new persistency mechanism

@@ -1,4 +1,4 @@
-:- module(rdf_e,[rdf_e/3]).
+:- module(rdf_e,[rdf_e/3,cache/3]).
 
 :- use_module(library('semweb/rdf_db'),
               [ rdf_global_id/2,
